\heading{2}{Jonah is swallowed by a great fish~--- he prays to the Lord and repents~--- Jonah is forgiven and spit out onto the dry land}

\begin{inparaenum}
  \verse{2:1} \ed{In English versions, this is \vref{Jon}{1}{17} and the following verses are \vref{Jon}{2}{1--10}. This version follows the Hebrew numbering scheme.}The \textsc{Lord} had prepared\halot{xxxx}{xxxx} a great fish\halot{xxxx}{xxxx} to swallow Jonah, and Jonah was in the belly of the fish for three days and three nights.%%
  \verse{2:2} Jonah prayed to the \textsc{Lord} his God from the belly of the fish,%%
  \verse{2:3} and said,%%
  
  \pvdb{``I cried because of my afflictions\footnotemark}{and the \textsc{Lord} answered me.}%%
  \fntalt{in my despair/distress}%%
  
  \pvdb{I cried for help from the belly of Sheol}{and He heard my voice.}%%
  
  \pvcb{\vn{2:4} You cast me into the deep~--- into the midst of the sea~---}{and the rushing\footnotemark\ waters surrounded me.}%%
  \fnthalot{xxxx}{xxxx deep?}%%
  
  \pvdb{All the billows and waves}{passed over me}%%
  
  \pvcb{\vn{2:5} and I said, `I am cast}{out of Your sight,}%%
  
  \pvdb{yet I will again look}{at Your holy temple.'\footnotemark}%%
  \fntlit{the temple of Your holiness}%%
  
  \pvcb{\vn{2:6} The waters encompassed me to my neck,}{the deep sea engulfed me.}%%
  
  \pvdb{The reeds\footnotemark\ were wrapped around my head}{\vn{2:7} I went down to the base of the mountains.}%%
  \fnthalot{xxxx}{xxxx seaweed?}%%
  
  \pvdb{The land, its bars\footnotemark}{came up to me forever.}%%
  \fntie{prison}%%
  
  \pvdb{But You brought up my life from destruction,}{O \textsc{Lord} my God.}%%
  
  \pvcb{\vn{2:8} When my soul was faint}{I remembered the \textsc{Lord}.}%%
  
  \pvdb{My prayer came to You,}{to Your holy temple.}%%
  
  \pvcb{\vn{2:9} xxxx}{xxxx}%%
  
  \pvcb{\vn{2:10} With a voice of thanksgiving}{I will sacrifice to You.}%%
  % xxxx Let me sacrifice to you with a voice of thanksgiving?    
  
  \pvdb{xxxx}{xxxx''}%%
  
  \verse{2:11} The \textsc{Lord} spoke to the fish and it vomited\alt{spit} Jonah onto the dry land.%%
\end{inparaenum}
