\heading{15}{Samuel commands Saul to annihilate the Amalekites~--- Saul and the people disobey God, spare Agag, and bring back some of the animals to sacrifice~--- Saul is sharply reprimanded by Samuel, denies any wrongdoing, and is stripped of his royalty}

\begin{inparaenum}
    \verse{15:1} And Samuel spake unto Saul, ``The \textsc{Lord} didst send me to anoint thee king over His people~--- over Israel~--- therefore, now, hearken unto the voice\footnote{sound} of the words of the \textsc{Lord}:%%
    \verse{15:2} Thus saith the \textsc{Lord} of Hosts, `I have reviewed\footnote{looked after} that which Amalek did to Israel, that which he placed for himself in the way going up out of Egypt.%%
    \verse{15:3} Go now, and smite Amalek and all that he hath; show no mercy\footnote{have no pity on them}. Put to death every man, woman, infant, suckling, ox, sheep, camel, and ass.'''%%
    \verse{15:4} So Saul summoned the people and inspected them in Telaim: 200,000 footmen and 10,000 men of Judah.%%
    \verse{15:5} Then Saul came unto the city of Amalek and waited in a valley.%%
    \verse{15:6} And Saul saith to the Kenite, ``Go, turn aside. Leave from the midst of Amalek lest I consume thee as well, for thou didst show kindness unto the children of Israel when they came up out of Egypt.'' So the Kenite turned aside from the midst of Amalek.%%
    \verse{15:7} Then Saul smote Amalek from Havilah to Shur (on the borders of Egypt).%%
    \verse{15:8} He caught Agag, king of Amalek, alive; but all the people were destroyed by the sword.%%
    \verse{15:9} But Saul~--- and the people as well~--- had pity on Agag, the best of the flocks, herds, garments, rams, and all that was beautiful and they would not destroy them.%%
    \verse{15:10} Then the word of the \textsc{Lord} came unto Samuel, saying,%%
    \verse{15:11} ``I'm sorry that I made Saul king: he hath forsaken Me in that he did not keep My commandments.'' And Samuel was grieved and cried unto the \textsc{Lord} all night.%%
    \verse{15:12} So Samuel rose early in the morning to meet Saul. And it was declared to Samuel, saying, ``Saul hath come in to Carmel and is setting up to himself a monument. Then he shall go round, pass over, and go down into Gilgal.''%%
    \verse{15:13} Samuel came to Saul, and Saul saith to him, ``Blessed be thou of the \textsc{Lord}; I have done what the \textsc{Lord} hath said.''%%
    \verse{15:14} So Samuel said, ``Why then do I hear bleating of sheep and the noise of a herd?''%%
    \verse{15:15} And Saul said, ``We brought them from Amalek because the people had pity on the best of the flocks and herds so that we could sacrifice them to the \textsc{Lord} thy God; the rest we destroyed.''%%
    \verse{15:16} Samuel said to Saul, ``Hold on,\footnote{Relax, wait} let me tell you what the \textsc{Lord} told me tonight.'' And Saul said, ``Go on.''%%
    \verse{15:17} Samuel said, ``When you were little\footnote{in your own eyes}, were you not made head of the tribes of Israel and anointed by the \textsc{Lord} to be king over Israel?%%
    \verse{15:18} And the \textsc{Lord} sent you on a journey and said, `Go and utterly destroy the Amalekites and all that they hath.'%%
    \verse{15:19} Why then did you not obey the voice of the \textsc{Lord}, but instead did fly on their spoil and do evil in the eyes of the \textsc{Lord}?''%%
    \verse{15:20} And Saul said to Samuel, ``I \emph{have} hearkened to the voice of the \textsc{Lord}. I went the way the \textsc{Lord} sent me, and I have brought Agag, king of Amalek, and destroyed the Amalekites.%%
    \verse{15:21} And the people took the spoil of the flocks and herds\footnote{Idiomatically rendered in definite plural}~--- the choicest\footnote{chief, first, best} part of the devoted things\footnote{i.e., that which should have been destroyed}~--- to sacrifice unto the \textsc{Lord} their God in Gilgal.''%%
    \verse{15:22} And Samuel saith:
    
:``Doth the \textsc{Lord} delight in sacrifices and burnt offerings So much as He delighteth\footnote{Poetic: therefore the verb's repetition is implied.} in hearkening to the voice\footnote{Implied: commandments} of the \textsc{Lord}?

:Listen! Obedience is better than sacrifice\footnote{Interesting Hebrew grammatical construct here. It is literally ``Obedience sacrifice good.'' That is how comparatives are done.}. Paying attention is more important than ram's fat.%%
    \verse{15:23} :Since rebellion is like the sin of witchcraft And stubbornness is like idolatry:

:Because you've rejected the word of the \textsc{Lord} He hath you from being king.%%
    \verse{15:24} Then Saul said to Samuel, ``I have sinned, for I transgressed the commandment of the \textsc{Lord}\footnote{Literally, I have passed over (or by) the command (or mouth) of the \textsc{Lord}} and thy words\footnote{Repetitious because they're the same in this instance.\footnote{cf. DyC 1:38} Shows a serious lack of understanding on Saul's part.} in that I showed reverence to\footnote{feared} the people by hearkening\footnote{listening} to their voice.\footnote{i.e., instead of the \textsc{Lord}}%%
    \verse{15:25} Please forgive me\footnote{Literally, Now, please bear my sin} and come with me as I bow down before the \textsc{Lord}.''%%
    \verse{15:26} Samuel said to Saul, ``No, I'm not going with you\footnote{Literally, I will not turn back with thee} because you've rejected the word of the \textsc{Lord} and because the \textsc{Lord} hath rejected you from being king over Israel.''%%
    \verse{15:27} Then Samuel turned around to go, but he\footnote{Saul} caught hold of his\footnote{Samuel's} robe's mantle\footnote{upper skirt} and it rent.%%
    \verse{15:28} And Samuel said to him, ``The \textsc{Lord} hath rent the Kingdom of Israel from thee today and given it to thy neighbor (who is better than thee).%%
    \verse{15:29} Additionally, the eminence\footnote{preeminence, perpetuity, everlastingness} of Israel\footnote{Is this a name-title of the \textsc{Lord}?} neither lies nor repents for He is not a man, that He is penitent.''\footnote{repents, is repentant}%%
    \verse{15:30} And he said, ``I have sinned. Come, now, and honor me before the elders of the people\footnote{This is horribly pretentious on Saul's part: if Samuel wouldn't bow down to the \textsc{Lord} with Saul, why in the world would he honor him before the elders of the people and all of Israel? It's like Saul doesn't think before he speaks.} and before Israel. Come with me as I bow down before the \textsc{Lord}.''%%
    \verse{15:31} So Samuel turned back after Saul as Saul bowed before the \textsc{Lord}.\footnote{I know, I'm as shocked as you.}%%
    \verse{15:32} And Samuel said, ``Bring Agag, king of Amalek, unto me.'' And Agag came in unto him delicately and said, ``Certainly the bitterness of death has past.''%%
    \verse{15:33} Samuel said, ``As thy sword hath bereaved women, even so shall thy mother be bereaved among women,'' and Samuel hewed Agag into pieces before the \textsc{Lord} in Gilgal.%%
    \verse{15:34} Samuel went to Ramath and Saul went to his house in Gibeah of Saul.%%
    \verse{15:35} Samuel never again came to see Saul, even until his death; nevertheless, Samuel mourned for Saul. And the \textsc{Lord} was sorry that he had made Saul king over Israel.%%
\end{inparaenum}
