\heading{1}{Those who follow the Lord are blessed and happy~--- the Lord knows the way}

\begin{inparaenum}
    \pvac{\vn{1:1} Happy\footnotemark\ is the man}{who does not walk}{in the counsel of the wicked.}%%
    \fntalt{Blessed}
    
    \pvbb{Who doesn't stand in the way of sinners}{and doesn't sit in the seat of scorners.}%%
    
    \pvab{\vn{1:2} Truly his joy\footnotemark\ is in the law of the \textsc{Lord}~---}{he ponders\footnotemark\ day and night on His law.}%%
    \fntalt{pleasure, delight}%%
    \fnthalot{c) \textbf{read in an undertone} \haref{Ps}{1}{2}; d) \textbf{ponder (by talking to oneself)} with accusative \haref{Is}{33}{18}, with \textit{b}$^\textit{e}$ \haref{Ps}{63}{7}; e) \textbf{plan} with accusative \haref{Ps}{2}{1}; f) \textbf{speak}, proclaim \haref{Is}{59}{3}}%%
    
    \pvab{\vn{1:3} He will be like a tree}{planted by streams of water\footnotemark}%%
    \fnted{So even in the heat it will grow.}%%
    
    \pvbb{which gives fruit in its season;}{its leaves will not wither.}%%
    
    \pvba{He will succeed in everything he does.}%%
    
    \pvba{\vn{1:4} The wicked aren't so!\footnotemark}%%
    \fntlit{Not thus the wicked!}%%
    
    \pvbb{Because they're like chaff}{that the wind blows away.}%%
    
    \pvab{\vn{1:5} Because of this, the wicked won't rise up in judgment,\footnotemark}{neither sinners in the assembly\footnotemark\ of the righteous.}%%
    \fnted{They won't be able to object during judgment. They will be invalid witnesses.}%%
    \fntca{\septuagint\ \Greek{ἐν βουλῇ} cf 1}{The Septuagint gives ``in the council of,'' compare verse~1}%%
    
    \pvab{\vn{1:6} For the \textsc{Lord} knows the way of the righteous;}{and the way of the wicked is lost.}%%
\end{inparaenum}
