\heading{110}{Christ will sit at the Father's right hand and will have the Melchizedek priesthood}%
\cf{Appendix~\ref{app:psalm-110}}%
\ed{``The most venerable tradition in Israel is to become both a king and a priest.'' ---~Steven D.\ Ricks, 30 September 2013}%

\begin{inparaenum}
  \noindent\verse{110:1} A psalm of David.\smallskip%%

  \pb A declaration%
  \alt{prophecy, word, utterance, oracle}
  of the \lord\ to my lord:%
  \cf{Appendix~\ref{app:adonai}}%
  \ed{Or any king, not necessarily David. Since this is a messianic psalm, it directed at Christ. That said, this is referring to a person to whom the psalmist, David, is subordinate to. It could very easily be Saul in the earlier days of David's life (cf.\ verse~3).}%
  \pa ``Sit at my right hand%%

  \pb until I make your enemies\pa a footstool for your feet.%%

  \pa \verse{110:2} Your mighty scepter\ed{symbolic of kingship}\pa the \lord\ will send from Zion\pa%
  to rule in the midst of your enemies.%
  \ed{Understood: that you may rule in the midst of your enemies.}%%

  \pa \verse{110:3} Your people are willing volunteers%
  \alt{free-will offering}%%
  \pa in the day of your strength;%
  \ie{When you go out to way, your people volunteer willingly.}%%

  \pb in royal robes%
  \alt{holy splendor/honor}%
  \ca{prb l c \fragheb\ mlt Mss \symmachus\ Hier \Hebrew{בהררי}}{probably read with the Cairo Genizah and multiple manuscripts from Symmachus' Greek translation of the Old Testament and Hieronymus give \Hebrew{הרר} [instead of \Hebrew{הדר}]}%
  \pa from the womb, from the morning light,%
  \halot{xxxx}{(reddish) \textbf{(light before) dawn}}%%
  \pa is the dew%
  \ca{\missing\ \septuagint; prp \Hebrew{כְּטַל}}{missing in the Septuagint; probably ``as the dew''}%
  \alt{light rain, strength, fortitude, might (in a non-sexual way)}
  of your youth.%%

  \pa \verse{110:4} The \lord\ has sworn\pa and will not retract:%%

  \pb You are forevermore a priest\pa after the Order of Melchizedek.%%

  \pb \verse{110:5} The Lord%
  \ca{\fragheb\ mlt Mss \Hebrew{יהוה}}{multiple manuscripts from the Cairo Genizah have the Tetragrammaton}
  at your right hand\pa shall smite\alt{beat} kings in the day of His anger.%%

  \pa \verse{110:6} He will judge among the nations;\pa He will fill%
  \ed{Understood (from the parallel in the first hemistich): the nations}%
  with corpses;\pa He will smite in pieces the the great ones\alt{heads}\pa over the great earth.%%

  \pa \verse{110:7} He will drink from the brook by the road,\pa therefore he will lift up the head.''%
  \ie{once He's been made great}%
  \ca{\fragheb\ \Hebrew{ראשׁי}, 2 Mss \peshitta\ \Hebrew{רֹאשׁוֺ}, 3 Mss + \Hebrew{הללויה} (2 Mss om in 111,1)}{The Cairo Genizah contains ``my head,'' two manuscripts of the Peshitta have ``his head,'' and three manuscripts include ``Hallelujah'' (two of which omit it in \vref{Ps}{111}{1})}%%
\end{inparaenum}
