\heading{23}{The Lord is my shepherd~--- He will comfort and provide}

\begin{inparaenum}
    {\noindent\verse{23:1} A psalm of David:}%%
    
    \pvcb{The \textsc{Lord} is my shepherd: I will not go lacking.}{\vn{23:2} He makes me to lie down in grassy pastures.\footnotemark}%%
    \fntca{huc tr \Hebrew{׃}}{hither transpose a \textit{sof pasuq} (symbol for end of verse)}%%
    
    \pvcb{He leads me beside calm waters.}{\vn{23:3} He rescues my soul.}%%
    
    \pvcb{He leads me in\footnotemark\ paths of righteousness}{for His name's sake.}%%
    \fntalt{on}%%
    
    \pvbc{\vn{23:4} Although I walk}{through the valley of darkness\footnotemark\footnotemark}{I won't fear evil}%%
    \fnted{Compare the Ugaritic \textit{\d{s}lmt} (sometimes with a \textit{waw} as \textit{\d{s}lmwt}) (to the Hebrew \Hebrew{צַלְמָוֶת}) meaning ``deepest darkness.''}%%
    \fnted{Classically understood as ``shadow (\Hebrew{צל}) of death (\Hebrew{מות})'' but a bit better as ``darkness.''}%%
    
    \pvcc{because You are with me.}{Your rod and your staff}{comfort me.}%%
    
    \pvbb{\vn{23:5} You prepare a table before me}{in the presence of my enemies.}%%
    
    \pvcb{You have anointed my head with oil;}{my cup overflows.\footnotemark}%%
    \fnted{probably with wine}%%
    
    \pvbb{\vn{23:6} Surely, goodness and loving-kindness\footnotemark\ will follow me}{all the days of my life.}%%
    \fntca{\septuagint\peshitta\ + suff 2 sg}{the Septuagint and the Peshitta have a second singular pronominal suffix [i.e., ``your loving-kindness'']}%%
    
    \pvcb{I will live in the house of the \textsc{Lord}}{for the rest of my days.}%%
\end{inparaenum}
