\heading{24}{Ancient equivalent of a temple recommend interview~--- only those who are ritualistically and ethically clean may enter the temple~--- the King of Glory is the Lord}

\begin{inparaenum}
  {\noindent\verse{24:1} A psalm of David:}%%
  
  \pvcb{The earth and its fullness are the \textsc{Lord}'s:}{the world and its inhabitants.}%%
  
  \pvbb{\vn{24:2} For\footnotemark\ it was He that founded it upon the sea~---}{He established it upon the rivers.}%%
  \fntca{\missing\ \septuagint\symmachus\theodotion}{missing in the Septuagint, Symmachus' and Theodotion's Greek translations of the Old Testament}%%
  
  \pvcb{\vn{24:3} Who shall ascend to the \textsc{Lord}'s mountain?}{Who shall stand in His place of holiness?}%%
  
  \pvbb{\vn{24:4} Those with clean palms and soles\footnotemark\ and a pure heart;}{who don't lift their souls in vanity}%%
  \fntie{ritualistically and ethically clean palms and soles. This is a higher order of cleanliness than the Mosaic Law which only requires ritualistic cleanliness.}%%
  
  \pvca{and don't swear\footnotemark\ deceitfully.}%%
  \fntie{in a covenantal sense}%%
  
  \pvbb{\vn{24:5} He shall bear a blessing from the \textsc{Lord}}{and righteousness from the God of his salvation.}%%
  
  \pvbb{\vn{24:6} This is the generation of those who seek Him~---}{that seek your face,\footnotemark\footnotemark\ O Jacob. Selah!}%%
  \fntca{\septuagint\ om suff, 2 Mss \septuagint\peshitta\ + \Hebrew{אֱלֹהֵי}, \targum\ suff 3 sg}{The Septuagint omits the [pronominal] suffix, two manuscripts and the Septuagint and Peshitta add ``of God,'' the Targum has a third person singular suffix [``his face'']}%%
  \fnted{It is shocking that any manuscript of the Septuagint would include ``the face of God'' because the Septuagint tends to exclude phrases referring to the anthropomorphic nature of God.}%%
  
  \pvac{\vn{24:7} O gates, lift up your heads!}{Be lifted up, O everlasting doors!\footnotemark}{The King of Glory shall come in!}%%
  \fnted{In the story of the battle of Baal and Yam has an interesting parallel to this story. On line~27 of column~iv it reads ``Lift up, gods, your heads,'' the same structure found here. This is of interest because it shows the same structure in a scriptural text of a pagan religion. See further in \textit{Canaanite Myths and Legends}, J.C.L. Gibson.}%%
  
  \pvac{\vn{24:8} Who is this, the King of Glory?}{The \textsc{Lord}: powerful and manly:\footnotemark}{the \textsc{Lord}, mighty in battle.}%%
  \fntalt{vigorous, heroic, mighty}%%
  
  \pvac{\vn{24:9} O gates, lift up your heads!}{Be lifted up, O everlasting doors!}{The King of Glory shall come in!}%%
  
  \pvac{\vn{24:10} Who is He, this King of Glory?}{The \textsc{Lord} of Hosts,}{He is the King of Glory. Selah!}%%
\end{inparaenum}
