\heading{41}{David discusses the virtues of a good person and how the Lord will uphold and bless such a person}

\begin{inparaenum}
  \noindent\verse{41:1} To the chief musician: a psalm of David.\smallskip%%
  
  \pa \verse{41:2} Oh the happiness of him who understands the poor~---\pa the \lord\ will deliver him in the day of evil.%%
  
  \pa \verse{41:3} The \lord\ shall preserve him~--- shall keep him alive. He shall be happy in the land.\pa You will not deliver his soul to his enemies' will.%%
  
  \pa \verse{41:4} The \lord\ will sustain him upon his sick bed.\lit{the bed of his sickness}\pa You've overturned all his bed in his sickness.%%
  
  \pb \verse{41:5} I~--- I~--- said, ``O \lord, be gracious to me.\pa Heal my soul because I've sinned against you.%%
  
  \pa \verse{41:6} My enemies speak evil against me:\pa `When will he die? When will his name perish?'%%
  
  \pa \verse{41:7} If someone comes to see me, he will speak falsely.\pa His heart gathers wickedness to himself. He goes outside and says it.%%
  
  \pa \verse{41:8} Everyone who hates me whispers together against me:\pa they devise evil against me.''%%
  
  \pa \verse{41:9} An ill%
  \halot{בְּלִיַּעַל}{\textbf{worthlessness}, \textbf{wickedness}\dots\ illness \haref{Ps}{41}{9}}%
  \ed{\Hebrew{בְּלִיַּעַל} from the Hebrew \Hebrew{יצל}: ``\textbf{help}, \textbf{be of use}''}
  word\alt{thing} is poured out on him%
  \ca{1 \Hebrew{בִּי} cf \septuagint}{1 has ``on me,'' compare the Septuagint}%
  \ed{This alternate reading is significant because it could change where the quotation marks go.}
  and he won't rise again from where he lay.%%
  
  \pa \verse{41:10} My ally in whom I trusted, who ate bread with me\pa has lifted his hand against me.%%
  
  \pb \verse{41:11} But You, \lord, be gracious to me and raise me up\pa that I may give recompense to them.%%
  
  \pa \verse{41:12} I know by this that You're pleased with me\pa because my enemy doesn't triumph\alt{shout} over me.%%
  
  \pa \verse{41:13} As for me, You've upheld me in my integrity\pa and established me before You forever.%%
  
  \pa \verse{41:14} Blessed be the \lord, the God of Israel\pa from eternity to eternity.%%
  
  \hfil Amen and amen.\hfil%%
\end{inparaenum}
