\heading{46}{God is our refuge and strength~--- His mighty power enumerated~--- be still and come to know God}

\begin{inparaenum}
  {\noindent\verse{46:1} To the music director.\cf{\vref{Ps}{45}{1}} For the sons of Korah. For the \Hebrew{עֲלָמוֺת}.\halot{xxxx}{unexplained term in performance} A song.}%%
  
  \pvbb{\verse{46:2} God is our refuge and strength,}{well proven as a help in distresses.}%%
  
  \pvbb{\verse{46:3} Therefore, we will not fear though the earth moves}{and the mountains shake in the midst\footnotemark\ of the seas.}%%
  \lit{heart}%%
  
  \pvbc{\verse{46:4} The waters shake and foam;}{the mountains shake as it swells.\footnotemark}{Selah!}%%
  \ca{ins \Hebrew{יעקב} \Hebrew{אלהי} \Hebrew{משגב־לנו} \Hebrew{עמנו} \Hebrew{צבאות} \Hebrew{יהוה} sec 8.12}{insert ``The \textsc{Lord} of Hosts is with us. The God of Jacob is our refuge'' according to verses~8 and~12 [like a chorus]}%%
  
  \pvcb{\verse{46:5} The river,\footnotemark\ its streams, make\footnotemark\ the city of God rejoice:}{the holy dwelling-places of the Most High.}%%
  \halot{xxxx}{river around city of God \haref{Ps}{46}{5}}%%
  \ed{given in piel which can be causative}%%
  
  \pvbb{\verse{46:6} God is in the midst of it: it will not be shaken.}{God will help at dawn.\footnotemark}%%
  \lit{at the dawn of the morning.}%%
  
  \pvbb{\verse{46:7} The nations raged and the kingdoms tottered.}{He speaks and the earth melts.}%%
  
  \pvbc{\verse{46:8} The \textsc{Lord} of Hosts is with us.}{The God of Jacob is our refuge.}{Selah!}%%
  
  \pvcb{\verse{46:9} Come and see the works of the \textsc{Lord}\footnotemark}{who has set desolations in the earth.}%%
  \ca{mlt Mss \septuagint\super{A\textit{L}}\peshitta\ \Hebrew{אֱלֹהִים} ut 66,5}{multiple manuscripts, the Septuagint (codex Alexandrinus and textus Graecus ex recensione Luciani), and the Peshitta have ``God,'' as in \vref{Ps}{66}{5}}%%
  
  \pvba{\verse{46:10} He makes war cease to the end of the earth.}%%
  
  \pvcb{He smashes the bows and cuts up the spears.}{He burns chariots\footnotemark\ with fire.}%%
  \ca{\septuagint\ \Greek{καὶ θυρεούς} = \Hebrew{וַעֲגִלוֺת} cf \targum}{the Septuagint has ``shields'' (equivalent to the Hebrew word for the same), compare the reading found in the Targum}%%
  
  \pvbb{\verse{46:11} Be still and know that I am God.}{I am exalted among the nations. I am exalted in the earth.}%%
  
  \pvbc{\verse{46:12} The \textsc{Lord} of Hosts is with us.}{The God of Jacob is our refuge.}{Selah!}%%
\end{inparaenum}
