\heading{47}{Praise God~--- He rules over all}

\begin{inparaenum}
    {\noindent\verse{47:1} To the music director.\footnote{cf~\vref{Ps}{45}{1}} A psalm for the sons of Korah.}%%
    
    \pvbb{\vn{47:2} All the people, clap your hands!}{Shout to God\footnotemark\ with a shout of joy!}%%
    \fntca{\missing\ pc Mss}{missing in a few manuscripts}%%
    
    \pvbb{\vn{47:3} Because the \textsc{Lord}, the Most High,\footnotemark\ is awe-inspiring,}{a great king over the whole earth.}%%
    \fntalt{is exalted}%%
    
    \pvbb{\vn{47:4} He subjugates the people beneath us}{and the people\footnotemark\ under out feet.}%%
    \fnthalot{1.\ \textbf{people} (ethnic community) (archaic word) \haref{Gn}{25}{23} \& often;---2.\ \textbf{people} (in general, French \textit{on}) $\dagger$ \haref{Pr}{11}{26}}%%
    
    \pvbc{\vn{47:5} He has chosen our\footnotemark\ heritage\footnotemark\ for us,}{the pride\footnotemark\footnotemark\ of Jacob whom He loves.}{Selah!}%%
    \fntca{\septuagint\peshitta\ suff 3 sg}{the Septuagint and Peshitta have a third singular pronominal suffix [i.e., ``his heritage'']}%%
    \fnthalot{(inalienable) \textbf{hereditary possession}, \textbf{heritage}, acquired by individual or family by conquest or inheritance, both property (i.e., land and buildings) and (movable) goods}%%
    \fntalt{loftiness, arrogance}%%
    \fnted{other translations give ``excellency''}%%
    
    \pvcb{\vn{47:6} God has gone up amidst shouting,}{the \textsc{Lord} amidst\footnotemark\ the sound of a shofar.}%%
    \fntlit{with}%%
    
    \pvbb{\vn{47:7} Praise\footnotemark\ God!\footnotemark\footnotemark\ Praise!\footnotemark}{Praise our King! Praise\footnotemark}%%
    \fntalt{sing, play an instrument [for all uses of ``praise'' in this verse]}%%
    \fnted{understood: in song}%%
    \fntca{pc Mss ´\Hebrew{לֵא}, Ms \septuagint\ \Hebrew{לֵאלֹהֵנוּ}}{a few manuscripts have the preposition \textit{to}, and a manuscript and the Septuagint have ``to our God''}%%
    \fntca{\peshitta\ \textit{b\v{s}wb\d{h}'} in splendore}{the Peshitta says ``in splendor'' [instead of ``Praise!'']}%%
    \fntca{\missing\ 2 Mss \peshitta}{missing in two manuscripts and the Peshitta}%%
    
    \pvbb{\vn{47:8} because God is the king\footnotemark\ of the whole earth.}{Praise with insight.\footnotemark\footnotemark}%%
    \fntca{mlt Mss + \Hebrew{עַל־}}{multiple manuscripts add ``on'' [making it ``God is king over the whole earth,'' ``of'' simply being understood by context]}%%
    \fnthalot{unclarified musical term in Psalms; suggesting: cultic-song; passage for learning; wisdom-song put to music; in contrast or opposition to \textit{ma\'sk\^il} `with insight,' verbal root \textit{\'skl} hiphil participle}%%
    \fnthalotu{Septuagint \Greek{συνέσεως} [prudence], \Greek{εἰς σύνεσιν} [in prudence], Vulgate \textit{intellectus} [understanding], Jerome \textit{eruditio} [learning], \Hebrew{טָבָא} \Hebrew{שׂכְלָא}\hspace*{0em}/\hspace*{0em}\Hebrew{סִ}; technical term but meaning unclear: ``cult song'' (Kittel), ``memory passage'' (Maag 193f), wisdom song performed to music\dots\ ---in contrast or opposition to \Hebrew{מַשְׂכִּיל} with insight, I \Hebrew{שׂכל} hiphil participle $\dagger$}%%
    
    \pvbb{\vn{47:9} God rules over the nations,\footnotemark}{God\footnotemark\ sits on his holy throne.}%%
    \fntca{2 Mss \septuagint\super{RAa1} + \Hebrew{כָּל־}}{two manuscripts and the Septuagint (codex Veronensis, codex Alexandrinus, xxxx) add ``whole/all/entire'' [i.e., ``over all the nations'']}%%
    \fntca{\missing\ Ms, frt dl}{missing in a manuscript, perhaps deleted [so simply ``He sits on\dots'']}%%
    
    \pvbb{\vn{47:10} the nobles\footnotemark\ people are gathered\footnotemark}{with\footnotemark\footnotemark\ the people\footnotemark\ of the God of Abraham.}%%
    \fntalt{noblemen, generous, willing, noble}%%
    \fnted{The niphal form of \Hebrew{אסף} means ``to gather'' when the verb is intransitive (as it is here). Were this verb transitive it could mean ``to conspire against'' or ``to unite against.''}%%
    \fnted{prepositions in poetry are generally understood through context}%%
    \fntca{prp \Hebrew{אָהֳלֵי} \Hebrew{עִם}}{proposed ``with the tent of''}%%
    \fntca{\septuagint(\peshitta) \Greek{μετά} = \Hebrew{עִם}; 1 prb \Hebrew{עַם} \Hebrew{עִם} (hpgr)}{the Septuagint and Peshitta have ``with'' [instead of ``people'' (\Hebrew{עַם})]; one probably has ``with the people'' (haplography)}%%
    
    \pvcb{Because the protection\footnotemark\ of the earth is God's.}{He is greatly exalted.}%%
    \fntalt{shield, refuge (metaphorically)}%%
\end{inparaenum}
