\heading{93}{The Lord is exalted~--- He is eternal~--- even the rivers praise Him}

\begin{inparaenum}
    \pvab{\vn{93:1} \footnotemark The \textsc{Lord} reigns. He has clothed Himself with loftiness/excellency(?).}{The \textsc{Lord} has clothed Himself, He's girded Himself with strength.}%%
    \fntca{\septuagint\ pr \Greek{εἰς τὴν ήμέραν τοῦ προσαββάτον, ὄτε κατῴκισται ἡ γῆ}$\cdot$ \Greek{αἾνος ᾠδῆς τῷ Δαυιδ}}{the Septuagint inserts the following before everything else: ``In the day of xxxx, xxxx xxxx Earth$\cdot$ xxxx xxxx xxxx David''}%%
    
    \pvbc{Also, the earth\footnotemark\ is established, it will not stagger.}{\vn{93:2} Your throne is established since then~---}{You are from eternity.}%%
    \fntalt{continents}%%
    
    \pvac{\vn{93:3} The rivers have been lifted up, O \textsc{Lord}.}{The rivers have lifted up their voices.}{The rivers have lifted up their waves.\footnotemark\footnotemark}%%
    \fntca{\missing\septuagint*}{[this last bit] missing in the Septuagint (textus Graecus originalis)}%%
    \fnthalot{\textbf{pounding} (of waves) \haref{Ps}{93}{3}. $\dagger$}%%
    
    \pvac{\vn{93:4} Than the sound of many waters,}{than the mighty surf,\footnotemark}{mighty in heaven\footnotemark\ is the \textsc{Lord}.}%%
    \fntlit{breakers of the sea}%%
    \fnthalot{\textbf{height} (i.e. place on high): --- 1.\ \textbf{elevation of ground} \haref{2K}{19}{23}, \textit{r\=o'\v s m\superit{e}r\^om\^im} \haref{Pr}{8}{2}; --- 2.\ highly placed location, \textbf{place on high} \haref{Is}{22}{16}; =~exalted \haref{Je}{17}{12}; --- 3.\ \textbf{upward} (adverb accusative) \haref{2K}{19}{22}; --- 4.\ \textbf{high social position} \haref{Jb}{5}{11}; --- 5.\ moral: \textit{mimm\=ar\^om} from a high position, loftily \haref{Ps}{73}{8}; --- 6.\ \textit{m\=ar\^om} =~heaven: a)~windows of heaven \haref{Is}{24}{18}, host of heaven \haref{Is}{24}{21}; b)~home of God (singular and plural) \haref{Is}{33}{5}; c)~to heaven \textit{lamm\=ar\^om} \haref{Is}{38}{14}, from heaven \textit{mimm\=ar\^om} \haref{2S}{22}{17}.}%%
    
    \pvac{\vn{93:5} The testimonies have been very reliable.\footnotemark}{Holiness is to your beautiful house,}{O \textsc{Lord}, forever.\footnotemark}%%
    \fntalt{stable}%%
    \fntlit{for length of days.}%%
\end{inparaenum}
