\heading{97}{Praise God~--- He is greater than the pagan gods}

\begin{inparaenum}
    \pvab{\vn{97:1} The \textsc{Lord} reigns. Let the earth be glad,}{let the many islands\footnotemark\footnotemark\ be joyful.\footnotemark}%%
    \fnthalot{the islands and coasts of the Mediterranean are for the Old Testament the extremes of the western world.}%%
    \fntalt{Phoenicians}%%
    \fntalt{rejoice.}%%
    
    \pvab{\vn{97:2} Clouds and darkness surround Him~---}{righteousness and judgment are the support of His throne.}%%
    
    \pvab{\vn{97:3} Fire goes before Him}{and devours\footnotemark\ round about His\footnotemark\ enemies.}%%
    \fntalt{scorches}%%
    \fntca{pc Mss \Hebrew{סְבִיבָיו}}{a few manuscripts have ``round about His'' [justifying the present (understood) rendering]}%%
    
    \pvab{\vn{97:4} His lightning has lit the world~---}{the earth saw and xxxx-root?.\footnotemark}%%
    \halot{xxxx}%%
    
    \pvab{\vn{97:5} The mountains melted like wax\footnotemark before the \textsc{Lord},\footnotemark}{from before\footnotemark\ the Lord of the whole earth.}%%
    \fnthalot{always metaphor of melting}%%
    \fntca{prb dl}{[``before the \textsc{Lord}''] should probably be deleted}%%
    \fntca{\missing\ \peshitta}{[``from before'' is] missing in the Peshitta}%%
    
    \pvab{\vn{97:6} The heavens announce His righteousness,}{all the people see His glory.}%%
    
    \pvac{\vn{97:7} All the idol servants are ashamed~---}{those who boast themselves of pagan gods:~---\footnotemark}{bow\footnotemark\ to Him, all you gods.\footnotemark\footnotemark}%%
    \fnthalot{always contemptuously as nonentities, idols}%%
    \fntca{\septuagint\peshitta\ imp}{the Septuagint and Peshitta have this in imperative form}%%
    \fntca{\septuagint(\peshitta) \Greek{οἱ ἄγγελοι αὐτοῦ}}{the Septuagint (and Peshitta) have ``angels''}%%
    \fnted{This is typical behavior for the Septuagint to avoid any reference to the anthropomorphic nature of God.}%%
    
    \pvac{\vn{97:8} Zion has heard and rejoiced, [xxxx: why not ``Hear and rejoice, O Zion!'']}{the daughters of Judah are joyful,}{because of Your judgments, O \textsc{Lord},}%%
    
    \pvab{\vn{97:9} because You, O \textsc{Lord},\footnotemark\ are Most High over all the earth;}{You have been greatly exalted over all [other(?)] gods.}%%
    \fntca{\missing\ Ms, dl m cs}{missing in a manuscript, should be deleted to preserve the meter case}%%
    
    \pvac{\vn{97:10} You who love the \textsc{Lord}, hate evil:}{He protects the souls of His devout,}{He delivers them from the power of the wicked.}%%
    
    \pvab{\vn{97:11} Light is sown\footnotemark\ for the righteous,}{and joy for the righteous\footnotemark\ of heart.}%%
    \fntca{1 c Ms Vrs \Hebrew{זָרַח} ut 112,4}{one with all or most translations has ``to shine'' [``has shown'' or ``is shining''] as in \vref{Ps}{112}{4}}%%
    \fntalt{just, upright}%%
    
    \pvab{\vn{97:12} Rejoice, O righteous, in the \textsc{Lord}!}{Praise at the memory\footnotemark\ of His holiness!}%%
    \fntlit{remembrance}%%
\end{inparaenum}
