\heading{26}{The people install Uzziah as king~--- Uzziah, like his father, was righteous and God made him prosper~--- he goes to war, becomes vain, and tempts God~--- he makes an illegal sacrifice and is cursed with a serious skin condition}

\begin{inparaenum}
    \verse{26:1} All the Judahites\lit{the people of Judah} took Uzziah, who was sixteen years old, and crowned\alt{made, installed} him king instead of his father, Amaziah.%%
    \verse{26:2} He built Eloth and restored it to Judah after the king lied with his fathers.\ie{after death}%%
    
    \verse{26:3} Uzziah was sixteen years old when he became king. He ruled in Jerusalem for fifty-two years. His mother's name was Jecholiah of Jerusalem.%%
    \verse{26:4} He did what was right in the eyes of \textsc{God} like everything that his father Amaziah had done.%%
    \verse{26:5} He worshiped\alt{sought, cared about, inquired regarding} God in the days of Zechariah (who understood\alt{perceived, paid attention to, considered, gave heed to, noticed} the visions of God). In the days he worshiped the \textsc{Lord}, God made him prosper.%%
    
    \verse{26:6} He went and fought\alt{do battle with, come to close quarters with} against the Philistines and made a breach in\alt{broke down} the wall of Gath, the wall of Jabneh, the wall of Ashdod. He built cities in Ashdod and Philistia.%%
    \verse{26:7} God came to his aid\alt{supported, helped} against the Philistines, the Arabians living in Gur-Baal, and the Maonites.%%
    \verse{26:8} The Ammonites gave a present to Uzziah and his name went to the entrance of Egypt because he became incredibly\lit{lifted way up, escalated. From \Hebrew{עלה} meaning to lift up.} strong.%%
    \verse{26:9} Uzziah built towers\halot{in a defensive wall} in Jerusalem by the corner gate, the valley gate, and at the corner, and he fortified\alt{strengthened} them.%%
    \verse{26:10} He built towers in the desert and hewed out\alt{quarried; other translations render this as ``dug,'' but \textsc{halot} only mentions stonework.} many wells because he had a lot of cattle, both in the lowland and in the plains;\alt{on the plateau, \Hebrew{מִישׁוֺר} meaning both plain and plateau.} he had\understood\ serfs\halot{peasant, not owning land, belonging to landlord} and wine-growers\alt{vinedresser} in the mountains and in Carmel because he loved growing stuff.%%
    
    \verse{26:11} Uzziah had an army\halot{alt., wealthy landowner, qualified, fit for military service; (large) landowner, obligated to military service and the furnishing of a certain number of men; then valiant man without regard to property; all the armed host of a people and a satrapy (a provincial governor in ancient Persia).} who was making war, going out by troop as commissioned\lit{listed by name} by the hand of Jeiel the scribe and Maaseiah the record-keeper and by the hand of Hananiah, the second after the king.%%
    \verse{26:12} The whole number of the head honchos\lit{chief fathers} of the elite troops was 2\thinspace600.%%
    \verse{26:13} With regard to the strength\alt{power; lit., hand} of the army,\alt{host} 307\thinspace500~warriors who made war with great ability\alt{power, capacity, strength, means} to support\alt{help, come to the aid of} the king against the enemies.%%
    \verse{26:14} For the entire army, Uzziah prepared shields, spears, helmets, scale-armor,\alt{coat of mail} bows, and slinging-stones for them.%%
    \verse{26:15} In Jerusalem, he made war machines\halot{(skillfully contrived) \textbf{war-machines}, spec. catapults}\ed{Not a ballista. The earliest ballistas were invented circa 400~\textsc{bc}.}~--- thought out\alt{invented} by technicians~--- to be placed\understood\ on the towers and the corners to shoot arrows and boulders.\lit{great stones} His name spread some distance\lit{out a distance} because he was extraordinarily helped  until he became strong.%%
    \verse{26:16} However, when he become strong, his heart was lifted up unto destruction. And he transgressed against the \textsc{Lord} his God by going unto the temple of the \textsc{Lord} and offering incense upon the altar of incense.%%
    \verse{26:17} And Azariah the priest went in after him with the priests of the \textsc{Lord}: eighty valiant men.%%
    \verse{26:18} And they withstood Uzziah the king and said unto him, ``This is not for thee, Uzziah, to burn incense unto the \textsc{Lord}; rather for the priests~--- the sons of Aaron~--- that are set apart\alt{consecrated, sanctified} to burn incense. Leave this holy place\lit{Go out from this sanctuary} for you have transgressed. And neither shall this be for thine honor from the \textsc{Lord} God.''%%
    \verse{26:19} Uzziah was wroth (and in his hand he had a censer of incense). And while he was angry at the priests, leprosy appeared\lit{rose up} in his forehead in the presence of the priests in the House of the \textsc{Lord} next to the altar of incense.%%
    \verse{26:20} Azariah, the chief priest, and all the priests looked upon him, and lo! he was leprous in his forehead. So they thrust him out from thence, yea, even he\ie{Uzziah} hurried out because the \textsc{Lord} has smitten him.%%
    \verse{26:21} Uzziah the king was leper unto his dying day~--- he dwelt in a separate house (being a leper) because he was cut off from the House of the \textsc{Lord}. And Jotham, his son, took over\lit{was over} the king's affairs,\lit{house} judging the people of the land.%%
    \verse{26:22} The rest\alt{remainder} of the history of Uzziah, the first and the last, was written by the prophet Isaiah, the son of Amoz.%%
    \verse{26:23} Uzziah lay with his fathers,\ie{in death} and they buried him with his fathers in the kings'\lit{that the kings have} cemetery,\lit{field of the graves} for the said, ``He had a skin condition.'' His son, Jotham, reigned in his stead.%%
\end{inparaenum}
