\heading{6}{xxxx}

\begin{inparaenum}
    \verse{6:1} And when Sanballat, Tobiah, Geshem the Arabian, and the rest of our enemies had heard\footnote{When it was heard by} that I had built the wall and that there was no breach in it (also, up until that time\footnote{until then} I had not set up the gates)%%
    \verse{6:2} that Sanballat and Geshem sent unto me, saying, ``Come, and we shall meet together in the villages of\footnote{in} the valley of Ono,'' but they were thinking of doing me wrong.\footnote{thinking to do evil unto me.}%%
    \verse{6:3} So I sent messengers unto them, saying, ``I am doing a great work so I cannot come down. Why should the work cease\footnote{\textit{sabbath} is used here. That is the concept being conveyed.} while I leave it and come down to you?''%%
    \verse{6:4} So they sent unto me~--- as I've said\footnote{according to this word}~--- four times, and I told them likewise.%%
    \verse{6:5} So Sanballat sent his servant unto me~--- as I've said~--- the fifth time with an open letter in his hand,%%
    \verse{6:6} and it was written therein: ``It hath been heard among the nations, and Gashmu hath said, `You and the Jews are thinking of rebelling. Therefore, you are building the wall and hast been a king unto them. Even according to these words!%%
    \verse{6:7} Moreover, thou hast appointed prophets who\footnote{to} call for thee in Jerusalem, saying, ``A king is in Judah!''' Now, it hath been heard by the king, even according to these words. Come now and let us reason together.''%%
    \verse{6:8} So I sent unto him, saying, ``No, it hath not been as thou hast said\footnote{according to these words that thou hast said} because you've made this up\footnote{devised} from thine own heart.%%
    \verse{6:9} For all of them are making us afraid, saying, `Their hands are too weak\footnote{feeble} for the work~--- that's why it's not done.' O God, strengthen my hands!''\footnote{Some slight eisegesis to get ``O God'' into this clause, but other translations seem to agree with me. Logic: why would he be asking his enemies to strengthen his hands?}%%
    \verse{6:10} Then I entered the house of Shemaiah, son of Delaiah (son of Mehetabeel), who was shut up,\footnote{locked in, under house arrest, not going outside} and he said,%

    \pvcb{``Let us meet at the House of God,}{inside the temple.}%

    \pvcb{So we shall shut the doors of the temple}{because they are coming to slay thee.}%

    \pvca{Even by night are they coming to slay thee.''}\\*%%
    \verse{6:11} And I said, ``Would a man like me flee?\footnote{A man such as I (or, like me)~--- would he flee?} And who, like me, would go in unto the temple to\footnote{and} live? I will not go.''\footnote{go in}%%
    \verse{6:12} And I perceived that lo! God had not sent him because he had prophesied this word against me: Tobiah and Sanballat had hired him.\footnote{That's some serious transgression of taking the name of the Lord in vain.}%%
    \verse{6:13} He was hired to frighten me.\footnote{Therefore was he hired, that I might be afraid} That I might do so: sin~--- that I might have something to go off of to spread an evil report so that they could reproach me.%%
    \verse{6:14} My God, remember Tobiah and Sanballat according to these, their works. And also the prophetess Noadiah and the rest of the prophets who would have made me afraid.%%
    \verse{6:15} So the wall was finished in fifty-two days: on the twenty-fifth of Elul.%%
    \verse{6:16} And when all of our enemies had heard it, and when all the nations round about us were afraid (and much cast down in their own perception) for they knew that this work was brought to pass by our God.%%
    \verse{6:17} Additionally, in those days, the noblemen of Judah sent a lot of letters to Tobiah, so those who are with Tobiah came to them.%%
    \verse{6:18} For there are many in Judah sworn to him: \footnote{for}he hath a son-in-law of Shechaniah, son of Arah; his son Johanan hath married the daughter of Meshullam, son of Berechiah.%%
    \verse{6:19} They also spoke before me of his goodness and then reported my words to him. So Tobiah sent letters to make me afraid.%%
\end{inparaenum}
