\heading{1}{The Lord foreordains Jeremiah to be a prophet~--- He comforts him and blesses him to fulfill his calling~--- xxxx}

\begin{inparaenum}
  \verse{1:1} The words of Jeremiah, son of Hilkiah, of the priests who were in Anathoth in the land of Benjamin;%%
  \verse{1:2} to whom the word of the \textsc{Lord} came in the days of Josiah, son of Amon, king of Judah, in the thirteenth year of his reign.%%
  \verse{1:3} xxxx\smallskip%%
  
  \pa \verse{1:4} The word of the \textsc{Lord} came to me, saying,%%
  
  \pa \verse{1:5} ``Before I formed%
  \ed{The verb used here, \Hebrew{יָצַר}, is defined in \textsc{halot} as generally meaning ``\textbf{form}, \textbf{shape}.'' However, when the subject is God, it means ``\textbf{create}, \textbf{form}'' but is noted as being an ``older, concrete synonym of \textit{b\=ar\=a'}''; since Jeremiah comes so late in the Hebrew period, it is doubtful, from a language standpoint, that the Lord is talking about creating. Additionally, from a theological and biological standpoint, it would be inconsistent to say that the Lord created Jeremiah in his mother's womb. This could only work poetically, not literally.}
  you in the womb, I knew you.\pa Before you came out of the womb, I sanctified you.%% xxxx: define the three types of ``knowing''
  
  \pa I made you a prophet to the nations.''%
  \ca{\septuagint\superit{C}\super{a1} sg}{the Septuagint (textus Graecus in genere Catenarum traditus a1) has this in singular [i.e., ``to the nation'']}
  \smallskip%%
  
  \noindent\verse{1:6} I said, ``Ah, \textsc{Lord} God! I can't speak because I'm a child.''%%
  
  \verse{1:7} The \textsc{Lord} said to me,\smallskip%%
  
  \pb ``Don't say, `I'm a child,'%
  \ca{pc Mss \septuagint\peshitta\ + \Hebrew{כִּי}}{a few manuscripts of the Septuagint and Peshitta add ``because'' [e.g., ``Don't say [it's] because I'm a child'']}%%
  
  \pb because you will go to everyone I will send you to,\pa and because\ed{repeated} you'll say everything I command you.%%
  
  \pa \verse{1:8} Don't be afraid of them\lit{of their faces}\pa because I am with you to deliver you,''%%
  
  \pa declares the \textsc{Lord}.\pa \verse{1:9} The \textsc{Lord} reached out%
  \lit{stretched out His hand}
  \ca{\septuagint\ + \Greek{πρός με}}{the Septuagint adds ``to me'' [i.e., The \textsc{Lord} stretched out His hand to me'']}
  and touched my mouth. And the \textsc{Lord} said to me,%%
  
  \pb ``I've put\alt{given} my words in your mouth.\pa \verse{1:10} Look, I've entrusted you this day%%
  
  \pa over the nations and kingdoms to tear up\alt{uproot, pull up, drive out} and tear down,\alt{break up, demolish, destroy} to destroy and to demolish, to build and to plant.''\smallskip%%
  
  \verse{1:11} %%
  \verse{1:12} %%
  
  \verse{1:13} %%
  \verse{1:14} xxxx\smallskip%%
  
  \pb xxxx\pa xxxx%%
  
  \pa \verse{1:15} xxxx\pa xxxx\pa xxxx%%
  
  \pb xxxx\pa xxxx%%
  
  \pb xxxx\pa xxxx%%
  
  \pa \verse{1:16} xxxx\pa xxxx%%
  
  \pb xxxx\pa xxxx%%
  
  \pb \verse{1:17} xxxx\pa xxxx%%
  
  \pa xxxx\pa xxxx%%
  
  \pa \verse{1:18} xxxx\pa xxxx%%
  
  \pb xxxx\pa xxxx%%
  
  \pb xxxx\pa xxxx%%
  
  \pa \verse{1:19} xxxx\pa xxxx%%
\end{inparaenum}
