\heading{33}{xxxx}

\begin{inparaenum}
  \verse{33:1} %%
  \verse{33:2} %%
  \verse{33:3} %%
  
  \verse{33:4} Therefore, thus says the \lord, the God of Israel, regarding the houses of this city of the kings of Judah which are broken down because of the siege ramps\halot{xxxx}{xxxx} and the sword:%%
  \verse{33:5} ``They are coming to way with the Chaldeans to fill them with the dead bodies I've slain in Mine anger and fury. Because of all this wickedness, I have hidden my face from this city.%%
  \verse{33:6} Look, I will xxxx xxxx and xxxx. I will heal them and reveal to them an abundance of peace and truth.%%
  \verse{33:7} I will free the captives of Judah and Israel. I will build them up as they were at first.%%
  \verse{33:8} I will cleanse them from all their iniquities which they have sinned against me. I will also pardon all their iniquities which they have sinned and transgressed against me.%%
  \verse{33:9} It shall be to me a joyous name of praise and beauty to all the nations of the earth who shall hear\ed{Does this apply to only the nations who hear of this goodness, or does it apply to all people because everyone will hear of this goodness?} of the good I do to them. However, they shall also fear and tremble because of all the good and peace that I do to you.''%%
  
  \verse{33:10} Thus says the \lord, ``It is again heard in this place of which you say, `It is waste: devoid of man and beast;' in the cities of Judah and in the desolated streets of Jerusalem which are devoid of man, inhabitant, and beast:%%
  \verse{33:11} An exultant voice, even a joyous voice, the voice of the bride and groom, saying, `Thanks to the \lord\ of Hosts, for the \lord\ is good. For His everlasting kindness to those who bring thanks to the house of the \lord. For I will turn back the land's captivity and make it as it used to be as at first,'' says the \lord.%%
  
  \verse{33:12} %%
  \verse{33:13} %%
  
  \verse{33:14} ``Listen. The days are coming,'' confirms the \lord, ``when I shall perform xxxx the good deed that I said to Israel and Judah's family.%%
  \verse{33:15} In those days, at that time, I will cause a righteous branch to grow up from David who shall produce judgment and righteousness in the world.%%
  \verse{33:16} In those days, Judah shall be saved and shall dwell in safety. It shall be called \textit{The \lord\ is our righteousness}.''%%
  
  \verse{33:17} For thus says the \lord: ``It shall never be taken away from David that someone sit upon the throne of the family of Israel.%%
  \verse{33:18} Neither shall it be taken from the priests and Levites to offer up burnt offerings, to make gifts, nor to do daily sacrifices.''%%
  
  \verse{33:19} The \lord's words came to Jeremiah, saying, ``xxxx''%%
  \verse{33:20} ``Thus says the \lord: `If you break my covenant during the day or during the nighttime so that xxxx\halot{xxxx}{xxxx \Hebrew{לְבִלְתִּי}} there's not day and night when there's supposed to be.\lit{in this season.}%%
  \verse{33:21} Additionally, my covenant will be broken with my servant David such that one of his sons shall not sit on his throne. xxxx%%
  \verse{33:22} I will multiply the progeny of my servant David and my Ministers the Levites as unnumbered as the hosts of Heaven and as unmeasurable as the sand of the sea.'\thinspace''%%
  
  \verse{33:23} The word of the \lord\ to Jeremiah, saying,%%
  \verse{33:24} ``Have you not considered what this people have spoken? They say, `The \lord\ has rejected the two families that He has chosen.' They despise my people and shall no longer be considered a nation before them.''%%
  
  \verse{33:25} Thus says the \lord: ``If my morning and evening covenant\ie{a covenant that is effectual both day and night~--- an eternal covenant.} is not, then the statues of Heaven and Earth shall not stand.%%
  \verse{33:26} Furthermore, I will reject\alt{cast away, refuse} the descendants of Jacob and my servant David. I will not\ed{xxxx is this supposed to be negated?} take his children to be rulers over the children of Abraham, Isaac, and Jacob because I will return them from captivity and have mercy on them.''%%
\end{inparaenum}
