\section{17}\label{1 Kings 17}
\heading{Elijah prophesied no rain~--- Elijah stays with the widow of Zarephath~--- the widow's son dies~--- the Lord heals the lad}
\begin{enumerate}[align=center]
    \verse{17:1} %%
    \verse{17:2} %%
    \verse{17:3} %%
    \verse{17:4} %%
    \verse{17:5} %%
    \verse{17:6} %%
    \verse{17:7} %%
    \verse{17:8} And the word of the \textsc{Lord} came unto him, saying,%%
    \verse{17:9} ``Arise. Go unto Zarephath~--- that is, unto Zidon~--- and dwell there. Behold! I have there commanded a widow woman to sustain thee.''%%
    \verse{17:10} So he arose and went unto Zarephath, came unto the city's entrance, and lo! there was a widow woman gathering sticks. So he called unto her and said, ``Please bring me a little water in a vessel so I can drink.''%%
    \verse{17:11} So she went to bring it and he calleth after her, and said, ``Please bring me some bread in thy hand.''%%
    \verse{17:12} Then she said, ``As the \textsc{Lord} thy God liveth, I don't even have a cake, but only a handful of meal in a pitcher and a little oil in a dish. Behold, I am gathering two sticks so that I can go in, prepare it for myself and my son, so that we can eat it and then die.''%%
    \verse{17:13} Elijah said unto her, "Fear not: go and do as I've said, only make for me from thence a little cake first and bring it out to me. Then make for thee and thy son last.%%
    \verse{17:14} ``For thus saith the \textsc{Lord} God of Israel, `The pitcher of meal shall not be consumed and the oil dish shall not be lacking until the day the \textsc{Lord} makes it rain on the face of the land.' ''%%
    \verse{17:15} So she went and did as Elijah had said. And she ate~--- she, he, and her household~--- for days.%%
    \verse{17:16} And the pitcher of meal was not consumed, neither did the oil dish lack, even according to the word of the \textsc{Lord} that He had spoken by the hand of Elijah.%%
    \verse{17:17} After these things, the son of the woman (the mistress of the house) became sick. And his sickness was so severe that there was no breath left in him.%%
    \verse{17:18} And she said unto Elijah, ``What's this to me and thee, O man of God? You came to me to make me remember my iniquity and to kill my son!''%%
    \verse{17:19} So he said to her, ``Give me your son.'' And he took him from her bosom, and took him to the upper chamber where he was abiding, and laid him on the bed.%%
    \verse{17:20} Then he cried unto the \textsc{Lord} and said, ``O \textsc{Lord} my God, have you really brought evil upon the widow with whom I sojourn by killing her son?''%%
    \verse{17:21} He stretched himself upon the child thrice and cried to the \textsc{Lord} and said, ``O \textsc{Lord} my God, please let this child's soul return to him!''%%
    \verse{17:22} And the \textsc{Lord} heard the voice of Elijah and the child's soul returned and he lived.%%
    \verse{17:23} And Elijah took the child and brought him down from the upper room into the house and gave him to his mother. And Elijah said, ``Look, your son lives.''%%
    \verse{17:24} And the woman said to Elijah, ``By this I now know that thou art a man of God and that the word of the \textsc{Lord} in thy mouth is true.''%%
\end{enumerate}
