\section{5}\label{Ezra 5}
\heading{Haggai and Zechariah prophesy~--- Zerubbabel and others begin construction of the temple~--- Tatnai, Shethar-boznai, and others write a letter to King Darius asking for an order to stop the Jews}
\begin{enumerate}[align=center]
    \verse{5:1} Haggai the prophet and Zechariah the son of Iddo prophesied in the name of the God of Israel concerning the Jews in Judah and in Jerusalem.%%
    \verse{5:2} Then Zerubbabel, son of Shealtiel, and Jeshua, son of Jozadak, rose and began to rebuild the house of God in Jerusalem. And the prophets supported them.%%
    \verse{5:3} At that time, Tatnai, the governor on the other side of the river, came to them with Shethar-boznai and their companions; and they said to them, ``Who gave you orders to build this house\footnote{temple} and to finish this wall?''%%
    \verse{5:4} They then said thus unto them, ``What are the names of the men who are building this house?''%%
    \verse{5:5} The eye of their God had been upon the elders\footnote{hoar-headed} of the Jews and they had not caused them to stop until the matter went to Darius. They then sent back a letter concerning this.%%
    \verse{5:6} A copy of the letter that Tatnai the governor on the other side of the river, Shethar-boznai, and his companions the Apharsachites\footnote{title for an official} who are on the other side of the river sent unto Darius the king.%%
    \verse{5:7} In the letter they sent to him was written: ``Peace be unto Darius the king.%%
    \verse{5:8} Be it known to the king that we have gone to the province of Judah, to the great temple of God, and it is build with square stones, wood is placed in the walls, and this work is done speedily. This work prospers in their hand.%%
    \verse{5:9} Then we asked these elders~--- thus did we ask them: `Who hath made a decree for you for this house to be built and this wall to be completed?'%%
    \verse{5:10} Additionally, we asked them for their names, to make it known unto you, so we can write the names of their leaders.\footnote{the men who are at the head}%%
    \verse{5:11} Thus have they returned the word, saying, `We are the servants of the God of heaven and earth. We have been building the house for many years'\footnote{from before this year} and that a great king of Israel had built and finished it.%%
    \verse{5:12} But after that, they made the God of heaven angry and He delivered them into the hands of the Chaldean Nebuchadnezzar, king of Babylon, who demolished this house and removed the other people to Babylon.%%
    \verse{5:13} But in the first year of Cyrus, king of Babylon, Cyrus, the king, gave an order to build this house of God.%%
    \verse{5:14} Furthermore, the gold and silver vessels from the house of God that Nebuchadnezzar removed from the temple in Jerusalem and brought to the temple in Babylon, Cyrus, the king, removed them from the temple in Babylon and brought them to Sheshbazzar whom he had appointed as governor.%%
    \verse{5:15} He\footnote{Cyrus} said to him, ``Lift up these vessels and go and put them in the temple in Jerusalem. Let the house of God be build on its place.''%%
    \verse{5:16} Then did Sheshbazzar come and lay the foundations of the house of God in Jerusalem. From then until now it has been being built and is not finished.%%
    \verse{5:17} And now, if it be good to the king, let an investigation\footnote{search, inquiry} be made into the treasury of the house of the king in Babylon whether it be that Cyrus the king made a decree to build this house of God in Jerusalem. Let the king's will concerning this be sent to us.''%%
\end{enumerate}
