\heading{6}{Darius starts an investigation to find if sanctions were ever given to the Jews to build a temple~--- evidence is found that Cyrus gave sanctions through a decree~--- Darius states that this decree is still in effect~--- he commands that the Jews be provided with whatever they need to build the temple~--- the temple is completed and dedicated}

\begin{inparaenum}
  \verse{6:1} King Darius then made a decree and they investigated in the records house where the treasures of Babylon were laid up.%%
  \verse{6:2} A scroll was found in Achmetha, a fortress in the province of Media, in the midst of which was written a record:%%
\end{inparaenum}

\begin{quotation}
  \begin{inparaenum}\setcounter{enumi}{2}
    \verse{6:3} ``In the first year of Cyrus, the king, King Cyrus made an order regarding the house of God in Jerusalem: let the house be built in the place where they are sacrificing sacrifices. Let its foundations be strongly laid: its height sixty cubits and its breadth sixty cubits.%%
    \verse{6:4} Three layers of square stones and a layer of new wood. Let the expense be charged\lit{given} to the king's house.%%
    \verse{6:5} Additionally, the gold and silver of the house of God which Nebuchadnezzar removed from the Jerusalem temple and brought to Babylon, let it be returned to the Jerusalem temple: you shall put it in the house of God.%%
    
    \verse{6:6} And now, Tatnai, governor of the trans-Euphrates,\lit{on the opposite side of the river} Shethar-boznai, and your companions the Apharsachites\ed{officials in Assyria} of the trans-Euphrates, stay away from this.\lit{be ye far from thence}%%
    \verse{6:7} Leave behind the work of the house of God; let the governor of the Jews and the elders of Judah build this house of God in its place.%%
    \verse{6:8} From me I give an order regarding what you should do to the work of those elders of Judah and the building of this house of God: the riches of the king which are on the other side of the river, the expenses be speedily given to these men that their work cease not.%%
    \verse{6:9} What they need, both young bullocks, rams, and lambs, for burnt offerings to the God of heaven; wheat, salt, wine, and oil, according to the saying of the Levite priest who is in Jerusalem, let it be given to them daily without negligence%%
    \verse{6:10} so that they can offer sweet odors to the God of heaven and pray for the life of the king and his sons.%%
    \verse{6:11} From me I give a command to anyone\lit{all mankind} who changes this decree: let wood be pulled out from their house and let them be impaled\lit{in impaling let them be impaled} and their house shall be made a dunghill for this.%%
    \verse{6:12} God caused His name to dwell there. He overthrew all the kings and people who try and cause His hand to change and hurt this house of God in Jerusalem. I, Darius, have made a decree~--- let it be done with all diligence.''\lit{let it be speedily done}%%
  \end{inparaenum}
\end{quotation}

\begin{inparaenum}\setcounter{enumi}{12}
  \verse{6:13} Then Tatnai, the governor of the trans-Euphrates, Shethar-boznai, and their companions speedily did as King Darius had sent.%%
  \verse{6:14} The elders of the Jews were continuing to build and prosper because of the prophecies of Haggai the prophet and Zechariah the son of Iddo. Since the decree of the God of Israel, Cyrus, Darius, and Xerxes\lit{Artaxerxes (\Hebrew{ְאַרְתַּחְשַׁ֖שְׂתְּא}), but historically this is more probably Xerxes.} king of Persia, they built and finished.\ed{Understood: the house of God in Jerusalem}%%
  \verse{6:15} This house was completed on the third day of the month Adar which was in the sixth year of the reign of King Darius.%%
  
  \verse{6:16} The sons of Israel, the priests, Levites, and the rest of the sons of the captivity made a joyous dedication to this house of God.%%
  \verse{6:17} They brought one hundred bullocks, two hundred rams, four hundred lambs, and twelve young goats (for a sin offering for all Israel according to the number of tribes of Israel) near to this temple of God for its dedication.%%
  \verse{6:18} They put their priests in their divisions, the Levites in their courses over the work of God that is in Jerusalem according to the writing in the scroll of Moses.%%
  
  \verse{6:19} The exiles\lit{children of the exiles} held\lit{made} Passover on the fourteenth of the first month%%
  \verse{6:20} because the priests and the Levites together\lit{as one man} were ritualistically clean. They were all clean. They slaughtered the Paschal lamb for all of the exiles, for all of their brethren the priests, for themselves.%%
  \verse{6:21} The children of Israel~--- the ones returning from exile, everyone who had separated themselves from the\ed{The definite article is used to clearly show that the state of uncleanliness was of the foreign nations, not of the returning Israelites.} state of ritualistic uncleanliness of the nations of the land~--- to worship the \textsc{Lord}, the God of Israel, ate.\alt{offered sacrifice; this is justified contextually because in verse~20 they slew the Paschal lamb, but since the lamb was already slain it makes more sense for them to eat than to again offer sacrifice.}%%
  \verse{6:22} They held with joy a festival of unleavened bread for seven days, for the \textsc{Lord} had made them happy.\alt{rejoice, joyful, glad} The \textsc{Lord}\understood\ made the Assyrian king change to their side,\lit{turn} to strengthen their hands in the work of the House of God, the God of Israel.%%
\end{inparaenum}
