\heading{38}{xxxx}

\begin{inparaenum}
  \pa \verse{38:1} The \textsc{Lord} answered Job from the gale\halot{xxxx}{heavy whirlwind} and said,%%
  
  \pb \verse{38:2} ``Who is this who darkens counsel\pa with words%
  \halot{xxxx}{\textbf{word} (utterance rather than lexical unit)}
  without knowledge?%%
  
  \pb \verse{38:3} Gird up your loins%
  \halot{xxxx}{\textbf{loins} (body between ribs and hipbones)}
  like a youth%
  \ca{Ms \Hebrew{כְגִיבּוֺר}}{a Hebrew manuscript has ``like a warrior''}
  \pa and I will inquire of\alt{ask, consult, demand, request} you and cause you to know.%%
  
  \pb \verse{38:4} Where were you when I laid the foundation of the earth?\pa Tell it, if you know.\lit{have a knowledge of understanding.}%%
  
  \pb \verse{38:5} Who placed its measurements, if you know?\pa Or, who's stretched it out on its measuring line?%%
  
  \pb \verse{38:6} On what have its pedestals\alt{sockets} been sunk?\pa Or, who has laid its cornerstone?%%
  
  \pb \verse{38:7} When the morning stars sang\alt{shouted} together for joy together\pa and all the children of God shouted for joy,%%
  
  \pb \verse{38:8} and He shut up of the sea with doors%
  \ed{not in construct, so it's ``sea with doors''}
  \pa when bursting forth, it went out of the womb.\ie{of the sea.}%%
  
  \pb \verse{38:9} When I made clothing for a cloud\pa and thick darkness its swaddling;%%
  
  \pb \verse{38:10} when I established\alt{set} my statute upon it\pa and place bars and doors,\ed{understood: on/over it}%%
  
  \pb \verse{38:11} and said, `You shall come to here, but no farther.\pa And here is placed on the haughtiness of your proud waves.'\lit{in the pride/haughtiness of your waves.'}%%
  
  \pb \verse{38:12} Have you, from your days, commanded the morning?\pa Have you made the light before dawn to know its place?%%
  
  \pb \verse{38:13} xxxx\pa xxxx%%
  
  \pb \verse{38:14} xxxx\pa xxxx%%
  
  \pb \verse{38:15} xxxx\pa xxxx%%
  
  \pb \verse{38:16} xxxx\pa xxxx%%
  
  \pb \verse{38:17} xxxx\pa xxxx%%
  
  \pb \verse{38:18} xxxx\pa xxxx%%
  
  \pb \verse{38:19} xxxx\pa xxxx%%
  
  \pb \verse{38:20} xxxx\pa xxxx%%
  
  \pb \verse{38:21} xxxx\pa xxxx%%
  
  \pb \verse{38:22} xxxx\pa xxxx%%
  
  \pb \verse{38:23} xxxx\pa xxxx%%
  
  \pb \verse{38:24} xxxx\pa xxxx%%
  
  \pb \verse{38:25} xxxx\pa xxxx%%
  
  \pb \verse{38:26} xxxx\pa xxxx%%
  
  \pb \verse{38:27} xxxx\pa xxxx%%
  
  \pb \verse{38:28} xxxx\pa xxxx%%
  
  \pb \verse{38:29} xxxx\pa xxxx%%
  
  \pb \verse{38:30} xxxx\pa xxxx%%
  
  \pb \verse{38:31} xxxx\pa xxxx%%
  
  \pb \verse{38:32} xxxx\pa xxxx%%
  
  \pb \verse{38:33} xxxx\pa xxxx%%
  
  \pb \verse{38:34} xxxx\pa xxxx%%
  
  \pb \verse{38:35} xxxx\pa xxxx%%
  
  \pb \verse{38:36} xxxx\pa xxxx%%
  
  \pb \verse{38:37} xxxx\pa xxxx%%
  
  \pb \verse{38:38} xxxx\pa xxxx%%
  
  \pb \verse{38:39} xxxx\pa xxxx%%
  
  \pb \verse{38:40} xxxx\pa xxxx%%
  
  \pb \verse{38:41} xxxx\pa xxxx%%
\end{inparaenum}
