\heading{42}{Job answers the Lord's questions from chapter~38~--- xxxx}

\begin{inparaenum}
  \verse{42:1} Job answered the \textsc{Lord}, and said,\smallskip%%
  
  \pb \verse{42:2} ``I know that You can do anything,\alt{all things}\pa no plan\alt{intention, concept, device} is impossible for you.%%
  
  \pb \verse{42:3} Who's this who hides counsel without knowledge?\pa Therefore, I have declared and not understood; things too wonderful to me that I don't know.%%
  
  \pb \verse{42:4} Please hear me and I will speak~---\pa I will ask You and You will let me know.%%
  
  \pb \verse{42:5} By the hearing of the ear I'd heard of You,\pa but now my eyes have seen You.%%
  
  \pb \verse{42:6} Therefore, I blame myself and have repented\pa in dust and ashes.''\smallskip%%
  
  \verse{42:7} After the \textsc{Lord} had spoken these words to Job,\ed{Isn't this the wrong direction?} the \textsc{Lord} said to Eliphaz the Temanite: ``My anger burns against you and your two friends because you haven't spoken well of Me like My servant Job.%%
  \verse{42:8} So now, take seven bulls and seven rams and go to My servant Job and make a burnt offering\ed{for purification or guilt} for yourselves; My servant Job will pray for you because I will definitely accept him\lit{his face:} without treating you disdainfully because you haven't spoken well\lit{that which is right} of Me like My servant Job.''%%
  \verse{42:9} So Eliphaz the Temanite, Bildad the Shuhite, Zophar the Naamathite, went and did as the \textsc{Lord} had told them, and the \textsc{Lord} accepted Job.%%
  \verse{42:10} The \textsc{Lord} turned Job's fortune when he prayed for his friends, and the \textsc{Lord} added to Job twice as much as he'd before had.%%
  \verse{42:11} All his brothers, sisters, %%
  
  \verse{42:12} %%
  \verse{42:13} %%
  \verse{42:14} %%
  \verse{42:15} There was not found in all the land a woman as hot as Job's daughters; and their father gave them an inheritance\ed{probably an equal inheritance} among their brothers.%%
  
  \verse{42:16} %%
  \verse{42:17} Job died, old and full of days.%%
\end{inparaenum}
