\heading{1}{xxxx}

\begin{inparaenum}
  \hfil\verse{1:1} Obadiah's vision:\hfil%%
  
  \smallskip\pvaa{Thus says the Lord \textsc{God} to\footnotemark\ Edom:}%% The \smallskip needs to be here and not on the previous line, otherwise it messes up the centering of the title.
  \alt{concerning}%%
  
  \pvbb{``We have heard\footnotemark\ a report from the \textsc{Lord},}{a messenger has been sent out throughout the nations.}%%
  \ed{The critical apparatus makes frequent comparison to Jer~49, but is not cited here. The reader is encouraged to study the two chapters side-by-side, specifically \vref{Jer}{49}{9--15} as they correspond to \vref{Obad}{1}{1--6}.}%%
  
  \pvba{Get up! Let's go against her\footnotemark\ to battle.}%%
  \ie{Edom}%%
  
  \pvab{\verse{1:2} I've made you small among the nations~---}{you are greatly despised.}%%
  
  \pvaa{\verse{1:3} The arrogance\footnotemark\ of your heart has tricked\footnotemark\ you.\footnotemark}%%
  \halot{xxxx}{\textbf{presumption}}%%
  \alt{deceived}%%
  \lit{lifted you up.}%%
  
  \pvbb{O dweller in the clefts in the rock,}{his dwelling-place is lofty.}%%
  
  \pvbb{He says in his heart,}{`Who will bring me down to earth?'}%%
  
  \pvab{\verse{1:4} Although you soar like the eagle,\footnotemark}{and though you place your nest among the stars,}%%
  \alt{vulture}%%
  
  \pvbb{I will bring you down from there,''}{declares the \textsc{Lord}.}%%
  
  \pvbb{\verse{1:5} ``If robbers come in to you~---}{if plunderers come in by night~---}%%
  
  \pvcb{(how you've been cut off!)}{will they not steal enough for themselves?}%%
  
  \pvcb{If gleaners\footnotemark\ come to you,}{won't they leave gleanings?}%%
  \alt{grape harvesters}%%
  
  \pvbb{\verse{1:6} How Esau has been searched!}{his treasures are sought out!}%%
  
  \pvbb{\verse{1:7} You've been sent out to the border}{by all of your allies.\footnotemark}%%
  \lit{the men of your treaty.}%%
  
  \pvcb{They've forgotten you. The men of your peace}{have eaten\footnotemark}%%
  \ed{The stichs have been slightly reworked. It is literally ``They've forgotten you. They have eaten [next stich] the men of your peace\dots.''}
  
  \pvcb{your bread.}{They've placed a trap\footnotemark\ under you.}%%
  \halot{xxxx}{\textbf{(man-)trap}}%%
  
  \pvca{There is no understanding in him.}%%
  
  \pvab{\verse{1:8} On that day, shall I not,''}{declares the \textsc{Lord},}%%
  
  \pvbb{``destroy the wise men out of Edom,}{and understanding from the mountain of Esau?}%%
  
  \pvab{\verse{1:9} Your mighty men, O Teman, shall be struck down}{in order that every man in the mountain of Esau shall be cut off}%%
  
  \pvbb{by being murdered.\footnotemark\ \verse{1:10} Because of violence against your brother Jacob,}{shame shall cover you, and you shall be cut off forever.}%%
  \alt{slaughtered.}%%
  
  \pvab{\verse{1:11} In the day of your standing on the other side,}{in the day when non-Israelites took his property captive,}%%
  
  \pvbb{when\footnotemark\ foreigners entered his gate,}{and they cast lost on Jerusalem,}%%
  \ed{understood}%%
  
  \pvba{even you were one of them!}%%
  
  \pvbb{\verse{1:12} But you looked on the day of your brother,\footnotemark}{on the day of his estrangement;}%%
  \ca{prp \Hebrew{בְאחיך}}{it has been proposed to be ``on your brother''}%%
  
  \pvcb{neither did you rejoice over the children of Judah}{in the day of their going astray,}%%
  
  \pvcb{nor make your mouth great}{in the day of distress,}%%
  
  \pvbb{\verse{1:13} nor come through the gate of My people}{in the day of their disaster,\footnotemark}%%
  \halot{xxxx}{(final) \textbf{disaster}}%%
  
  \pvcb{nor did you, even you, look on its wickedness}{is the day of its disaster,}%%
  
  \pvcb{nor did you send against its outer wall}{in the day of its disaster,}%%
  
  \pvbb{\verse{1:14} nor stand on the crossroads}{to cut off its fugitives,}%%
  
  \pvcb{nor delivered its survivors}{in the day of its distress,}%%
  
  \pvbb{\verse{1:15} because the day of the \textsc{Lord} is near}{on all the nations.}%%
  
  \pvcb{As you've done, it's done to you:}{your reward will return on your own head.}%%
  
  \pvab{\verse{1:16} Because as you've drunk on My holy mountain,}{so shall all the nations drink unceasingly\footnotemark~---}%%
  \ca{mlt Mss \Hebrew{סָבִיב}}{multiple Hebrew manuscripts have ``all around'' [instead of ``unceasingly/continually'']}%%
  
  \pvbb{they shall drink and slurp,\footnotemark}{they shall be like they were note.}%%
  \ca{prp \Hebrew{ונָעו}}{it has been proposed to be ``and stumble/teeter/shake/tremble''}%%
  
  \pvac{\verse{1:17} On mount Zion}{there's deliverance}{and holiness; [why would it be ``there is holiness''?]\footnotemark}%%
  \ca{add?}{[this hemistich] added?}%%
  
  \pvbb{the house of Jacob shall possess}{their possessions.}%%
  
  \pvab{\verse{1:18} The house of Jacob shall be a[?] fire,}{the house of Joseph shall be a flame,}%%
  
  \pvbb{the house of Esau shall be stubble:\footnotemark}{they shall burn among them and consume them.}%%
  \alt{straw}%%
  
  \pvbb{There will be no survivors}{in the house of Esau}%%
  
  \pvba{because the \textsc{Lord} has spoken.}%%
  
  \pvbb{\verse{1:19} The south [subject?] shall possess mount Esau [make consistent with this one] and the lowlands of the Philistines,}{they shall possess}%%
  
  \pvba{the field\footnotemark\ of Ephraim and the field of Samaria and of Benjamin\footnotemark\ and Gilead.}%%
  \ca{\septuagint\ \Greek{τὸ ὄρος}, frt l \Hebrew{הַר}}{the Septuagint has ``the mountain,'' perhaps read ``mount''}%%
  \ca{xxxx}{xxxx}%%
  
  \pvba{\verse{1:20} The captives of this force of the children of Israel (that are of the Canaanites) are to[?] Zarephath,}%%
  
  \pvba{and the captives of Jerusalem who were in Sepharad shall possess the cities of the south.}%%
  
  \pvbb{\verse{1:21} And saviors shall go up on Mount Zion}{to judge the mountain of Esau,}%%
  
  \pvca{and the kingdom shall be the \textsc{Lord}'s.}%%
\end{inparaenum}
