\heading{12}{The Lord promises blessings upon Abram~--- Abram and his company journey to Canaan~--- Abram build altars and calls upon the Lord~--- }

\begin{inparaenum}
    \verse{12:1} The \textsc{Lord} said to Abram, ``Go\lit{for yourself} out from your land, from your family, from your father's house, to the land that I will show you.%%
    \verse{12:2} I will make a great nation of you. I will will you. I will make your name great. You will be a blessing.%%
    \verse{12:3} I will bless those who bless you and curse those that curse you.\lit{and those that curse you I will curse.} In you\ie{Because of you, Through you} shall all of the families of the earth be blessed.''%%
    \verse{12:4} Abram went on according as the \textsc{Lord} had spoken to him, and Lot went with him. Abram was 75~years old when he left\lit{went out of} Haran.%%
    \verse{12:5} Abram took his wife (Sarai), his nephew\lit{brother's son} Lot, all the goods\halot{furnishings, gear, utensils} that they had acquired,\alt{gathered} and the people\halot{acquire people, rear persons (slaves?)} whom they had gathered\ed{Converted? Bought?} in Haran, and they went to go towards the\understood\ land of Canaan, and they came into the land of Canaan.%%
    \verse{12:6} Abram passed from one side of the land to the other, to the place of Shechem, to the great tree\alt{oak} of Moreh. And the Canaanite was then in the land.%%
    \verse{12:7} The \textsc{Lord} appeared to Abram, and said, ``I will give this land to your posterity.'' He\ed{Abram} built an altar there to the \textsc{Lord} who had appeared to him.%%
    \verse{12:8} He moved on from there to the mountain on the east of Bethel, and pitched\lit{stretched out} his tent with Bethel to the west and Ai\alt{Hai} to the east. He built an altar there to the \textsc{Lord} and called on the name of \textsc{God}.%%
    \verse{12:9} Abram traveled, going and traveling to the south.%%
    
    \verse{12:10} There was a famine in the land, and Abram went towards Egypt to stay as a foreigner and sojourn there because the famine in the land was grievous.\alt{heavy, oppressive}%%
    \verse{12:11} As he got near\alt{drew near, approached} to enter Egypt, he said to his wife Sarai, ``Please, I know that you are a beautiful woman to look at.%%
    \verse{12:12} When the Egyptians see you, they will say, `This is\understood\ his wife,' and they will kill me and keep you alive.%%
    \verse{12:13} Please say that\understood\ you're my sister in order that everything will be fine on my account\alt{for my sake} and that\understood\ my soul may live on account of you.''\lit{on your account, for your sake.}%%
    \verse{12:14} When Abram entered Egypt, the Egyptians saw the woman, that she was really beautiful.%%
    \verse{12:15} %%
    \verse{12:16} %%
    \verse{12:17} %%
    \verse{12:18} %%
    \verse{12:19} %%
    \verse{12:20} %%
\end{inparaenum}
