\heading{17}{The Lord commands Abram to be perfect~--- continuation of Abrahamic covenant~--- name changed to Abraham~--- Canaan given to Abraham~--- covenant of circumcision set forth~--- Sarai changed to Sarah~--- Isaac promised~--- covenant to continue through him~--- Abraham and his household are circumcised}

\begin{inparaenum}
    \verse{17:1} Abraham was 99 years old when the \textsc{Lord} appeared to him. He said, ``I am God Almighty\footnote{Omnipotent~--- pl. for violence XXXX=[check unabridged Kohler-Baumgartner]}~--- walk before Me and be perfect.%%
    \verse{17:2} I will give\footnote{set up, establish} My covenant between us and I will greatly multiply you.''%%
    \verse{17:3} Abraham fell on his face as God spoke with him, saying,%%
    \verse{17:4} ``Pay attention. My covenant is with you~--- you shall be like a father of a multitude of nations.%%
    \verse{17:5} You shall no longer be called Abram, but rather Abraham, for I have made you a father of a multitude of nations.%%
    \verse{17:6} You shall be exceedingly fruitful. You shall become many nations. Kings will be among your posterity.\footnote{lit., Kings will come from you.}%%
    \verse{17:7} I will establish a covenant between us and also between your offspring.\footnote{progeny} It shall be an everlasting covenant: to be the God of you and your children.%%
    \verse{17:8} I will give you and your descendants\footnote{lit, seed after thee} the land of your sojournings and the land of Canaan as an everlasting possession for I have become their God.''%%
    \verse{17:9} God said to Abraham, ``You and the generations that follow shall honor this covenant.%%
    \verse{17:10} This is My covenant, between Me and you and your progeny,\footnote{lit., your seed after you} that you shall observe carefully: every male shall be circumcised.''%%
    \verse{17:11} You shall circumcise the flesh of your foreskin as a sign of the covenant between Me and you.%%
    \verse{17:12} Every male in your generations shall be circumcised when they are eight days old: those born in a house and those, not your offspring (the children of a resident alien), that are bought with money.%%
    \verse{17:13} Those born in a house and those bought with your money shall certainly be circumcised. My covenant shall be an everlasting covenant in your flesh.%%
    \verse{17:14} The uncircumcised male, the flesh of whose foreskin is uncircumcised, shall be cut off\footnote{Great word choice} from My people: he has broken My covenant.%%
    \verse{17:15} God said to Abraham, ``Your wife will no longer be known as Sarai, but Sarah.%%
    \verse{17:16} I will bless her. Additionally, I will give you a son from her. I will bless her that she shall become a nation~--- people's kings will be among her posterity.''%%
    \verse{17:17} And Abraham fell on his face and laughed. He said in his heart, ``Shall a son be born to a hundred year old man? And shall a ninety year old woman bear him?''\footnote{More lit., Shall one be born to an hundred year old? Shall a daughter who is ninety bear?}%%
    \verse{17:18} Abraham said to God, ``O that Ishmael might live in Your presence!''%%
    \verse{17:19} And God said, ``On the contrary, your wife, Sarah, shall bear a son. You shall name him Isaac. I will establish My covenant with him and it shall be an everlasting covenant with his descendants.%%
    \verse{17:20} Concerning Ishmael, I have heard you. Look, I will bless him, make him fruitful, greatly multiply him: he shall have\footnote{beget, sire} twelve princes. I will make him a great nation.%%
    \verse{17:21} My covenant shall be with Isaac whom Sarah shall bear to you at an\footnote{this} appointed time next year.''%%
    \verse{17:22} He finished speaking with him and God left\footnote{went up from} Abraham.%%
    \verse{17:23} Abraham took his son, Ishmael, all who were born in his house, those bought with his money (all of the men in Abraham's house) and circumcised them\footnote{lit., the flesh of their foreskins} that day like God had told him to.%%
    \verse{17:24} Abraham was ninety-nine years old when the flesh of his foreskin was circumcised.%%
    \verse{17:25} His son, Ishmael, was thirteen when the flesh of his foreskin was circumcised.%%
    \verse{17:26} \footnote{In}That same day, Abraham and his son, Ishmael, were circumcised.%%
    \verse{17:27} All of the men in the house~--- those born in the house and those bought with a resident alien's money~--- were circumcised as well.\footnote{lit., with him}%%
\end{inparaenum}
