\heading{26}{Isaac and his family go down to live in Gerar per the Lord's instructions~--- Isaac says that his wife is his sister; Abimelech calls him out on it~--- digging of wells and related disputations}

\begin{inparaenum}
  \verse{26:1} There was a famine in the land, besides the previous famine that had happened\lit{been} in Abraham's days. So Isaac went to Gerar, to Abimelech king of the Philistines.\lit{to Abimelech king of the Philistines to Gerar.}%%
  \verse{26:2} But the \lord\ appeared to him and said, ``Don't go down to Egypt; stay in the land that I will tell you.%%
  \verse{26:3} Stay\halot{xxxx}{\textbf{stay as foreigner and sojourner}} in this land. I will be with you. I will bless you, because I will give all these lands to you and your posterity~--- I will carry out the oath\alt{swearing} that I swore to your father Abraham.%%
  \verse{26:4} I will multiply your posterity as the stars in the sky.\alt{heavens} I will give all these lands to your posterity. All the nations of the earth shall wish a blessing to themselves [because of] your posterity.\ed{This is a bit confusing and \textsc{halot} isn't too helpful. It gives this \textit{hithpael} form as meaning ``\textbf{wish a blessing to oneself} (to one another?), with \textit{b}\superit{e} \haref{Gn}{22}{18} (Abraham)\haref{}{26}{4} (Isaac)'' which is where I derived the present translation from.}%%
  \verse{26:5} Because that Abraham\ca{\sampen\septuagint\ + \Hebrew{אָבִיךָ}}{the Samaritan Pentateuch and Septuagint add ``your father''} heard My voice and obeyed My statutes and My commandments and My laws.''%%
  \verse{26:6} Isaac lived in Gerar.%%
  \verse{26:7} The men of the place asked him about his wife, and he said, ``She is my sister'' because he was afraid to say ``My wife,\ca{ins c \sampen\septuagint\ \Hebrew{הִיא}}{insert with the Samaritan Pentateuch and Septuagint ``she [is]''} lest the men of the place kill me for Rebecca because she's beautiful.''\lit{of good appearance.''}%%
  \verse{26:8} When he'd been there a while,\lit{When the days became long to him there} Abimelech king of the Philistines looked down through his window and saw Isaac was fondling\halot{xxxx}{with \textit{'}\superit{\=e}\textit{t} \textbf{fondle} (a woman) \haref{Gn}{26}{8}} his wife Rebecca.%%
  \verse{26:9} Abimelech called to Isaac, and said, ``She's definitely\lit{surely} your wife! Why\lit{How} did you say, `She's my sister'?'' And Isaac said to him, ``Because I said, `Lest I die for her.'\thinspace''%%
  \verse{26:10} Abimelech said, ``What is this thing you've done to us? One of the people, like nothing,\lit{like a little thing} could have laid with your wife, and you would have brought guilt on us.''%%
  \verse{26:11} So Abimelech commanded all the people,\ca{\sampen\septuagint\ \Hebrew{עַמּוֺ}}{the Samaritan Pentateuch and Septuagint have ``his people''} saying, ``Whoever touches this man or his wife shall certainly be killed.''\lit{shall die the death.''}%%
  \verse{26:12} Isaac sowed in that land and found a hundredfold in that field; the \lord\ blessed him.%%
  \verse{26:13} The man became great, becoming continually greater, until he had become very great.%%
  \verse{26:14} He had possession of a flock and possession of a herd and many servants. The Philistines envied him.%%
  \verse{26:15} All the wells that his father's servants had dug during his father Abraham's days; the Philistines stopped them and filled them with dust.%%
  \verse{26:16} Abimelech said to Isaac, ``Leave us\lit{Go from us} because you're much mightier than we are.''%%
  \verse{26:17} So Isaac went from that place and encamped in Garar and stayed there.%%
  \verse{26:18} Isaac returned and dug the water wells that they'd\ca{prb l c \sampen\septuagint\vulgate\ \Hebrew{עַבְדֵי} cf 15.19.25 et \peshitta\targum\super{J}}{probably read with the Samaritan Pentateuch, Septuagint, and Vulgate which have ``servants'' [in construct, so ``his father Abraham's servants'']; compare verses~15, 19, and~25 and the Peshitta and Targum} dug in the days of his father, Abraham, which the Philistines had stopped after Abraham's death. He named them\lit{called their names} after the names his father had called them.%%
  \verse{26:19} Isaac's servants dug in the valley and they there found a well of running water.%%
  \verse{26:20} The shepherds of Gerar conducted a lawsuit\halot{xxxx}{\textbf{dispute}, \textbf{quarrel} (in public, w. words, complaints, assertions, reproaches); \textbf{conduct a} (legal) \textbf{case}, \textbf{lawsuit}: absolute (state) \haref{Gn}{26}{21}} with Isaac's shepherds, saying, ``The water is ours!''\lit{To us the water.} And he called the name of the well ``Quarrel'' because they'd quarreled with him.%%
  \verse{26:21} They dug another well and they also quarreled for it.\ed{Possibly ``[T]hey also disputed over it.''} He called the place Sitnah.\lit{Hatred.}\halot{xxxx}{accusation}%%
  \verse{26:22} He moved on from there and dug another well which they didn't quarrel over.\lit{and they didn't dispute about it.} He called the place Rehoboth,\lit{Wide open} and said, ``For now the \lord\ has made room\alt{opened wide} for us and we shall be fruitful in this land.''%%
  \verse{26:23} He went from there to Beersheba.%%
  \verse{26:24} The \lord\ appeared to him in that night, and said, ``I am the God of your father, Abraham. Don't be afraid because I am with you and have blessed you. I will multiply your posterity on account of my servant Abraham.''%%
  \verse{26:25} He built an altar there and called on the name of the \lord. He pitched his tent there and Isaac's servant dug a well there.%%
  \verse{26:26} Abimelech went to him from Gerar along with\lit{and} his friend Ahuzzath and the captain of his army, Phichol.%%
  \verse{26:27} Isaac said to them, ``Why have you come to me, seeing as you hate me and have driven me away from you?''%%
  \verse{26:28} They said, ``We've really seen that the \lord\ is with you; so we say, `Please let there be a covenant between us\ca{\missing\ Vrs, dl?}{[``between us'' is] missing in multiple versions, deleted?} and you\lit{between you} so let us make a covenant with you:%% Keep "\lit{between you}" because it matters in the context of the critical apparatus note.
  \verse{26:29} don't do evil with us because we haven't hurt you;\lit{touched you}\alt{done you harm} and because we have only done good with you and sent you away in peace. You are now blessed of the \lord.''%%
  \verse{26:30} He made a feast for them, and they ate and drank.%%
  \verse{26:31} And they got up in the morning\ed{\Hebrew{שׁכם} can have the connotation to be early, although that's not given with every definition.} and swore to each other,\alt{one to another}\ca{\sampen\septuagint\ \Hebrew{לְרֵעֵהוּ}}{the Samaritan Pentateuch and Septuagint have ``to his neighbor''} and Isaac sent them and they went away from him in peace.%%
  \verse{26:32} In that day, Isaac's servants came and told him about\halot{xxxx}{on account of} the well that had been dug, and they said to him, ``We have found water.''%%
  \verse{26:33} He called it Sheba, thus the name of the city is Beersheba to this day.%%
  
  \verse{26:34} Esau was 40~years old when took to wife Judith, the daughter of Beeri the Hittite, and Basmath, the daughter of Elon the Hittite.\ca{frt l c \sampen\septuagint*\peshitta\ \Hebrew{הַחִוִּי}}{perhaps read with Samaritan Pentateuch, Septuagint, and Peshitta ``Hivite''}%%
  \verse{26:35} They were a bitterness\alt{affliction, grief} to Isaac and Rebecca's spirit.%%
\end{inparaenum}
