\heading{3}{The snake convinces Eve to eat the forbidden fruit~--- she does so and then convinces Adam to do likewise~--- they are reproved by the Lord~--- the snake is cursed~--- curses are given to Adam and Eve~--- they are driven out of the Garden of Eden~--- the Lord has cherubim protect the tree of life}

\begin{inparaenum}
    \verse{3:1} The snake was subtler\alt{cleverer, shrewder} than all the other field animals that the \textsc{Lord} God had created. He\footnote{Theological implication here is that the snake, representing the adversary, is male; otherwise, ``she'' or ``it'' would be acceptable pronouns.} said to the woman, ``Has God really said that you shall not eat from every tree of the Garden?''%%
    \verse{3:2} The woman replied to the snake, ``We may eat the fruit of the trees\lit{From the fruit of the trees} of the garden,%%
    \verse{3:3} `but from the fruit of the tree that's in the middle of the garden,' God said, `don't eat it and don't touch it or else you'll die.'\thinspace''%%
    \verse{3:4} The snake said to the woman, ``You won't certainly die%%
    \verse{3:5} because God knows that\heb{כִּי}{See further in Waltke and O'Connor 39.3.4.} when\lit{in the day} you\heb{plural}{} eat from it, your eyes will be opened and you'll be like God, knowing good and evil.''%%
    \verse{3:6} The woman saw that the tree was good for food, that it is wanting\alt{craving, longing, pleasant, beautiful, lovely} to the eyes~--- that the tree is pleasurable to make one wise. So she took some of its fruit and ate. She also gave some of it\understood\ to her husband who was\understood\ with her, and he ate.%%
    \verse{3:7} Both of their eyes were opened and they knew that they were naked. They sewed together fig leaves and made loincloths\alt{aprons, girdles} for themselves.%%
    \verse{3:8} They heard the voice of the \textsc{Lord} God, walking in the breeze of day.\footnote{It is unclear if this is the evening or morning wind. The Hebrew is \Hebrew{לְרוּחַ הַיּוֺם}, literally meaning ``in/during the wind/breeze of day.''} Adam and his wife were hiding themselves from the face of the \textsc{Lord} God who\understood\ was in the midst of the Garden.%%
    \verse{3:9} The \textsc{Lord} God called to Adam and said to him, ``Where are you?''%%
    \verse{3:10} And he replied, ``I heard your voice in the Garden and I was afraid because I'm naked~--- that's why\understood\ I hid.''\lit{hid myself.}%%
    \verse{3:11} He said, ``Who told you that you're naked? Have you eaten from the tree that I commanded you to not eat from?''\lit{From the tree that I commanded you to not eat from, have you eaten?}%%
    \verse{3:12} Adam said, ``The woman, whom you gave to me, she gave me fruit\understood\ from the tree and I ate.''%%
    \verse{3:13} The \textsc{Lord} God said to the woman, ``What is this that you've done?'' And the woman said, ``The snake tricked\alt{deceived. The niphal meaning is to ``give oneself false hopes.''} me and I ate.''%%
    \verse{3:14} The \textsc{Lord} God said to the snake,\smallskip%%
    
    \pvbd{``Because you've done this,}{you shall be cursed}{more than every animal,\footnotemark}{and more than every field animal.}%%
    \fntlit{from all the animals}%%
    
    \pvbc{You shall go about on your belly}{and you shall eat loose soil\footnotemark}{all the days of your life.}%%
    \fnthalot{loose \textbf{soil}, crumbling to \textbf{dust}.}%%
    
    \pvad{\vn{3:15} I will put enmity\footnotemark}{between you and the woman~---}{between your seed}{and her seed.}%%
    \fntalt{hostile intention}%%
    
    \pvbb{He will crush your head,}{you will snap at\footnotemark\ his heel.''}\bigskip%%
    \fntalt{snatch at}%%
    
    {\noindent\verse{3:16} He said to the woman:}%%
    
    \pvbc{``I will greatly multiply}{your hardships\footnotemark\ and pregnancy.\footnotemark}{You will bear children in pain.\footnotemark}%%
    \stepcounter{footnote}%%
    \fntalt{pain}%%
    \stepcounter{footnote}%%
    \fnthalot{traditionally \textbf{pregnancy}; other: \textbf{sensory pleasure}}%%
    \stepcounter{footnote}%%
    \fnted{This raises an interesting theological question: Could Eve previously bear (had previously born?) children without pain? Or is this an anagogical sense that teaches that child birth in Heaven is painless?}%%
    
    \pvbb{You shall long\footnotemark\ for your husband}{and he shall govern you.''}\bigskip%%
    \stepcounter{footnote}%%
    \fnthalot{\Hebrew{שׁוק}: by-form of Heb.\ II \Hebrew{שׁקק}; Arb.\ \textit{\v s\=aqa(w)} I and II to fill with longing, desire, craving\dots\ Eth.\ \textit{\d{s}\v eh\v eqa} to wish, desire}%%
    
    \verse{3:17} He said to Adam, ``Because you have listened to what your wife has said\lit{the voice of your wife} and have eaten from the tree that I commanded you,\lit{saying} `Don't eat from it,'%%
    
    \pvba{The ground will be cursed on your account.}%%
    
    \pvbb{You shall eat in hardship and pain\footnotemark}{all the days of your life.}%%
    \stepcounter{footnote}%%
    \fntheb{עֶצֶב}{means both of these.}
    
    \pvab{\vn{3:18} Thorn bushes and thistles\footnotemark\footnotemark\ shall sprout up before you.\footnotemark}{You shall eat the green plants\footnotemark\ of the field.}%%
    \stepcounter{footnote}%%
    \fnthalot{Subspecies of \textit{Centaurea pallescens}: a sort of \textbf{thistle}}%%
    \stepcounter{footnote}%%
    \footnotetext{Both of these are singular, but better understood as plural.}%%
    \stepcounter{footnote}%%
    \fntlit{to you}
    \stepcounter{footnote}%%
    \fnthalot{coll.\ \textbf{green plants}: weeds, grass, vegetables, cereals, growing during rainy season, not perennials}
    
    \pvab{\vn{3:19} By the sweat of your face}{shall you eat bread}%%
    
    \pvbb{until you return to the ground}{because you've been taken from out of it.}%%
    
    \pvbb{Because you are dirt,}{and to dirt shall you return.''}%%
    
    \verse{3:20} Adam called his wife\lit{called the name of his wife} Eve because she is the mother of every living thing.%%
    \verse{3:21} The \textsc{Lord} God made animal skin shirts\halot{long \textbf{shirt-like (under-)garment}} for Adam and his wife, and He clothed them.%%
    
    \verse{3:22} The \textsc{Lord} God said, ``Surely Adam has become like one of us,\ed{This is an interesting theological point (``one of us'') because it assumes more than one God. However, as noted in Gen 1\thinspace:\thinspace1\textit{a}, \Hebrew{אֱלֹהִים} is always conjugated in the 3ms.} understanding good and evil. So now, lest he reaches\lit{stretches} his hand out to also take from the tree of life, eats, and lives forever~---''%%
    \verse{3:23} The \textsc{Lord} God drove him out of the Garden of Eden to till the ground where he'd been brought.%%
    \verse{3:24} He banished\alt{drove out} Adam. He had cherubim and the flame\halot{Metaphorical for ``blade.''} of a sword sojourn to the east of the Garden of Eden, turning this way and that to guard the path to\understood\ the tree of life.%%
\end{inparaenum}
