\heading{49}{xxxx}

\begin{inparaenum}
  \verse{49:1} ``%%
  \verse{49:2} %%
  \verse{49:3} %%
  \verse{49:4} %%
  \verse{49:5} %%
  \verse{49:6} %%
  \verse{49:7} %%
  \pb \verse{49:8} To Judah~--- your brothers will praise you.\pa Your hand will be on your enemies' necks.%%
  
  \pa Your siblings\lit{Your father's children} will bow down to you.%%
  
  \pa \verse{49:9} Judah is like a young lion\pa gone up from the prey, my son.%%
  
  \pa Bent and crouched like a lion\pa or a lioness.\pa Who shall wake\alt{stir} him?%%
  
  \pa \verse{49:10} Neither shall the scepter\pa nor the lawgiver leave Judah%%
  
  \pa Until Shiloh\ed{xxxx} comes. The nation's obedience shall be to Him.%%
  
  \pa \verse{49:11} He'll tie his donkey to the vine\pa and his\ie{his donkey's} colt to the choice vines.%%
  
  \pa He'll wash his\ed{His?} clothing in wine~---\pa even in the blood of grapes.%%
  
  \pa \verse{49:12} Eyes red with wine;\pa teeth white with milk.%%
  
  \verse{49:13} %%
  \verse{49:14} %%
  \verse{49:15} %%
  \verse{49:16} %%
  \verse{49:17} %%
  \verse{49:18} %%
  \verse{49:19} %%
  \verse{49:20} %%
  \verse{49:21} %%
  \pb \verse{49:22} Joseph is a fruitful son,\pa even a fruitful son by a well;%%
  
  \pa xxxx%%
  
  \pa \verse{49:23} Archers have provoked, shot at,\pa and hated him,%%
  
  \pa \verse{49:24} but his bow is fair,\pa his arms and hands are strong%%
  
  \pa because of the Mighty One of Jacob. From there is the shepherd, the stone of Israel.%%
  
  \pa \verse{49:25} The God of your father will help you:\pa the Almighty will bless you%%
  
  \pa with blessings from heaven above, with blessings of the deep lying under,%%
  
  \pa from the breast and womb.%%
  
  \pa \verse{49:26} Your father's blessings shall be greater than\pa that of your forebearers%%
  
  \pa even unto the heights\alt{bounds} of the everlasting hills.%%
  
  \pa These shall be upon Joseph's head~---\pa on the crown of him who was separated from his brothers.%%
  
  \verse{49:27} %%
  \verse{49:28} %%
  \verse{49:29} %%
  \verse{49:30} %%
  \verse{49:31} %%
  \verse{49:32} %%
  \verse{49:33} %%
\end{inparaenum}
