\heading{7}{The Lord gives instructions on how many and of what kind of animals to bring onto the Ark, and how long the Flood shall last~--- everything is gathered into the Ark~--- the Flood~--- effects of the Flood~---}

\begin{inparaenum}
    \verse{7:1} The \textsc{Lord} said to Noah, ``You and all of your household, go into the Ark because I've seen you\ed{singular} righteous before Me in this generation.%%
    \verse{7:2} Of all the large, clean\ie{cultically, ritualistically} animals, you shall bring seven pairs\lit{seven seven} with\lit{to} you, male and female.\ed{What kind of split? It would make sense to bring more females than males.} But of the beasts that are not clean, you shall bring\understood\ two, male and female.%%
    \verse{7:3} Additionally, bring\understood\ the birds of the sky by sevens, male and female, in order to keep offspring alive on the face of all the earth.%%
    \verse{7:4} Because \lit{in}seven days hence I will let it rain for forty days and forty nights on the earth, until\ed{The conjuctive \textit{waw} (\Hebrew{וְ})) is generally rendered ``and,'' but can also be used to show the consequences of actions, so I feel justified in rendering it as ``until'' here.} everything living\halot{\textbf{what subsists}, \textbf{what is living}} that I have made is destroyed from off the face of the land.''%%
    \verse{7:5} So Noah did according to everything that the \textsc{Lord} had commanded him.%%
    
    \verse{7:6} Noah was\lit{was a son of} 600~years old when the deluge of waters was\ed{Not necessarily ``began'' because the Flood only lasted 40~days. This is simply saying that the Flood happened during the 600$^\text{th}$ year of Noah's life.} on the earth.%%
    \verse{7:7} Noah went in, and his sons and his wife and his sons' wives went in\understood\ with him into the Ark, from the presence of the waters of the Flood.\lit{deluge}%%
    \verse{7:8} The clean animals, the unclean animals, the birds, every reptile of the ground~---%%
    \verse{7:9} they came in pairs\lit{two by two} to Noah into the Ark, male and female, as God had commanded Noah.%%
    \verse{7:10} And after seven days, the waters of the Flood were on the earth.%%
    \verse{7:11} In the six hundredth year of Noah's life, in the second month, on the seventeenth day of the month, in that day there was\understood\ a breach\halot{a forced breach} of all the fountains\alt{springs} of the great deep\halot{\textbf{primeval ocean}, \textbf{deep} Gn~$1_2$\dots\ \textbf{deeps of sea} Ex~$15_5$ (quasi-mythological)\dots\ \textbf{subterranean water} Dt~$8_7$} and the windows\halot{through which rain falls} of the sky were opened.%%
    \verse{7:12} The rain was on the earth for forty days and forty nights.%%
    \verse{7:13} On the same day,\ie{the first day of the Flood} Noah went out with\lit{and} Shem, Ham, and Japheth (Noah's sons), Noah's wife, and the three wives of his sons with them, into the Ark.%%
    \verse{7:14} They, every living thing of their own species, every large animal of their own species, every reptile of their own species that flocks\halot{\textbf{swarm}, \textbf{teem} (of vast numbers of creatures in water, on ground in woods; in random movement)} on the earth, every bird of their own species~--- every bird,\halot{\textbf{birds}, (creatures) \textbf{with wings}} everything that has wings.%%
    \verse{7:15} They went in to Noah, into the Ark, in pairs\lit{two by two}~--- all living creatures in which is the breath of life.%%
    \verse{7:16} Those who came, male and female of every animal, came according as God had commanded them. And the \textsc{Lord} shut them in.%%
    \verse{7:17} The Flood was on the earth for forty days, and the waters multiplied and lifted up the Ark above the ground.\lit{from upon the ground/earth.}%%
    \verse{7:18} The waters were mighty and multiplied greatly upon the earth, and the Ark went about on the surface\lit{face} of the waters.%%
    \verse{7:19} The waters were incredibly mighty\alt{The waters swelled/rose greatly} on the earth, and covered all the high mountains that are under the heavens~---%%
    \verse{7:20} the waters swelled up fifteen cubits\footnote{6.858\thinspace m (22.5\thinspace ft)} and covered the mountains.%%
    \verse{7:21} Every creature that moves\lit{swarms, teems} on the earth died: birds,\lit{of birds; this is true for most of the list.} large animals, animals, everything swarming thing\halot{tiny animals occurring in large numbers, in water Gn~$1_{20}$, in air Lv~$11_{20}$, on ground Gn~$7_{21}$.} that swarm on the earth, and all mankind%%
    \verse{7:22} %% add punctuation to end of v 21
    \verse{7:23} %%
    \verse{7:24} %%
\end{inparaenum}
