\heading{2}{Elijah is taken by a fiery carriage into heaven~--- the people doubt him and send a search party out to find his body~--- they're unsuccessful~--- the people petition Elisha to make their water potable, which he does~--- some children mock Elisha and are eaten by bears}

\begin{inparaenum}
  \verse{2:1} When the \textsc{Lord} took Elijah into Heaven in the gale,\alt{heavy windstorm} Elijah and Elisha departed from Gilgal.%%
  \verse{2:2} Elijah said to Elisha, ``Please stay here, because the \textsc{Lord} has sent me to Bethel.'' Elisha responded, ``As the \textsc{Lord} lives and as you\lit{your personality, individuality, life, desire, mood, state of mind, will, man, person} live, I will not leave\alt{abandon} you.'' So they went down to Bethel.%%
  \verse{2:3} The sons of the prophets who were in Bethel came out to Elisha, and said, ``Are you aware\lit{Do you know} that the \textsc{Lord} is taking your master from being in charge\lit{from your head} today?'' and he said, ``I know. Be quiet!''\alt{silent}%%
  \verse{2:4} Elijah said to him, ``Elisha, please stay here because the \textsc{Lord} has sent me to Jericho.'' And he responded, ``As the \textsc{Lord} lives and as you live, I will not leave you.'' So they came to Jericho.%%
  \verse{2:5} The prophets' children (in Jericho) approached Elisha, and said to him, ``Are you aware that the \textsc{Lord} is taking your master from being in charge today?'' and he said, ``I know. Be quiet!''%%
  \verse{2:6} Elijah said to him, ``Please stay here because the \textsc{Lord} has sent me to Jordan.'' And he responded, ``As the \textsc{Lord} lives and as you live, I will not leave you.'' So they both went on.%%
  \verse{2:7} Fifty men, sons of the prophets, went and stood aside, far off; and both of them stood by the Jordan.%%
  \verse{2:8} Elijah took his mantle,\ed{Traditionally rendered as \textit{mantle} and kept that way here.}\lit{robe (of state), fur garment} rolled it up, beat\alt{strike, hit} the water with it, and it divided on this side and the other.\alt{here and there} And they crossed over on dry ground.%%
  \verse{2:9} When they'd crossed over, Elijah said to Elisha, ``Ask what you'd have me do for you before I'm taken from you.'' And Elijah said, ``Please let there be a double portion of your spirit upon me.''%%
  \verse{2:10} He said, ``You've asked for difficult\alt{hard} thing. If you see me taken from you, then it shall be so to you; otherwise,\lit{and if not} it shall not be so.''%%
  \verse{2:11} As they were going about talking, a fiery chariot and fiery horses came\lit{separated} between them both, and Elijah went up in a gale to Heaven.%%
  \verse{2:12} Elisha saw it and cried out, ``My father, my father! The chariot of Israel and its horsemen!''\alt{riders} And he didn't see him again. He took hold\lit{strength} of his clothing and ripped it in two.\lit{two pieces}%%
  \verse{2:13} He picked up Elijah's mantle that had fallen off him and went back and stood on the bank of the Jordan.%%
  \verse{2:14} He took Elijah's mantle that had fallen off him, struck the water, and said, ``Where is the \textsc{Lord}, the God of Elijah? Where is He?''\lit{even He} He struck the water and it divided here and there, and Elisha crossed over.%%
  \verse{2:15} The children of the prophets in Jericho, on the opposite bank, saw him and said, ``Elijah's spirit has rested on Elisha!'' They came out to meet him and bowed to the ground in front of him.%%
  \verse{2:16} They said to him, ``Please pay attention. There are fifty men with your servants, qualified people\lit{sons, children}~--- please let them go. Seek your master lest the Spirit of the \textsc{Lord} takes him and throws him down onto a mountain or into a valley.'' And he replied, ``You shall not send.''\ie{You shall not send any men to go try and find him. It would be futile.}%%
  \verse{2:17} They strongly urged him until he was ashamed and said, ``Send.'' So they sent fifty men who\lit{and they} sought for three days, but didn't find him.%%
  \verse{2:18} They returned to him, and he was living in Jericho. He said to them, ``Didn't I tell you, `Don't go?'\thinspace''%%
  
  \verse{2:19} The men of the city said to Elisha, ``Look, please, the city's position is good, as my master sees it, but the water is bad, the land is barren.''%%
  \verse{2:20} He replied, ``Bring me a new dish and put salt in it.'' So they brought it to him.%%
  \verse{2:21} He went out to the water's source and threw salt in it,\lit{there} and said, ``Thus saith the \textsc{Lord}, `I've made this water drinkable. Death and sterility shall no longer come from it.'\thinspace''%%
  \verse{2:22} The waters are drinkable even\understood\ until now, according to the words that Elisha had spoken.%%
  
  \verse{2:23} And he went from there up to Bethel, and as he was going up on the highway, little boys came out of the city and mocked him, saying unto him, ``Go up, bald head! Go up, bald head!''%%
  \verse{2:24} And he turned back, faced them, and cursed them in the name of the \textsc{Lord}. And two mama bears came forth and out of the woods and tore 42 children among them.%%
  \verse{2:25} He then went from there up to Mount Carmel, and from there he returned to Samaria.%%
\end{inparaenum}
