\heading{38}{xxxx}

\begin{inparaenum}
    \verse{38:1} %%
    \verse{38:2} %%
    \verse{38:3} %%
    \verse{38:4} %%
    \verse{38:5} %%
    \verse{38:6} %%
    \verse{38:7} %%
    \verse{38:8} %%
    \verse{38:9} %%
    
    \verse{38:10} %%
    \verse{38:11} %%
    \verse{38:12} %%
    \verse{38:13} %%
    
    \verse{38:14} %%
    \verse{38:15} %%
    \verse{38:16} %%
    
    \verse{38:17} Thus says my master, the \textsc{Lord}: ``Aren't you he of whom I spake in past times through the hands of my servants, the prophets of Israel, who prophesied for years and bade them in those days that I would bring you against them?%%
    
    \verse{38:18} It shall be that in that day, when Gog comes against the land of Israel,'' the Lord \textsc{God} confirms, ``that My face will show how angry I am.%%
    \verse{38:19} In my zeal and my fiery wrath I have spoken: `There will be a great commotion throughout the land of Israel in that day.%%
    \verse{38:20} All the fish in the ocean, the birds in the sky, the beasts in the field, everything creeping along, all the men on the earth shall shake\footnote{alt., quake} at my presence. The mountains shall be thrown down, the cliffs shall fall, every wall shall fall to the earth.'%%
    \verse{38:21} I will proclaim through all the mountains\footnote{lit., my mountains. Since everything is the Lord's, \textit{the} mountains works just fine.} that a sword shall be brought against him,'' confirms the Lord \textsc{God}. ``Every man's sword shall be brought against his brother.%%
    \verse{38:22} I will judge\footnote{alt., plead} him through pestilence\footnote{alt., \textit{with him} or \textit{for him} through/with pestilence} and bloodshed, an abundance of rain and hailstorms, fire and brimstone, will I rain on him and the people who are with him.%%
    \verse{38:23} I have boasted, sanctified, and made Myself known in the eyes of many nations: thall shall know that I am the \textsc{Lord}.''%%
\end{inparaenum}
