\heading{1}{The Lord continues to love His people who in turn reject Him~--- the Lord enumerates the ways in which His people have rejected Him~--- they offer impure sacrifices}

\begin{inparaenum}
  \verse{1:1} The pronouncement\lit{burden}\alt{oracle, prophecy} of the \textsc{Lord} to Israel by the hand of Malachi:\smallskip%%
  
  \pvab{\vn{1:2} ``\thinspace`I have loved you,' says the \textsc{Lord}.}{Yet you say, `How have you loved us?'}%%
  
  \pvbd{`Isn't Esau Jacob's brother?'}{A declaration of the \textsc{Lord}.\footnotemark}{`I have loved Jacob}{\vn{1:3} and hated Esau.}%%
  \fntca{dl m cs}{deleted to preserve the meter case}%%
  
  \pvbb{I have made his mountains desolate\footnotemark}{and his inheritance [shall be/will go?] to the jackals\footnotemark\footnotemark\ of the desert.'}%%
  \fntalt{deserted}%%
  \fnthalot{xxxx}{\textit{Canis aureus}}%%
  \fntca{frt l \Hebrew{נָתַתִּי}, prp \Hebrew{נְוֺת} cf \peshitta}{perhaps read ``I will give,'' it has been proposed to be ``xxxx'' compare the Peshitta}%%
  
  \pvab{\vn{1:4} But Edom says, `We have been decimated,\footnotemark}{but we will return and rebuild the ruined sites,'\footnotemark}%%
  \fntalt{shattered}%%
  \fntalt{desolate places}%%
  
  \pvbb{thus says the \textsc{Lord} of Hosts:}{`They shall build and I will destroy.'}%%
  
  \pvbb{They've called\footnotemark\ them `a territory of wickedness,'}{`a people whom the \textsc{Lord} has cursed forever.'}%%
  \fnted{This is not ``They've called \emph{to}'' because there is no expressly stated subject.}%%
  
  \pvab{\vn{1:5} Your eyes have seen, and you shall say,}{`The \textsc{Lord} is exalted beyond the border of Israel.}%%
  
  \pvbb{\vn{1:6} A son honors his\footnotemark\ father,}{a servant his master.}%%
  \fnted{understood}%%
  
  \pvcb{But if I am a father,}{where is My honor?}%%
  
  \pvcb{And if I am a master,}{where is My fear?\footnotemark}%%
  \fntalt{the reverence toward Me?}%%
  
  \pvcb{The \textsc{Lord} of Hosts says to you,}{``Priests have despised My name,''}%%
  
  \pvcb{yet you say,}{``How have we despised Your name?''}%%
  
  \pvbb{\vn{1:7} You have offered on My altar}{polluted\footnotemark\ bread.}%%
  \fnthalot{xxxx}{(cultic) \textbf{pollution}}%%
  
  \pvcb{Yet you say,}{``How have we polluted you?''}%%
  
  \pvcb{In that you say,\footnotemark\ ``The altar of the \textsc{Lord}}{is despised.''\footnotemark}%%
  \fntlit{By your saying, By your having said, By saying}%%
  \fntalt{contemptible.'}%%
  
  \pvbc{\vn{1:8} If you bring}{the blind\footnotemark to sacrifice~---}{there is no wrong in this.\footnotemark}%%
  \fntie{a blind animal}%%
  \fntlit{isn't it evil?}%%
  
  \pvcc{If you bring}{the lame\footnotemark\ or ill~---}{there is no wrong in this.}%%
  \fnted{could refer to an animal}%%
  
  \pvcc{Please offer it to the governor:\footnotemark}{will he be pleased with you, or will he accept you?'}{says the \textsc{Lord} of Hosts.\footnotemark}%%
  \fnthalot{xxxx}{a rather vague title}%%
  \fntlit{lift up your face?'}%%
  \fntca{dl m cs}{delete [this hemistich] to preserve the meter case}%%
  
  \pvbb{\vn{1:9} `And now, please, appease\footnotemark}{God\footnotemark\ and He will graciously provide for us.}%%
  \fntalt{flatter, entreat}%%
  \fntlit{the face of God}%%
  
  \pvcc{This has been from your hand:\footnotemark}{will He accept your offering?'}{says the \textsc{Lord} of Hosts.\footnotemark}%%
  \fntie{You did this:}%%
  \fntlit{will He lift up from your faces?'}%%%
  \fntca{dl}{delete [this hemistich]}%%
  
  \pvbb{\vn{1:10} `Who's even among you who would close the doors?\footnotemark}{You would not kindle on My altar without cause.\footnotemark}%%
  \fntie{of the temple?}%%
  \fntalt{in vain.}%%
  
  \pvcc{I have no pleasure in you,'}{says the \textsc{Lord} of Hosts.\footnotemark}{`I do not want an offering\footnotemark\ from you\footnotemark}%%
  \fntca{add}{[this hemistich] added}%%
  \fntalt{oblation, sacrifice, allegiance}%%
  \fntlit{your hand.}%%
  
  \pvbb{\vn{1:11} \footnotemark because from sunrise to sunset}{My name is great among the nations.}%%
  \fntca{11--13 prb add}{verses~11--13 probably added}%%
  
  \noindent In every place, incense and a clean\halot{xxxx}{\textbf{(cultically) clean}} offering are brought in to My name, because My name is great among the nations,' says the \textsc{Lord} of Hosts.%%
  \verse{1:12} `But you are profaning it by saying, ``The table of the Lord is impure;\alt{polluted, desecrated} its fruit, its food,\ca{prp \Hebrew{כֻלּוֺ}}{it has been proposed to be ``all of his''} is contemptible.''\ca{prb l \Hebrew{וְנִבְזֶה} (\Hebrew{ניבו} ortum ex dttg \Hebrew{נב}) cf \peshitta\targum}{probably read ``and is contemptible (``its fruit'' arising from dittography of \textit{nb}), compare the Peshitta and Targum}%%
  \verse{1:13} You say, ``What a bother!'' and become enraged\lit{sniff} at it,'\ca{1 \Hebrew{אותִי} (Tiq soph)}{1 has ``at Me'' (the Tiqqun sopherim)} says the \textsc{Lord} of Hosts. `You've brought in stolen goods, the lame, and the sick, and you bring in your offering. Should I accept\alt{be please with} this from your hand?' says the \textsc{Lord}.\ca{frt ins c pc Mss \septuagint\peshitta\super{W} \Hebrew{צְבָאוֺת}}{perhaps insert ``of Hosts'' with a few Hebrew manuscripts, the Septuagint, and Peshitta}%%
  
  \verse{1:14} \ca{14~add}{verse~14 added}`Cursed is the deceiver who has\lit{when there's} a male\halot{xxxx}{\textbf{male animal}, especially \textbf{ram}} in his flock and vows\ie{that it's clean} and\ed{understood: then} sacrifices a defective thing\halot{xxxx}{\textbf{defect}} to the Lord,\ca{prp c mlt Mss \Hebrew{לַיהוָה} vel \Hebrew{לִי}}{it has been proposed to be the Tetragrammaton in multiple Hebrew manuscripts} because I am a great king,' says the \textsc{Lord} of Hosts, `and My name is revered among the nations.'\thinspace''%%
\end{inparaenum}
