\heading{3}{The Lord will send a messenger to prepare His way~--- His people will be refined and purified~--- men rob God by offering poor sacrifices~---  xxxx}

\begin{inparaenum}
    \pvab{\vn{3:1} ``I will send My messenger}{and He shall prepare the\footnotemark\ way before Me.}%%
    \fntlit{a}%%
    
    \pvbb{The Lord whom you seek}{shall suddenly enter His temple.\footnotemark}%%
    \fnted{reversed the order of these stichs for idiomaticy}%%
    
    \pvca{The angel of the covenant in whom you delight~---He cometh,'' says the \textsc{Lord} of Hosts.\footnotemark}%%
    \fntca{prb add}{[this line] probably added}%%
    
    \pvab{\vn{3:2} ``But who shall endure\footnotemark\ the day of His coming?}{Who shall be standing when He appears?}%%
    \fntalt{abide, live through}%%
    
    \pvbb{For He is like refiner's fire}{and fuller's lye.\footnotemark}%%
    \fntalt{soap; the primary ingredient being lye}%%
    
    % Verse 3 should not be indented.
    \verse{3:3} For He shall sit as a refiner and purifier of silver and He shall purify the children of Levi and purge them like gold and silver. And they shall be an offering unto the \textsc{Lord} in righteousness.%%
    \verse{3:4} The offering of Judah and Jerusalem shall be a sweet savor unto the \textsc{Lord} even as in days of old~--- as in previous years.\smallskip%%
    
    \pvab{\vn{3:5} I will draw near to you for judgment.}{I have been a hasteful\footnotemark\ witness}%%
    \fnthalot{xxxx}%%
    
    \pvbb{against the sorcerers and adulterers,}{against those who swear falsely,}%%
    
    \pvbb{and against those who monetarily extort their employees\footnotemark~---}{widows, orphans, those who turn away the resident aliens,\footnotemark}%%
    \fntlit{oppressors of their hireling's wages}%%
    \fnted{the most exposed (or least protected) groups of people}%%
    
    \pvbb{and those who don't fear Me,''}{says the \textsc{Lord} of Hosts.}%%
    
    \pvab{\vn{3:6} ``For I, the \textsc{Lord}, have not changed.}{And you, children of Jacob, have not been consumed.}%%
    
    \pvab{\vn{3:7} Even from the days of your fathers have you shied away\footnotemark\ from My statutes}{and not taken heed.\footnotemark}%%
    \fntalt{turned away/aside}%%
    \fntca{frt ins \Hebrew{מִשְׁמַרְתִּי} cf 14}{perhaps insert ``to my commands,'' compare verse~14}%%
    
    \pvbb{Turn your back on Me and I will turn My back on you,''}{says the \textsc{Lord} of Hosts.}%%
    
    \pvbb{``Yet you say,}{`How do we turn away?'\footnotemark}%%
    \fntca{add?}{[this line] added?}%%
    
    \pvab{\vn{3:8} Will a man deceive\footnotemark\footnotemark\ God?}{Because you have been deceiving Me.}%%
    \fnthalot{\textbf{deceive} \haref{Ma}{3}{8f} \haref{Pr}{22}{23}; or perhaps \textbf{rob}}%%
    \ed{The critical apparatus gives \Hebrew{עקב} (grasp by the heel, cheat) instead of \Hebrew{קבע} (deceive, rob) for all instances in this verse and verse~9. \textsc{halot} says that this is ``perhaps metathesis of \Hebrew{קבע} to avoid assonance of \Hebrew{יַעֲקֹב}.''}%%
    
    \pvbb{But you say, `How have we been deceiving you?'}{By tithes and offerings.\footnotemark}%%
    \fntca{frt 1 \Hebrew{וּבַתּרומה} \Hebrew{בַּמּעשׂר} cf \peshitta\targum\vulgate, sed cf 9\super{b--b}}{perhaps 1 has ``in tithes and in offerings,'' compare the Peshitta, Targum, and Vulgate, but compare 9\super{b--b} [see following]}%%
    
    \pvac{\vn{3:9} You are cursed with a curse,}{and you rob Me,}{even this whole nation.\footnotemark}%%
    \fntca{add? prb 1 \Hebrew{כלו} \Hebrew{הֲגַם} et tr ad fin 8}{added? probably 1 has ``Will they also all this?'' and transposed to the end of verse~8}%%
    
    \pvac{\vn{3:10} Bring all of the tithes}{into the storehouse:}{there is food in My house.}%%
    
    \pvbb{Please, try Me on this,''}{says the \textsc{Lord} of Hosts,}%%
    
    \pvbb{``If I will not open the windows of heaven for you}{and empty out upon you a blessing \textit{ad infinitum}.}%%
    
    \pvab{\vn{3:11} I will rebuke the devourer for your sake.}{He shall not destroy the fruit of your land,}%%
    
    \pvbb{neither shall the vine of your the field be barren,''}{says the \textsc{Lord} of Hosts.}%%
    
    \pvab{\vn{3:12} ``All of the gentile nations shall call you blessed}{because you shall be a desirable\footnotemark\ land,''}%%
    \fntalt{delightful}%%
    
    \pvaa{says the \textsc{Lord} of Hosts.}%%
    
    \pvab{\vn{3:13} ``Your words have been harsh\footnotemark\ against Me,''}{says the \textsc{Lord}.\footnotemark}%%
    \fntalt{strong}%%
    \ca{prb ins \Hebrew{צְבָאוֺת} cf \septuagint\superit{L}}{probably insert ``of Hosts,'' compare the Septuagint (textus Graecus ex recensione Luciani)}%%
    
    \pvbb{``Yet you say,}{`How have we spoken against You?'}%%
    
    \pvab{\vn{3:14} ``You've said, `It is vain to serve God.}{And what profit is there in keeping His charges}%%
    
    \pvbb{and walking mournfully\footnotemark}{before the \textsc{Lord} of Hosts?}%%
    \fntlit{walking like mourners}%%
    
    \pvab{\vn{3:15} We now declare}{the proud happy.\footnotemark}%%
    \fntalt{arrogant, proud, haughty, insolent}
    
    \pvbb{Moreover, those who work iniquity\footnotemark\ prosper\footnotemark}{and also, those who tempt God escape.'\thinspace''}%%
    \fntalt{do wickedness}%%
    \fntlit{are built up, we build up}%%
    
    \pvac{\vn{3:16} Then those who feared the \textsc{Lord} spoke}{one to another,}{and the \textsc{Lord} observed and heard it.}%%
    
    \pvbb{And a scroll of remembrance was written before Him}{for those who reverenced the \textsc{Lord} and pondered\footnotemark\ on His name.}%%
    \fntalt{consider, reflect on}%%
    
    \pvac{\vn{3:17} ``They shall be Mine,''}{says the \textsc{Lord} of Hosts,}{``in the day I shall make them a peculiar\footnotemark\ treasure.}%%
    \fnted{see further in Appendix~\ref{app:peculiar}}%%
    
    \pvbb{And I shall show mercy on them, just like a man shows mercy}{on his son who serves him.}%%
    
    \pvab{\vn{3:18} You will return and differentiate between those who are righteous and those who are wicked~---}{between those who serve God and those who don't serve Him.''}%%
    
    \pvad{\vn{3:19} \footnotemark Because the day comes,}{burning as a furnace,\footnotemark}{when all the proud}{and all who work wickedness shall be stubble.\footnotemark}%%
    \fnted{In English translations, this is the beginning of chapter~4.}%%
    \fntca{\septuagint* + \Greek{καὶ φλέζει αὺτούς}}{the Septuagint (textus Graecus originalis) adds ``xxxx''}%%
    \fntalt{straw.}%%
    
    \pvbb{``The day is coming when He shall burn them up,''}{says the \textsc{Lord} of Hosts.}%%
    
    \pvbb{``It shall leave them neither}{root nor branch.}%%
    
    \pvab{\vn{3:20} ``Risen to you who fear My name,}{the Sun\footnotemark\footnotemark\ of Righteousness has risen with healing in His wings.}%%
    \fnted{In Egyptian iconography, there is an image of a beneficent sun god with wings. This is likely the same kind of imagery.}%%
    \fnted{We cannot ascribe word play (i.e., sun/son) here because it doesn't exist in Hebrew (i.e., \Hebrew{שֶׁמֶשׁ}/\Hebrew{בן}).}%%
    
    \pvbb{And you shall go forth and grow up}{as calves of the stall.}%%
    
    \pvab{\vn{3:21} You shall tread down the wicked for they shall be ashes}{under the soles of your feet}%%
    
    \pvbb{In the day that I shall do this,''}{says the \textsc{Lord} of Hosts.}%%
    
    \pvab{\vn{3:22} \footnotemark``Remember the Law of Moses, My servant,}{that I commanded him,}%%
    \fntca{22~add; \septuagint* tr~22 post~24}{verse~22 added; the Septuagint (textus Graecus originalis) transposes verse~22 after verse~24}%%
    
    \pvbb{In Horeb, for all Israel~---}{The judgments and statutes.}%%
    
    \pvab{\vn{3:23} \footnotemark I will send you}{Elijah the prophet}%%
    \fntca{23sq add}{verse~23 [and?] following add[ed/ition?]}%%
    
    \pvbb{Before the coming of the day of the \textsc{Lord},}{which is great and dreadful.\footnotemark}%%
    \fntalt{awe-inspiring, terrifying.}%%
    
    \pvab{\vn{3:24} \footnotemark He shall turn the fathers' hearts to the children,}{and the children's hearts to their father,}%%
    \fntca{23sq add}{verse~23 [and?] following add[ed/ition?]}%%
    
    \pvbb{otherwise I will come and smite}{the earth with a curse.''\footnotemark}%%
    \fntalt{destruction.''}%%
\end{inparaenum}
