\heading{15}{Moses and the people sing a song praises the Lord for their deliverance from Egypt~--- they come to Marah, but the waters are bitter~--- Moses makes an executive decision and brings them to Elim}

\begin{inparaenum}
    \verse{15:1} Moses and the children of Israel then sang this song to the \textsc{Lord}, and they spoke, saying,\smallskip%%
    
    \pvbd{``I will sing to the \textsc{Lord}}{because he is greatly exalted.}{The horse and chariot}{he's thrown into the sea.}%%
    
    \pvab{\vn{15:2} My strength and song is \textsc{Jah}\footnotemark~---}{He has become my\footnotemark\ salvation.}%%
    \fntca{\missing\ \septuagint}{[``\textsc{Jah}''] is missing in the Septuagint}%%
    \fntca{\peshitta\ \textit{ln}}{the Peshitta has ``our''}%%
    
    \pvbb{This is my God: I will praise Him~---}{my father's God: I will exalt\footnotemark\ Him.}%%
    \fntalt{praise, extol}%%
    
    \pvab{\vn{15:3} The \textsc{Lord} is a man\footnotemark\ of war.}{His name is \textsc{Jehovah}.}%%
    \fntca{\sampen\peshitta\ \Hebrew{גִּבּוֺר}, \septuagint\ \Greek{συντρίβων}}{the Samaritan Pentateuch and Peshitta have ``hero,'' the Septuagint has ``crusher''}%%
    
    \pvad{\vn{15:4} Pharaoh's chariots and force}{he's cast into the sea.}{His chosen captains}{he's sunk in the Red Sea.}%%
    
    \pvab{\vn{15:5} The depths cover him;}{they went into the depths like a stone.}%%
    
    \pvad{\vn{15:6} Your right hand, O \textsc{Lord},}{has become glorious in power:}{Your right hand, O \textsc{Lord},}{has destroyed the enemy.}%%
    
    \pvac{\vn{15:7} You've demolished your opponents\footnotemark\ in the abundance of Your excellency;}{You've sent Your anger}{and it consumes them like chaff.}%%
    \fntlit{one who stands against you}%%
    
    \pvac{\vn{15:8} By the breath of Your nostrils}{You've dammed up\footnotemark\ the water~---}{the flowing waves stand like a dam,\footnotemark}%%
    \fntalt{piled up, gathered together}%%
    \fntalt{dike, wall}%%
    
    \pvba{the depths in the heart of the sea have congealed.\footnotemark}%%
    \fntalt{thickened.}%%
    
    \pvbd{\vn{15:9} The enemy said,}{`I pursue. I overtake.\footnotemark}{I allot plunder.}{My soul is overwhelmed by them.}%%
    \fntalt{catch up.}
    
    \pvbb{I draw my sword,}{my hand destroys them.'}%%
    
    \pvad{\vn{15:10} You have blown\footnotemark\ with Your breath,}{the sea covers them;}{they sank like lead}{in the mighty waters.}%%
    \fntalt{blasted}%%
    
    \pvad{\vn{15:11} Who is like You}{among the gods, O \textsc{Lord}?}{Who is like You,}{glorious in holiness,}%%
    
    \pvbd{awe-inspiring in praises,}{doing miracles?\footnotemark}{\vn{15:12} You've stretched out Your right hand}{and the earth has swallowed them.}%%
    \fntalt{marvels, wonders, extraordinary things, something extraordinary?}%%
    
    \pvad{\vn{15:13} In Your kindness You've led}{the people whom You've redeemed:}{You've guided them\footnotemark\ in Your strength}{to Your holy abode.}%%
    \fntca{\sampen\super{Mss}\targum\super{J} \Hebrew{נִחַלְתָּ}}{the Samaritan Pentateuch and Targum have ``taken possession''}%%
    
    \pvab{\vn{15:14} The nations\footnotemark\ have heard, they've trembled:}{the Philistines have been seized with anguish.\footnotemark}%%
    \fnted{people\thinspace$\rightarrow$\thinspace peoples\thinspace$\rightarrow$\thinspace nations}%%
    \fntalt{pain}%%
    
    \pvad{\vn{15:15} Then have the chiefs of Edom}{been amazed\footnotemark}{and the mighty ones of Moab}{have been seized with trembling.}%%
    \fnted{The first two hemistichs have been rearranged; it's literally ``Then been amazed * the chiefs of Edom''}
    
    \pvba{All the inhabitants of Canaan have melted away.}%%
    
    \pvad{\vn{15:16} Terror and dread}{have fallen on them.\footnotemark}{By the greatness of Your arm}{they are as silent\footnotemark\ as stone.}%%
    \fnted{The first two hemistichs have been rearranged; it's literally ``Have fallen on them * terror and dread.''}%%
    \fntalt{still}%%
    
    \pvbb{Until Your people pass over, O \textsc{Lord};}{until the people You've bought have passed over.}%%
    
    \pvad{\vn{15:17} You shall bring them in and plant them}{in the mountain of Your inheritance,}{in the place You've made}{for You to dwell in, O \textsc{Lord}.}%%
    
    \pvbd{A sanctuary of my Lord}{has been established by Your hand.}{\vn{15:18} The \textsc{Lord} shall reign}{forever and always.}%%
    
    \verse{15:19} Because Pharaoh's horse came in with his chariots and his horsemen into the sea, and the \textsc{Lord} turned the water of the sea back on them; and the children of Israel went on dry ground in the midst of the sea.''%%
    
    \verse{15:20} Miriam, the prophetess, Aaron's sister, took the timbrel\alt{tambourine} in her hand, and all the women following her went out with their timbrels and dances,%%
    \verse{15:21} and Miriam answered them:\smallskip%%
    
    \pvbd{``Sing to the \textsc{Lord}}{because He is greatly exalted!}{The horse and its chariot}{He's thrown into the sea.''}%%
    
    \verse{15:22} Moses brought Israel from the Red Sea, and they went out into the Shur desert. They went three days in the desert and couldn't find water.%%
    \verse{15:23} They came to Marah and hadn't been able to drink the water of Marah because they were bitter;\halot{\textbf{bitter} (in taste): water \haref{Ex}{15}{23}; \textit{m\^e hamm\=ar\^im} bitter-water (meaning ?)} hence they called it Marah.%%
    \verse{15:24} The people murmured\alt{grumbled} against Moses, saying, ``What shall we drink?''%%
    \verse{15:25} He\ca{\sampen\septuagint\peshitta\vulgate\ + \Hebrew{מֹשֶׁה}}{the Samaritan Pentateuch, Septuagint, Peshitta, and Vulgate add ``Moses''} cried to the \textsc{Lord}, and the \textsc{Lord} showed him a tree, and he threw it into the waters and the waters became sweet.\ed{There's a peculiar tabbing here in the Hebrew that's preserved with the following space.}\hspace*{4em}He there made a law and a legal decision for them. He put them to the test.%%
    \verse{15:26} He said, ``If you'll actually listen to the voice of the \textsc{Lord} your God and do what's right in His eyes and hearken to His commandments and obey all of His laws, I won't put any of the sickness I put on the Egyptians on you because I, the \textsc{Lord}, am your healer.''%%
    \verse{15:27} They came to Elim and there were twelve water fountains and seventy date palms,\halot{\textit{Phoenix dactylifera}} and they camped there by the waters.%%
\end{inparaenum}
