\section{Exodus 19}\label{exodus:19}
\heading{Israel to be a peculiar treasure, a kingdom of priests, a holy nation~--- they are sanctified~--- the Lord appears in Sinai}
\begin{enumerate}[align=center]
    \verse{19:1} In the third month since the departure of the children of Israel from the land of Egypt, \footnote{in that day}they came into the Sinai wilderness.%%
    \verse{19:2} They left Rephidim, came into the Sinai wilderness, and camped\footnote{in the Sinai wilderness} before the mountain.%%
    \verse{19:3} And Moses went up to God, and the \textsc{Lord} called to him from the mountain, saying, ``Thus shall you say to the children of Jacob and tell to the sons of Israel:%%
    \verse{19:4} `You~--- you've seen what I've done to the Egyptians, how I bear you on eagles' wings, and how I have brought you to Myself.%%
    \verse{19:5} And now, if you will truly obey Me and keep\footnote{observe} My commandments, you shall be a treasure\footnote{to Me}\footnote{Refers to a temple treasure, therefore having covenantal implications.} from among all people:\footnote{above all people} for all the earth is Mine.%%
    \verse{19:6} You shall be\footnote{to Me} a kingdom of priests and a holy nation.' These are the words which you shall say to the children of Israel.''%%
    \verse{19:7} So Moses entered in\footnote{i.e., to the camp} and called the elders of the people. He laid before them all these things that the \textsc{Lord} had commanded him.%%%
    \verse{19:8} Then all the people answered together, saying, ``We will do everything the \textsc{Lord} has said.'' So Moses returned to tell the \textsc{Lord} what the people had said.\footnote{the words of the people}%%
    \verse{19:9} The \textsc{Lord} said to Moses, ``I will come to you in a thick cloud so that the people may hear my words when I speak with you. So shall they believe you forever.'' Then Moses told the words of the people unto the \textsc{Lord}.%%
    \verse{19:10} And the \textsc{Lord} said unto Moses, ``Go to the people and sanctify them today and tomorrow. Let them wash their clothes%%
    \verse{19:11} because on the third day the \textsc{Lord} will descend in the eyes of all the people before Mount Sinai.%%
    \verse{19:12} You shall set bounds around the people, saying, `Hearken to these words\footnote{lit., them} and don't go up into the mountain. Anyone touching the base of the mountain will certainly be put to death.%%
    \verse{19:13} No hand shall touch it\footnote{the mountain} or he\footnote{whomever touches it} will surely be stoned or shot:\footnote{with arrows} whether beast or man, they shall not live while the trumpet sounds\footnote{continues} as they approach the mountain.'\,''%% This needs work.
    \verse{19:14} So Moses went down from the mountain to the people and sanctified\footnote{consecrated, made holy} the people and they washed their clothes.\footnote{As a way of preparing themselves to go up into the mountain.}%%
    \verse{19:15} He said to the people, ``Prepare yourselves. Do not sleep with\footnote{lit., approach} a woman for three days.''%%
    \verse{19:16} On the third morning, when it was morning, there were noises and flashes of lightning, a thick cloud over the mountain, and the tremendously powerful sound of a trumpet:\footnote{One used for ritualistic purposes, a \Hebrew{שׁוֹפָר}.} and all the people in the camp were afraid.%%
    \verse{19:17} Moses brought the people out from the camp to meet God and they stationed themselves at the base of the mountain.%%
    \verse{19:18} All of Mount Sinai was smoking because the \textsc{Lord} descended on it in fire. And the smoke went up like the smoke of a furnace. The whole mountain shook.%%
    \verse{19:19} The sound of the trumpet grew continually louder\footnote{more and more powerful} while Moses spoke and God answered.\footnote{with a voice}%%
    \verse{19:20} The \textsc{Lord} descended on the summit of Mount Sinai and He\footnote{lit., the \textsc{Lord}} called unto Moses.\footnote{at the summit} And Moses went up.%%
    \verse{19:21} The \textsc{Lord} said to Moses, ``Go down and solemnly charge the people, otherwise many of them will break through, see the \textsc{Lord}, and perish.''\footnote{Not because of some great iniquity, merely because of the glory of the \textsc{Lord} and His inherent power.}%%
    \verse{19:22} Also, the priests who come near the \textsc{Lord} shall sanctify themselves, otherwise the \textsc{Lord} will tear them down.%%
    \verse{19:23} Moses said to the \textsc{Lord}, ``The people will not be able to come up to Mount Sinai because You have charged us, saying, `Make a border around the mount and sanctify it.'\,''%%
    \verse{19:24} The \textsc{Lord} said unto him, ``Go down~--- you and Aaron with you~--- and the priests will come up. Don't let the people break through to come up to the \textsc{Lord}, otherwise His glory will break out against them.''%%
    \verse{19:25} So Moses went down to the people and spoke to them.%%
\end{enumerate}
