\chap{21}

\heading{The Lord's law of servants, marriage, and the death penalty~--- eye for an eye, tooth for a tooth~--- damage caused by oxen}

\begin{enumerate*}[mode=unboxed]
    \verse{21:1} These are the judgments\footnote{laws} that you shall set before them.\footnote{Lit., ``their face''}%%
    \verse{21:2} If you acquire a Hebrew slave\footnote{male slave, servant. This person is subject to Jewish jurisprudence as evidenced by the use of the word \Hebrew{עבד}} six years, he shall serve thee; in the seventh year he shall go free without having to pay.\footnote{Like indentured servanthood.}%%
    \verse{21:3} If he comes alone, alone shall he go out. If he is married,\footnote{The husband/lord of a woman.} his wife shall go with him.%%
    \verse{21:4} If his master giveth him a wife and she bears sons or daughters to him, their\footnote{her} offspring\footnote{children} shall belong to the master and the man shall go forth by himself.%%
    \verse{21:5} If the servant actually says, ``I love my master, my wife, my children~--- I will not be freed.''%%
    \verse{21:6} Then his master will bring him\footnote{cause him to approach} to the presence of God, to the door, and his master shall pierce his ear with an awl and he shall be his\footnote{the master's} slave forever.%%
    \verse{21:7} If a man sells his daughter to be a handmaid, then she shall not go out as the bondsmen do.%%
    \verse{21:8} However, if she is unacceptable to her buyer\footnote{lit., master} (who has taken her for himself), then he shall let her be ransomed. However, because he has dealt with her unfairly, he shall not have power to sell her to foreigners.%%
    \verse{21:9} If he designates\footnote{assigns, but not necessarily \textit{marry}. \textit{Kohler-Baumgartner} simply says of \Hebrew{יִיעָדֶנָּה} (\Hebrew{יעד}) in the qal: designate, assign a woman} her to his son, then it shall be done to her according to the law of daughters.%%
    \verse{21:10} If he takes another for himself, then her food, clothing, and right to motherhood\footnote{alt., marital intercourse. Right to motherhood is more fitting contextually.}\footnote{How does this cover consent? Is that an issue in the ancient world?} shall not be taken away.%%
    \verse{21:11} But if he doesn't do these three things for her, she shall go free without money.%%
    \verse{21:12} Anyone who strikes\footnote{smite, (deal a) blow} a man to death shall certainly be killed.%%
    \verse{21:13} However, if he has not lain in wait, but God has delivered him\footnote{i.e., the killed} into his hand, then I will appoint a place where he can flee.%%
    \verse{21:14} If a man schemes\footnote{acts presumptuously} against his neighbor to kill him by deceit,\footnote{design, scheme, subtlety} you shall take him from my altar to kill him.%%
    \verse{21:15} He who strikes his father or his mother\footnote{\dots he\dots} shall surely be put to death.%%
    \verse{21:16} Anyone who kidnaps\footnote{steals} someone\footnote{to sell them} and he is caught with the man still in his possession:\footnote{power} he shall surely be put to death.\footnote{In other words, kidnapping is a capital offense.}%%
    \verse{21:17} He who curses\footnote{reviles, insults~--- but harsher than we tend to think} his father or his mother shall sure be put to death.%%
    \verse{21:18} And if men quarrel\footnote{fight} and one man hits another (whether with a stone or his fists), but not to kill him~--- merely to put him to bed~---%%
    \verse{21:19} if he rises and walks about outside\footnote{\textit{Koehler-Baumgartner} specifies this as the area outside a house.} on his staff,\footnote{For the sick.} then the smiter shall be declared exempt from punishment.\footnote{alt., acquitted of relevant charges} Only, he shall pay for his lost time until he is thoroughly healed.%%
    \verse{21:20} If a man strikes\footnote{hits} his male or female servant with a staff\footnote{rod} and that person dies under his hand then he will surely be punished in proxy for him.\footnote{or her.}%%
    \verse{21:21} But if a day or two goes by and he\footnote{i.e., the slave} gets up, he\footnote{the master} will not be punished for it is his\footnote{Understood: own.} property.%%
    \verse{21:22} If men fight and strike a pregnant woman so that she has a miscarriage,\footnote{Lit., ``her children go forth.''} but there is no further harm, he will surely be fined according to what the husband deems right, and he\footnote{the other man} will give according to what is assessed.%%
    \verse{21:23} If there is a serious injury or death, you shall take life in place of\footnote{for} life:%%
    \verse{21:24} an eye for an eye, a tooth for a tooth, a hand for a hand, a foot for a foot,\footnote{This is significantly more merciful than people let on. In other ancient civilizations (e.g., Hammurabi's code) there were different punishments depending on the person's social status.}%%
    \verse{21:25} burning for burning, wound for wound, strike for strike.%%
    \verse{21:26} If a man strikes the eye of his male or female servant and knocks it out,\footnote{Lit., destroys it} he will let him go free in penance for his eye.%%
    \verse{21:27} If the tooth of a male or female servant is knocked out, he must let them go free because of their tooth.%%
    \verse{21:28} If an ox\footnote{``Bull'' can be supplied for every following instance of ``ox'' in this chapter.} gores a man or a woman and that person dies, the ox will surely be stoned to death, his flesh shall not be eaten, and the owner shall not be liable.%%
    \verse{21:29} If the ox has been known to gore in the past and this was known\footnote{made known} to its master(s), and his master didn't keep him under guard, and it kills a man or a woman, the ox will be stoned and his master will be put to death as well.%%
    \verse{21:30} If a payment is laid upon him, he shall pay\footnote{give} a redemption\footnote{ransom} for his life according to everything that is laid upon him.%%
    \verse{21:31} Whether it runs down\footnote{alt., butts, thrusts, gores} a son or a daughter, then you shall do to it according to this judgment.%%
    \verse{21:32} If the ox runs down a male servant or a handmaid, then he shall pay the master thirty pieces of silver\footnote{Thirty shekels} and the ox shall be stoned.%%
    \verse{21:33} If a man uncovers\footnote{opens} or digs a pit, doesn't properly cover it,\footnote{Probably meant ``at night.''} and a donkey or an ox falls in,%%
    \verse{21:34} then the owner of the pit shall make it right: he shall give money to the owner\footnote{of the animal} and the dead animal shall be his.%%
    \verse{21:35} If a man's bull strikes the bull of his neighbor\footnote{associate} and he dies, then he shall sell the live ox and half the money and the dead ox.%%
    \verse{21:36} If it is known that the bull has previously run people down\footnote{lit., is known to run people down} and its owner has not kept watch of him, then the owner will have to compensate the owner of the dead ox and the dead ox will belong to him.\footnote{i.e., the one who paid.}%%
    \verse{21:37} If a man should steal an ox or a sheep\footnote{lamb} and slaughters\footnote{use it for food} or sells it, he shall recompense\footnote{make peace} with five oxen for the ox, or four sheep for the lamb.%%
\end{enumerate*}
