\heading{24}{Israel, by covenant, accept the Lord~--- Moses, Aaron, Nadab, Abihu, and the Seventy see God~--- the Lord calls Moses to the mountain to receive the stone tablets}

\begin{inparaenum}
    \verse{24:1} And He saith unto Moses: ``Go up unto the \textsc{Lord}~--- you, Aaron, and seventy of the elders of Israel~--- and you shall bow down\footnote{i.e., to worship} from afar.%%
    \verse{24:2} And Moses will approach the \textsc{Lord} alone. The people\footnote{i.e., Aaron and the Seventy} will not go up with him.''%%
    \verse{24:3} And Moses came and told the people all the words and judgments of the \textsc{Lord}; and all of the people answered with one voice, saying, ``All the words that the \textsc{Lord} hath spoken we will do.''%%
    \verse{24:4} Moses wrote all the words of the \textsc{Lord}. Then, he rose early in the morning, and built an altar at the foot of the mountain. He also made\footnote{built} twelve memorial stones\footnote{pillars} for the twelve tribes of Israel.%%
    \verse{24:5} And he sent lads\footnote{young men, servants} of the children of Israel and they offered burnt offerings and sacrificed sacrifices, peace offerings, and bulls.\footnote{oxen}%%
    \verse{24:6} Then Moses took half of the blood and placed it in bowls;\footnote{basins} and half\footnote{i.e., the other half} of the blood he sprinkled\footnote{cast} on\footnote{at} the altar.%%
    \verse{24:7} He took the scroll of the covenant and he read\footnote{it} out in the ears\footnote{hearing} of the people and they said, ``Everything that the \textsc{Lord} hath said we will do.''\footnote{hear, obey}%%
    \verse{24:8} And Moses took the blood and he sprinkled\footnote{cast} it upon the people and said, ``This is the blood of the covenant which the \textsc{Lord} hath made with you concerning all these things.''%%
    \verse{24:9} Then Moses, Aaron, Nadab, Abihu, and seventy elders of Israel went up%%
    \verse{24:10} and saw the God of Israel. Under His feet was a slab of sapphire as bright as heaven.%%
    \verse{24:11} He did not lay His hand on the eminent children of Israel. They saw God and they ate and they drank.\footnote{xxxx: reference to the sacrament?}%%
    \verse{24:12} The \textsc{Lord} said to Moses, ``Come up to me in the mountain\footnote{lit., mountain-ward} and be there. I will give you stone tablets, the Law and the Commandments that I have written for their instruction.''%%
    \verse{24:13} Moses rose up~--- and Joshua, his minister, too~--- and Moses went to the mountain of God.%%
    \verse{24:14} To the elders he said, ``Wait here\footnote{lit., in this place} for us until we return for you. And look, Aaron and Hur are here\footnote{Understood} with you, and whoever has any problems, bring it up with them.''\footnote{lit., approach them with it}%%
    \verse{24:15} Moses went up into the mountain, and the cloud covered the mountain.%%
    \verse{24:16} The glory of the \textsc{Lord} settled on Mount Sinai, and the cloud covered the mountain for six days. On the seventh day, He called to Moses from the thick\footnote{lit., midst} of the cloud.%%
    \verse{24:17} In the people's view, the appearance of the \textsc{Lord} in His glory was like fire consuming the peak of the mountain.%%
    \verse{24:18} Moses came into the thick of the cloud and climbed up the mountain. And Moses was in the mountain for forty days and forty nights.%%
\end{inparaenum}
