\heading{1}{The Lord reproves the people for having not rebuilt the temple, but rather their own homes~--- they obey the Lord and work on the temple}

\begin{inparaenum}
  \verse{1:1} In the second year of King Darius,\ed{ca.\ 522~\textsc{b.c.}} in the sixth month, on the first day of the month, the word of the \textsc{Lord} came by the hand of the prophet Haggai to Zerubbabel son of Shealtiel, governor of Judah, and to Joshua son of Jehozadak, the high priest, saying:%%
  \verse{1:2} ``Thus says the \textsc{Lord} of Hosts, saying: `This people has said, ``The time has not come, the time that the house of the \textsc{Lord} will be built!''\thinspace'\thinspace''%%
  
  \verse{1:3} The word of the \textsc{Lord} through\lit{by the hand of} the prophet Haggai, saying:\smallskip%%
  
  \pvcb{\vn{1:4} ``Is it time for you~--- you~---}{to live in your paneled houses\footnotemark}%%
  \fnted{probably one that is nicely finished on the outside, but probably on the inside as well}%%
  
  \pvda{while this house lies in ruin?''}%%
  
  \pvca{\vn{1:5} And so now, thus says the \textsc{Lord} of Hosts:}%%
  
  \pvda{``Take heed:\footnotemark}%%
  \fnted{For variant renderings, see Appendix~\ref{app:take-heed}.}%%
  
  \pvcb{\vn{1:6} you've sewn much, but brought in little.}{You eat, but you're not satiated.}%%
  
  \pvdb{You drink, but don't get drunk.\footnotemark}{You get dressed, but you're not warm.}%%
  \fnted{This seems to be the Lord's equivalent of ``go big or go home.'' The people aren't doing good~--- they're not even doing bad well. It seems to be that the Lord just wants them to commit: if you're going to be good, be good! If you're going to be bad, be bad! There is no efficacy in anything they do.}%%
  
  \pvdb{He who hires himself out for money, hires himself out}{for a bag that has holes in it.''}\bigskip%%
  
  \verse{1:7} Thus says the \textsc{Lord} of Hosts:\smallskip%%
  
  \pvda{Take heed.\footnotemark}%%
  \fntca{add? cf 5}{[Was this stich] added? Compare verse~5}%%
  
  \pvcc{\vn{1:8} Go up to the mountain}{and bring wood\footnotemark}{and build a house}%%
  \fntlit{a tree}%%
  
  \pvdc{that I may be pleased with it}{and be glorified,''}{says the \textsc{Lord}.}%%
  
  \pvcb{\vn{1:9} ``Turning about\footnotemark\ much, but it came to little;}{you brought it home and blew on it.}%%
  \fntalt{``Paying attention to, Worrying about}%%
  
  \pvdb{Why?''}{declares the \textsc{Lord} of Hosts.}%%
  
  \pvdb{``Because My house is in ruins}{and every man is running to his house.}%%
  
  \pvcb{\vn{1:10} Therefore, the heavens have withheld their dew\footnotemark\ from over you}{and the land has withheld its produce.}%%
  \fntalt{light rain}%%
  
  \pvcb{\vn{1:11} I proclaimed\footnotemark\ a drought on the land,}{the mountains, the grain,}%%
  \fntalt{called for}%%
  
  \pvdb{the new wine,\footnotemark\ the pure oil,\footnotemark}{and on\footnotemark\ that which the ground brings forth~---}%%
  \fnted{probably freshly squeezed; not fermented or only recently starting to ferment}%%
  \fntca{add?}{[``the oil''] added?}%%
  \fntca{ins c mlt Mss Vrs \Hebrew{־כָּל־}}{multiple Hebrew manuscripts and all or most translations insert ``all''}%%
  
  \pvdb{on mankind, on beasts,}{and on all the exertions\footnotemark\ of the hands.''}\bigskip%%
  \fntalt{labors, toils}%%
  
  \verse{1:12} Zerubbabel son of Shealtiel and the high priest, Joshua son of Jehozadak, and all the remnant of the people heeded the voice of the \textsc{Lord} their God and the words of the prophet Haggai as the \textsc{Lord} their God had sent them, and the people feared before the \textsc{Lord}.%%
  \verse{1:13} Haggai, the messenger of the \textsc{Lord}, in the \textsc{Lord}'s messages to the people, spoke, saying,\hspace*{2em} ``\thinspace`I am with you,' declares the \textsc{Lord}.''%%
  \verse{1:14} The \textsc{Lord} stirred\alt{raised} up the spirit of the governor of Judah, Zerubbabel son of Shealtiel, and the spirit of the high priest, Joshua son of Jehozadak, and the spirit of all the remnant of the people: they came in\ed{This preposition could easily be rendered ``on'' which would greatly change the meaning of this sentence.} and did work in the house of the \textsc{Lord} of Hosts, their God.%%
  
  \pvca{\vn{1:15} In the twenty-fourth day of the sixth month in the second year of King Darius.\footnotemark}%%
  \fntca{huc ins \caref{}{2}{15--19}}{hither insert \vref{Hag}{2}{15--19}}%%
\end{inparaenum}
