\heading{2}{The Lord encourages Zerubbabel, Joshua, and the people~--- He asks Haggai to question the priests concerning the Law~--- the Lord chooses Zerubbabel}

\begin{inparaenum}
    \noindent\verse{2:1} \ed{This verse does not start a new paragraph, rather it is a continuation of the previous chapter.}In the seventh month,\understood\ on the twenty-first day of the month, the word of the \textsc{Lord} came by the hand of Haggai the prophet, saying:%%
    \verse{2:2} ``Please speak to the governor of Judah, Zerubbabel son of Shealtiel, and the high priest, Joshua son of Jehozadak, and the remnant of the people, saying:\smallskip%%
    
    \pvca{\vn{2:3} `Who among you is left}%%
    
    \pvdb{who saw this house}{in its former glory?}%%
    
    \pvdb{And how do you see}{it now?\footnotemark}%%
    \fntalt{How do you look at it now?}%%
    
    \pvdb{Is it like nothing}{in your eyes?}%%
    
    \pvca{\vn{2:4} Now, be strong, Zerubbabel,' declares the \textsc{Lord}.\footnotemark}%%
    \fntca{add?}{[``declares the \textsc{Lord}''] added?}%%
    
    \pvdb{`Be strong, Joshua son of Jehozadak,}{the high priest;}%%
    
    \pvdb{be strong, all the people of the land,'}{declares the \textsc{Lord}.\footnotemark}%%
    \fntca{add?}{[``Joshua\dots\ \textsc{Lord}''] added?}%%
    
    \pvdb{`Work, because I am with you,'}{declares the \textsc{Lord} of Hosts.\footnotemark}%%
    \fntca{add?}{[``of Hosts''] added?}%%
    
    \pvca{\vn{2:5} \footnotemark The thing\footnotemark\ I covenanted with you when you came out of Egypt}%%
    \fntca{prb l \Hebrew{זֹאת} \Hebrew{הַבְּרִית}, al dl \Hebrew{את} = \Hebrew{(צב)אות}}{probably read ``this covenant,'' others delete \textit{\d tbaot} and leave \textit{et}}%% xxxx fix the latinization here
    \fntca{\missing\ \septuagint*, add}{[this stich] missing in the Septuagint (textus Graecus originalis) and was added}%%
    
    \pvdb{and My spirit remains in your midst~---}{don't be afraid.'\thinspace''}\bigskip%%
    
    \pvcb{\vn{2:6} For thus says the \textsc{Lord} of Hosts:}{``In just a little bit\footnotemark}%%
    \fntlit{``Yet once, it's little,}%%
    
    \pvdb{I will shake the heavens and the earth,}{the sea and the dry land.}%%
    
    \pvcb{\vn{2:7} I will shake all the nations,}{and the desire of the nations shall come in.}%%
    
    \pvdb{I will fill this house with glory,''}{says the \textsc{Lord} of Hosts.}%%
    
    \pvcb{\vn{2:8} ``The silver is Mine, the gold is Mine,''}{declares the \textsc{Lord} of Hosts.}%%
    
    \pvcc{\vn{2:9} ``The glory of this last\footnotemark\ house is greater}{than the first,''}{says the \textsc{Lord} of Hosts.}%%
    \fnted{moved ``last'' from the second hemistich to here for idiomaticy}%%
    
    \pvdb{``I will create\footnotemark\ peace in this place,''}{declares the \textsc{Lord} of Hosts.}%%
    \fntlit{give}%%
    
    \verse{2:10} On the twenty-fourth day\understood\ of the ninth month\understood\ in the second year of Darius, the word of the \textsc{Lord} came by the hand of Haggai the prophet, saying:%%
    \verse{2:11} ``Thus says the \textsc{Lord} of Hosts: `Please inquire\ed{This is direction being given to Haggai, not to the people through Haggai.} of the priests concerning\understood\ the Law, saying:%%
    \verse{2:12} ``If someone is carrying holy flesh in the fold\lit{skirt} of his garment and he touches bread with his skirt~--- or a boiled dish or wine or oil or any food~--- shall it become holy?'' And the priest will answer and say, ``No.''%%
    \verse{2:13} And Haggai will say, ``If someone who is cultically unclean shall touch any of these, will they become unclean?'' And the priest will answer and say, ``It shall be unclean.''%%
    \verse{2:14} And Haggai will answer and say,\smallskip%%
    
    \pvdc{``So is this people}{and so is this nation}{before Me,'' declares the \textsc{Lord}.}%%
    
    \pvdc{``So is everything they do with their hands}{and that which they offer there:}{it is unclean.}%%
    
    \pvaa{\vn{2:15} And now,\footnotemark}%%
    \fntca{tr 15--19 post 1,15a}{transpose verses~15--19 after \vref{Hag}{1}{15}}%%
    
    \pvab{please take heed}{from this day on}%%
    
    \pvab{before a stone was laid on a stone}{in the house of the \textsc{Lord}. \vn{2:16} When\footnotemark\footnotemark}%%
    \fntlit{From their being}%%
    \fntalt{From that time}%%
    
    \pvab{they come to a sheath of twenty,}{there are only ten;\footnotemark}%%
    \fnted{Meaning that when they try to produce something, they will only produce half; when they try to sanctify things, they will only be unclean.}
    
    \pvab{when they come to the wine vat to get fifty,}{there will only be twenty.}%%
    
    \pvab{\vn{2:17} I have smitten you with blight and mildew}{and hail, all the work of your hands.}%%
    
    \pvab{None of you are with Me,''}{declares the \textsc{Lord}.}%%
    
    \pvab{\vn{2:18} ``Please take heed}{from today onward}%%
    
    \pvaa{from the twenty-fourth day of the ninth month, from the day the temple of the \textsc{Lord} was founded: take heed.}%%
    
    \pvac{\vn{2:19} Is the seed\footnotemark\ still in the grain-pit?\footnotemark\footnotemark}{Until the vine, the fig, and the pomegranate,\footnotemark}{and the olive tree do not produce:\footnotemark\footnotemark}%%
    \fntca{prb l \Hebrew{םִגְרָע} deminutio, al ins \Hebrew{נִגְרַע}}{probably read ``diminished'' [probably ``Is the grain-pit yet diminished?''], while others insert ``diminish'' [i.e., ``Is the seed still diminished in the grain-pit?'']}%%
    \fntalt{storage chamber?}%%
    \fnted{This is a classic example of a non-idiomatic translation used in vogue: many English translations render this as ``barn'' even though the concept of a barn is anachronistic.}%%
    \fntca{add?}{[Was this hemistich] added?}%%
    \fntlit{bear}%%
    \fntalt{bring forth}%%
    
    \pvaa{from this day I will bless you.''}%%
    
    \pvba{\vn{2:20} The word of the \textsc{Lord} came to Haggai on the twenty-fourth day of the month,}%%
    
    \pvaa{saying: \vn{2:21} ``Speak to the governor of Judah, Zerubbabel, saying:}%%
    
    \pvcb{`I am shaking}{the heavens and the earth.w}%%
    
    \pvbb{\vn{2:22} I will overthrow the throne of the kingdoms;}{I will exterminate the strength of the kingdoms\footnotemark\ of the nations.}%%
    \fntca{prb dl, var lect}{[``of the kingdoms''] is probably deleted, [or is a] variant reading}%%
    
    \pvcc{I will overthrow the chariots and their riders:}{their horses and riders shall come down,}{each by his brother's sword.\footnotemark}%%
    \fntca{add?}{[this hemistich] added?}%%
    
    \pvcb{\vn{2:23} In that day,' declares the \textsc{Lord} of Hosts, `I will take you,}{Zerubbabel, son of Shealtiel, My servant,' the \textsc{Lord} declares,\footnotemark}%%
    \fntca{add?}{[``son of Shealtiel'' and ``the \textsc{Lord} declares''] added?}%%
    
    \pvcb{`and I will make you like a seal\footnotemark}{because I have chosen you,'}%%
    \fnthalot{xxxx}{for identification}%%
    
    \pvda{declares the \textsc{Lord} of Hosts.''}%%
\end{inparaenum}
