\heading{10}{Trumpets to be used in calling and assembling Israel~--- cloud removes from camp~--- Israel sets forth in their orders~--- the Ark of the Covenant goes before the people}

\begin{enumerate*}[mode=unboxed]
    \verse{10:1} The \textsc{Lord} spoke to Moses, saying,%%
    \verse{10:2} ``Make two silver trumpets\footnote{KB: long, straight instrument of metal for signaling.}\footnote{Not to be confused with a \textit{shofar}.} for yourselves. You shall make them of hammered\footnote{or embossed} metalwork. They shall serve you when the assembly is gathered and when the order of departure for the camp is given.%%
    \verse{10:3} When they blow them, the whole company shall gather to you at the entrance of the meeting tent.%%
    \verse{10:4} If they blow with one, then the princes, the heads of thousands, shall gather to you.%%
    \verse{10:5} When you've blown an alarm, the camps which are encamped eastward shall break camp and march on.%%
    \verse{10:6} When you blow the alarm the second time, the camps which are encamped southward shall set forward. They shall blow an alarm when they set forward.%%
    \verse{10:7} In the assembling of the congregation of people, you shall blow the horn, but you shall not shout.%%
    \verse{10:8} The descendants of Aaron the High Priest shall blow the trumpets. They shall be an eternal, diving statute for you.%%
    \verse{10:9} If you go to war in your land against the enemy who is being hostile toward\footnote{alt., in a state of conflict with} you, then you shall blow the trumpets and you shall be remembered by\footnote{alt., in the sight of} the \textsc{Lord} your God and be saved from your enemies.%%
    \verse{10:10} In the day that you rejoice, in your assemblies and in your new moons, you shall blow the trumpets because of and in addition to\footnote{\Hebrew{עַל}, here, carries with it both of these meanings.} your burnt offerings and your communional peace sacrifices. These\footnote{lit., they} shall be a reminder for you \footnote{xxxx: Check KB on \Hebrew{לִפְנֵי}.}of your God: I, the \textsc{Lord}, am your God.%%
    \verse{10:11} And in the second year, in the second month, on the twentieth of the month, that the cloud was taken up from the dwelling place of the reminders.\footnote{See Appendix~\ref{}}%%
    \verse{10:12} The children of Israel set forth on their journey from the wilderness of Sinai and the cloud rested in the wilderness of Paran.%%
    \verse{10:13} And the first went not before the face of the \textsc{Lord} by the hand of Moses.%%
    \verse{10:14} The banner of the camp of Judah set forth first off, according to their ranks; and over its host was Nahshon the son of Amminadab.%%
    \verse{10:15} Over the host of the tribe of the children of Issachar: Nethaneel the son of Zuar.%%
    \verse{10:16} Over the host of the tribe of the children of Zebulun: Eliab the son of Helon.%%
    \verse{10:17} The utensils of the tabernacle were taken down, and the sons of Gershon and the sons of Merari~--- who carried the tabernacle and its accouterments~--- set forth.%%
    \verse{10:18} The banner of the camp of Reuben traveled according to their ranks; and over its host was Elizur the son of Shedeur.%%
    \verse{10:19} Over the host of the tribe of the children of Simeon: Shelumiel the son of Zurishaddai.%%
    \verse{10:20} Over the host of the tribe of the children of Gad: Eliasaph the son of Deuel.%%
    \verse{10:21} The carriers of the Holy Place\footnote{Inferred: the Holy of Holies as well.}~--- the Kohathites~--- went forth; and they would be the ones to set up the tabernacle at their\footnote{i.e., the Israelites} coming to a particular spot.%%
    \verse{10:22} The banner of the camp of the children of Ephraim traveled according to their ranks; and over its host was Elishama the son of Ammihud.%%
    \verse{10:23} Over the host of the tribe of the children of Manasseh: Gamaliel the son of Pedahzur.%%
    \verse{10:24} Over the host of the tribe of the children of Benjamin: Abidan the son of Gideoni.%%
    \verse{10:25} The banner of the camp of the children of Dan set forth, which was the rear of all the camps throughout their armies: and over its rank was Ahiezer the son of Ammishaddai.%%
    \verse{10:26} Over the host of the tribe of the children of Asher: Pagiel the son of Ocran.%%
    \verse{10:27} Over the host of the tribe of the children of Naphtali: Ahira the son of Enan.%%
    \verse{10:28} Such were the journeys of the children of Israel according to their ranks when they set forth.''%%
    \verse{10:29} And Moses said to Hobab,\footnote{Jethro} the son of Reuel the Midianite, the father-in-law of Moses, ``We are setting off to the same place that the \textsc{Lord} said, `I will give to you.' Come with us and we will treat you kindly because the \textsc{Lord} hath dealt kindly with Israel.''%%
    \verse{10:30} But he said, ``I will not go; rather, I will go unto mine own land and to mine own home country.''%%
    \verse{10:31} And he\footnote{Moses} said, ``Please don't leave me because you know how we should encamp in the desert~--- you shall be our eyes.%%
    \verse{10:32} ``If you will go with us, it will be good to you in the same measure as the \textsc{Lord} will be\footnote{was. Some slight eisegesis here, but I feel the rendering is more accurate with an eternal perspective. However, it works perfectly well with ``was'' because the Israelites have already been delivered from Egypt.} with us.''%%
    \verse{10:33} So they set out from the mountain of the \textsc{Lord} a journey of three days, and the Ark of the Covenant of the \textsc{Lord} went before them three days to seek out a place of rest for them.%%
    \verse{10:34} And the \textsc{Lord}, as a cloud, went over them by day when they journeyed from the camp.%%
    \verse{10:35} And when the Ark went forth, Moses said, ``Rise up, O \textsc{Lord}, and let thine enemies be scattered; yea, let thine haters flee from before thy face.''%%
    \verse{10:36} And when it rested he said, ``Return, O \textsc{Lord}, to the tens of thousands of Israelites.''\footnote{10,000 is the largest root describing numbers in Biblical Hebrew.}%%
\end{enumerate*}
