\heading{12}{Miriam and Aaron complain against Moses~--- Miriam is cursed with leprosy~--- Moses prays and Miriam is healed}

\begin{inparaenum}
    \verse{12:1} Miriam and Aaron spoke against Moses because of the Kushite\ed{possibly ``Midianite''} woman whom he married; for he had taken a Kushite woman in marriage.%%
    \verse{12:2} They said, ``Has the \textsc{Lord} only spoken through Moses? Hasn't He also spoken through us?'' And the \textsc{Lord} heard it.%%
    \verse{12:3} Now, the man (Moses) was incredibly unassuming\alt{meek, humble}~--- more than any other man on the face of the earth.\ed{In other words, he is too humble to fight his own battles. Instead, he lets the Lord fight for him.}%%
    
    \verse{12:4} The \textsc{Lord} spoke suddenly unto Moses, Aaron, and Miriam, ``The three of you shall go forth unto the tent of meeting.'' So the three of them went forth.%%
    \verse{12:5} Then the \textsc{Lord} went down in the pillar of a cloud and stood in the door of the tent. He called for Aaron and Miriam and they both came out.%%
    \verse{12:6} He said,%%
    
    \pvbb{``Heed my words:}{if you have a prophet, the \textsc{Lord}}%%
    
    \pvbb{will reveal Himself to him in a vision.}{I will speak unto him in a dream.}%%
    
    \pvab{\vn{12:7} Is it not true that Moses is My servant?}{That of all My house he is faithful?}%%
    
    \pvab{\vn{12:8} I speak to him face to face: visibly. Not in riddles.}{He has seen the form of the \textsc{Lord}.}%%
    
    \pvbb{Why weren't you hesitant}{to speak against my servant Moses?''}%%
    
    {\noindent\verse{12:9} The \textsc{Lord} grew angry with them and He left.}%%
    \verse{12:10} The cloud turned away from the tent\ie{the Tabernacle; see Appendix~\ref{app:tabernacle}} and Miriam was leprous\ed{What is mentioned here as ``leprosy'' is a curable skin condition. See Appendix~\ref{app:psalm-110}.} as snow. Aaron turned to face Miriam and she was leprous.%%
    \verse{12:11} Moses said, ``My \textsc{Lord}, please do not lay this sin that we foolishly sinned upon us.%%
    \verse{12:12} Please, don't let her look like\lit{be like} a stillborn, half of whose skin is eaten when it comes out of its mother's womb.''\ie{a malformed stillborn.}%%
    \verse{12:13} Moses cried unto the \textsc{Lord}, saying, ``Oh God, please heal her!''\ed{\Hebrew{נָא} is used twice to emphasize Moses' plea.}%%
    
    \verse{12:14} So the \textsc{Lord} said to Moses, ``If her father had spat in her face, would she not be in disgrace for seven days? Let her be removed from the rest of the people and after that\understood\ she shall be received.''\ed{We don't know why nothing happened to Aaron. Perhaps it's because he merely gave in to peer pressure (as he was wont to do).}%%
    \verse{12:15} Miriam was confined outside the camp for seven days, and the people did not travel until she was received back into the camp.%%
    \verse{12:16} Then the people up and left Hazeroth and camped in the desert of Paran.%%
\end{inparaenum}
