\heading{14}{The people of Israel complain to Moses and Aaron, wishing to have died in Egypt~--- Joshua speaks up in favor of the land of Canaan~--- the Lord and Moses discuss the future of Israel~--- the men who gave a false report are reproved by Moses}

\begin{inparaenum}
    \verse{14:1} All the congregation raised their voices and the people cried on that night.%%
    \verse{14:2} All the children of Israel murmured against Moses and Aaron. The entire congregation said to them, ``Would that we had died in the land of Egypt. Or would that we had died in this desert.%%
    \verse{14:3} Why did the \textsc{Lord} bring us to this land? To fall by the sword? That our wives and little ones become prey? Wouldn't it be better for us to return to Egypt?''%%
    \verse{14:4} Everyone said to their brother, ``We will choose a leader and return to Egypt.''%%
    \verse{14:5} So Moses and Aaron fell on their faces before the entire congregation of the people of Israel.%%
    \verse{14:6} Joshua, the son of Nun, and Caleb, the son of Jephunneh, from among the explorers of the land, rent their clothes.%%
    \verse{14:7} They said to the whole congregation of the people of Israel, saying, ``The land that we have explored is really good.%%
    \verse{14:8} If the \textsc{Lord} is pleased with us and brings us to this land, then He will give it to us: the land that is flowing with milk and honey.%%
    \verse{14:9} Only, don't rebel against the \textsc{Lord}. Do not fear the people of the land for they shall be our food. Their shadow is turned away and the \textsc{Lord} is with us. Don't be afraid of them.''%%
    \verse{14:10} All the congregation said to stone them\footnote{Joshua and Caleb, possibly Moses and Aaron, too.} with stones. But the glory of the \textsc{Lord} appeared over the tabernacle\footnote{Probably over the Holy of Holies.} in front of all of the people of Israel.%%
    \verse{14:11} The \textsc{Lord} said to Moses, ``For how long will this people despise Me? How long will they not believe Me despite all the signs that I have performed in their midst?%%
    \verse{14:12} I will smite them with a pestilence\footnote{alt., plague} and I shall disinherit them. I shall make you\footnote{Just you.} a great nation, mightier and more powerful than them.''%%
    \verse{14:13} But Moses said to the \textsc{Lord}, ``The Egyptians shall hear how You have, by Your power, brought up this people out of their midst%%
    \verse{14:14} and they will tell the other inhabitants of the land how they have heard that you, \textsc{Lord}, are in this people's midst, that you have appeared eye to eye, that your cloud stands above them. That you walk before them, a cloud by day, a pillar of fire by night.%%
    \verse{14:15} If you kill this people together,\footnote{lit., as one man} then the nations will say that which they've heard of You, saying,%%
    \verse{14:16} `The \textsc{Lord} was not able to bring this people to the land that He swore to them because He has destroyed them in the desert.'%%
    \verse{14:17} And now, I plead with You, let the power of the \textsc{Lord} be great according as You have previously spoken, saying,%%
    \verse{14:18} `The \textsc{Lord} is full of patience, incredibly merciful, forgiving wrongdoing and sin, not clearing those who are guilty, visiting the wrongdoings unto the third and fourth generation.'%%
    \verse{14:19} Please, pardon the transgression of this people according to Your great kindness, even as You have borne this people from Egypt until now.''%%
    \verse{14:20} The \textsc{Lord} said, ``I have forgiven according to your word.%%
    \verse{14:21} And yet, even as I live, the whole earth will be filled with the glory of the \textsc{Lord}.%%
    \verse{14:22} All mankind\footnote{Referring simply to the Israelites} who see My glory and signs that I have done in Egypt and the desert, those who have tested Me these ten times and have not hearkened to My voice:%%
    \verse{14:23} surely they shall not see the land that I have sworn to their fathers. Everyone who has provoked me shall not see it.%%
    \verse{14:24} My servant Caleb, because there was another spirit in him and because he fully follows\footnote{is fully after} Me: I shall bring him into the land that he has entered. His posterity shall possess it,%%
    \verse{14:25} and the Amalekites and the Canaanites who are dwelling in the valley. On the morrow, turn yourselves unto the desert by the way of the Red Sea.''%%
    \verse{14:26} The \textsc{Lord} spoke unto Moses and Aaron, saying,%%
    \verse{14:27} ``How long do I have to bear this wicked congregation that murmur against me? I have heard the murmurings of the people of Israel which they whine against me.%%
    \verse{14:28} Say to them: `As I, the \textsc{Lord} live, as you have spoken in My ears, I will not do to you.%%
    \verse{14:29} Your corpses shall fall in this desert. All of your numbered ones, even all those numbered from a son twenty and up who has murmured against Me.%%
    \verse{14:30} You will not come into the land that I have raised My hand\footnote{Raised His arm to the square and sworn to them.}, which I have caused you to settle in with the exception of Caleb, the son of Jephunneh, and Joshua, the son of Nun.%%
    \verse{14:31} Your small children~--- who have become the objects of plunder, you said~--- I have brought them in. They know the land that you hate.%%
    \verse{14:32} Your corpses will fall in this desert.%%
    \verse{14:33} Your children shall be shepherds in this desert. I will bear your fornications until your corpses are consumed in the desert.%%
    \verse{14:34} By the number of the days in which you spied out the land (forty days), a day for a year\footnote{lit., a day for a year (again)} shall you bear your iniquities~--- forty years~--- and you shall know My displeasure.%% xxxx: This needs some serious work.
    \verse{14:35} I, the \textsc{Lord}, have spoken. Surely I shall do this thing to this whole, evil company XXXX'\,''%%
    \verse{14:36} The men that Moses sent to explore the land, they returned and caused the entire congregation to murmur against Moses by bringing an evil report of the land.%% xxxx: better word than evil? Change in next verse as well.
    \verse{14:37} Even those men who brought an evil report of the land died by the plague before the \textsc{Lord}.%%
    \verse{14:38} Joshua, the son of Nun, and Caleb, the son of Jephunneh, survived all the men who went out to explore the land.%%
    \verse{14:39} Moses spake these words to all the people of Israel and the people mourned exceedingly.%%
    \verse{14:40} They rose early in the morning and went up to the top of the mountain, saying ``Here we are. Because we have sinned, we have come up to the place that the \textsc{Lord} told us.''%%
    \verse{14:41} Moses said, ``Why is this? All of you are transgressing the \textsc{Lord}'s commands. It shall not succeed.\footnote{For two reasons: they had sinned and they would be consumed by the glory of the Lord.}%%
    \verse{14:42} Don't go up! The \textsc{Lord} is not in your midst and all of you are not smitten before your enemies.%% XXXX: This verse needs complete retranslation.
    \verse{14:43} The Amalekites and the Canaanites are before you there and you will fall on the sword because you have turned from following the \textsc{Lord} and therefore\footnote{Understood} the \textsc{Lord} is not with you.%%
    \verse{14:44} They have the audacity to go up to the base of the mountain. However, the Ark of the Covenant of the \textsc{Lord} and Moses did not leave the midst of the camp.%%
    \verse{14:45} The Amalekites and the Canaanites who are dwelling in the mountain shall come down and smite you and beat you even as far as Hormah.''%%
\end{inparaenum}
