\heading{27}{Zelophehad's daughters risk losing their inheritance because their father had no sons~--- they ask Moses about the situation, he inquires of the Lord, and an answer is given~--- Joshua is ordained}

\begin{inparaenum}
  \verse{27:1} The daughters of Zelophehad (the son of Hepher, the son of Gilead, the son of Machir, the son of Manasseh, from the families of Manasseh the son of Joseph) came\alt{drew} near, and these are the names of his daughters: Mahlah, Noah, Hoglah, Milcah, Tirzah.%%
  \verse{27:2} They stood before Moses and Eleazar the priest and the military\halot{xxxx}{xxxx} leaders and the entire congregation at the door of the Tabernacle,\lit{opening of the tent of meeting} saying,%%
  \verse{27:3} ``Out father died in the wilderness~--- he was not in the midst of the company of Korah who gathered together against the \textsc{Lord}, but from his own sins. And he had no sons.%%
  \verse{27:4} Why should the name of our father be removed from his family? Because he had no sons? Please, give property to us in the midst of our uncles.''\lit{father's brothers.}%%
  \verse{27:5} Moses brought their matter before the \textsc{Lord}.%%
  
  \verse{27:6} The \textsc{Lord} spoke to Moses, saying,%%
  \verse{27:7} ``The daughters of Zelophehad have rightly spoken. You shall surely give to them an inheritance among their uncles by giving their father's inheritance to them.%%
  \verse{27:8} %%
  \verse{27:9} If he doesn't have a daughter, his inheritance shall go to his brethren.\alt{[nearest?] male relative.}%%
  \verse{27:10} If he has no brothers, you shall give it to his uncles.%%
  \verse{27:11} If his father has no brothers, you shall give his inheritance to his next of kin and he shall possess it.'' Thus it was a statue of judgment to the children of Israel, as the \textsc{Lord} had commanded Moses.%%
  
  \verse{27:12} The \textsc{Lord} spoke\ca{\sampen\ ut 6\super{a} $\rightarrow$ 2 Mss \sampen\septuagint\ \Hebrew{וַיְדַבֵּר}}{the Samaritan Pentateuch as verse~6a, which says ``two Hebrew manuscripts, the Samaritan Pentateuch, and Septuagint have `The word of'\thinspace''} to Moses,\ca{Ms \sampen\ + \Hebrew{לאמר} cf 6}{a Hebrew manuscript and the Samaritan Pentateuch add ``saying,'' compare verse~6} ``Go up to this Mount Abarim and see the land that I have given to the children of Israel.%%
  \verse{27:13} You shall see it, and you also shall be gathered to your people like your brother, Aaron,%%
  \verse{27:14} because you've been obstinate against\understood\ My word in the Sin Desert in the strife of the congregation~--- to sanctify me\ca{\peshitta(\vulgate) \textit{wl' qd\v stwnnj} = \Hebrew{אוֺתִי} \Hebrew{וְלֹא־קִדַּשְׁתֶּם} cf \caref{Dt}{32}{51}; \septuagint\ + \Greek{οὐχ ἡγιάσατ με}}{the Peshitta (and Vulgate) have ``and have not sanctified me,'' compare \vref{Deut}{32}{51}; the Septuagint has ``xxxx''} at the waters before their eyes'' (them being\lit{they} the waters of Meribah in Kadesh in the Sin Desert).%%
  
  \verse{27:15} Moses spoke to the \textsc{Lord}, saying,%%
  \verse{27:16} ``O \textsc{Lord}, God of the spirits of all flesh, appoint a man over the congregation%%
  \verse{27:17} who will go out before them, who will come in before them, who will take them out, who will bring them in, so that the congregation of the \textsc{Lord} won't be like sheep without a shepherd.''%%
  \verse{27:18} The \textsc{Lord} said to Moses, ``Take Joshua, son of Nun, a man\ie{Joshua} in whom is the Spirit, and lay your hand on him\halot{xxxx}{(in consecration)}%%
  \verse{27:19} and cause him to stand before Eleazar the priest and \lit{before}the entire congregation, and appoint\alt{command, order, direct} him before their eyes.%%
  \verse{27:20} Put your honor on him so that the entire congregation of the children of Israel will obey him.%%
  \verse{27:21} He shall stand before Eleazar the priest, and you shall consult the Urim for judgment about him before the \textsc{Lord}: at his word they shall go out, at his word they shall come in~--- he and all of the children of Israel with him, even the entire congregation.''%%
  \verse{27:22} Moses did as the \textsc{Lord} had commanded him, and he took Joshua and caused him to stand before Eleazar the priest and the entire congregation,%%
  \verse{27:23} and laid his hands on him and commanded him, as the \textsc{Lord} had spoken\ca{nonn Mss \septuagint\targum\super{J}\vulgate\ \Hebrew{צִוָּה} cf 22a}{several Hebrew manuscripts, the Septuagint, Targum, and Vulgate have ``commanded,'' compare verse~22} through\lit{by the hand of} Moses.\ca{\missing\ \vulgate; nonn Mss \septuagint\targum\super{J} \Hebrew{אֶת־מ׳} cf \super{b}; \sampen\ + \Hebrew{אֵלָיו} \Hebrew{וַיֹּאמֶר} et \caref{Dt}{3}{21 b.\ 22}}{missing in the Vulgate; several Hebrew manuscripts, the Septuagint, and Targum have ``to Moses,'' compare the previous [\textsc{ca}] footnote; the Samaritan Pentateuch adds ``and He said to him'' and \vref{Deut}{3}{21--22}}%% xxxx: 21 b. 22? Is that "21 and the beginning of 22"?
\end{inparaenum}
