\section{Numbers 5}\label{Numbers 5}
\heading{How to deal with lepers~--- repentance necessary for forgiveness~--- dealing with women accused of immorality}
\begin{enumerate}[align=center]
    \verse{Numbers^5:1} %%
    \verse{Numbers^5:2} %%
    \verse{Numbers^5:3} %%
    \verse{Numbers^5:4} %%
    \verse{Numbers^5:5} %%
    \verse{Numbers^5:6} %%
    \verse{Numbers^5:7} %%
    \verse{Numbers^5:8} %%
    \verse{Numbers^5:9} %%
    \verse{Numbers^5:10} %%
    \verse{Numbers^5:11} And the \textsc{Lord} spake unto Moses, saying,%%
    \verse{Numbers^5:12} ``Speak unto the children of Israel and say unto them: `If any man's wife go astray and trespasses against him%%
    \verse{Numbers^5:13} in that a man lies sensually with her and it is hidden from the eyes\footnote{kept from the knowledge} of her husband~--- kept secret~--- and she is defiled, but there is no witness against her, neither is she caught in the act:%%
    \verse{Numbers^5:14} if the spirit of jealousy is passed over him and he suspects her of infidelity and she is defiled, or if the spirit of jealousy passes over him and he suspects her of infidelity and she is not defiled:\footnote{E.g., if he's jealous, whether or not she's defiled, the following needs to happen.}%%
    \verse{Numbers^5:15} the man shall bring his wife before the priest. He\footnote{the husband} shall bring an offering for her (a tenth of an ephah of barley wheat), but he shall not pour upon it any oil neither will he place incense upon it since it is a memorial offering to bring to mind\footnote{remembrance} wrong-doing.%%
    \verse{Numbers^5:16} %%
    \verse{Numbers^5:17} The priest shall take holy water in a jar of clay and from the dust that is on the floor of the Tabernacle. And the priest shall take it and put it\footnote{i.e., the dust} in the water.%%
    \verse{Numbers^5:18} And the priest shall make the woman stand before the \textsc{Lord}.\footnote{In front of the Holy Place. It is unknown whether or not she is allowed into the grounds of the Tabernacle complex.} And he will uncover the woman and he shall place in her hands the memorial offering as an offering of jealousy. XXXX%%
    \verse{Numbers^5:19} And the priest shall place her under oath and say to the woman, ``If no man has laid with thee, and if thou hast not gone astray to be unclean, you will be free from these waters of bitterness that caused the curse.\footnote{i.e., to take effect}%%
    \verse{Numbers^5:20} But if you have gone astray XXXX (under?) your husband, or if you have defiled yourself in that a man other than your husband has lain with you,\footnote{Lit., if a man other than your husband has given you his line}%%
    \verse{Numbers^5:21} the priest shall charge the woman\footnote{place the woman under oath} with an oath of cursing and the priest shall say to the woman, `May the \textsc{Lord} make you a curse and an oath amongst your people.' And XXXX your thigh to waste\footnote{fall} away and your belly to swell.%%
    \verse{Numbers^5:22} %%
    \verse{Numbers^5:23} And the priest shall write these words on a scroll and he shall wash them out into the bitter waters.%%
    \verse{Numbers^5:24} He will have the woman drink the bitter waters that cause the curse and the waters will enter into her.%%
    \verse{Numbers^5:25} And the priest shall take the offering of jealousy from the hand of the woman and he shall wave the offering before the \textsc{Lord} and he shall bring it\footnote{i.e., the offering} to the altar.%%
    \verse{Numbers^5:26} And the priest shall take the handful XXXX from the altar as her memorial offering and he will burn it on the altar, and afterwards he will have the woman drink the waters.\footnote{Understood: the waters of bitterness.}%%
    \verse{Numbers^5:27} And he shall cause her to drink the waters of bitterness if she has transgressed in that she is defiled. And the waters of bitterness that caused the curse shall enter into her and her stomach shall swell and her uterus shall waste away and she shall be a curse in the midst of her people.%%
    \verse{Numbers^5:28} And if the woman is not defiled in that she acted unfaithfully, but she is clean, she shall be guiltless,\footnote{clean} and she, conceiving, will give birth to a child.\footnote{i.e., eventually}%%
    \verse{Numbers^5:29} This is the law concerning those suspected of infidelity (when the wife goes astray while married to her husband and is defiled).%%
    \verse{Numbers^5:30} %%
    \verse{Numbers^5:31} Then the man is free from wrong-doing, but the woman will pay for her iniquity.%%
\end{enumerate}
