\section{Isaiah 2}\label{Isaiah 2}
\heading{Isaiah sees in vision the latter day temple, the gathering of Israel, the Millennium~--- the proud to be humbled at the Savior's Second Coming}
\begin{enumerate}[align=center]
    \verse{Isaiah^2:1} The thing that Isaiah the son of Amoz foresaw concerning Judah and Jerusalem.%
    % : represents a tab. There are no colons in this chapter.
    \verse{Isaiah^2:2} ::And in the last days:the mountain of the \textsc{Lord}'s house will be established%
    
    ::in the tops of the mountains.:It shall be lifted up above the hills.%
    
    ::All nations shall flow\footnote{Like a river} unto it.\verse{Isaiah^2:3} Tons of people will walk to it and say,%
    
    ::``Come, and let us ascend unto the \textsc{Lord}'s mountain~--- :to the house of the God of Jacob.%
    
    ::He will teach us of His ways.:We will walk in His paths\footnote{Theologically it should be ``path''}%
    
    ::For the law\footnote{teaching, instruction. Traditionally rendered ``law''} goes forth from Zion:and the word of the \textsc{Lord} from Jerusalem.''%
    \verse{Isaiah^2:4} He shall judge among the nations;: He's arbitrates between many people.%
    
    ::They'll forge ploughshares from their swords:and pruning hooks from their spears.%
    
    ::One nation shall not lift its sword against another,:neither shall they learn warfare anymore.%
    \verse{Isaiah^2:5} ::Come, O house of Jacob, that we may walk in the \textsc{Lord}'s light.%
    \verse{Isaiah^2:6} :For you have left your people~--- the house of Jacob~--- to themselves%
    
    :because they've been filled\footnote{Possibly missing ``with superstition.'' i.e., ``filled with superstition from the east.'' There is no expressly-stated object} from the east~--- :they're sorcerers like the Philistines.%
    
    :They please themselves\footnote{Or ``clasp hands'' or ``make sufficient''} with foreigner's\footnote{Outsiders of Israel~--- foreigners, infidels, pagans, etc.} children.\footnote{This is to be understood in a sexual context}%
    \verse{Isaiah^2:7} :Their land is full of silver and gold~--- :there is no end to their treasures.%
    
    :Their land is full of horses~--- :there is no end to their chariots.%
    \verse{Isaiah^2:8} :Their land is full of idols%
    
    :They bow down before the work of their own hands~--- :that which their fingers have made.%
    \verse{Isaiah^2:9} The low shall be bowed down; the haughty humbled. :Don't forgive them.%
    \verse{Isaiah^2:10} Enter into a boulder, :hide in the dust,%
    
    :from before the \textsc{Lord}'s face and the glory of His majesty.
    \verse{Isaiah^2:11} %
    \verse{Isaiah^2:12} %
    \verse{Isaiah^2:13} %
    \verse{Isaiah^2:14} %
    \verse{Isaiah^2:15} %
    \verse{Isaiah^2:16} \footnote{See Appendix \ref{app:isa-2-16}}Upon the ships of Tarshish\footnote{Going to Tarshish. Either in Asia Minor (where Paul was from) or in present-day Spain} :and upon the ships all the beautiful vessels.\footnote{excellent ships}%
    \verse{Isaiah^2:17} %
    \verse{Isaiah^2:18} %
    \verse{Isaiah^2:19} %
    \verse{Isaiah^2:20} %
    \verse{Isaiah^2:21} %
    \verse{Isaiah^2:22} %
\end{enumerate}
