\heading{36}{Rabshakeh comes to besieged Jerusalem, blasphemes against the Lord, implores the people to not listen to King Hezekiah}

\begin{inparaenum}
  \verse{36:1} In the fourteenth year of King Hezekiah,\ie{his reign} king of Judah, the Assyrian king, Sennacherib, came up against all the fortifications of Judah and conquered them.%%
  \verse{36:2} The king of Assyria sent Rabshakeh\ed{See further in Appendix~\ref{app:rabshakeh}}\ed{This is either his literal name or it should be rendered ``(his) great commander.''} with a strong force from Lachish to Jerusalem unto King Hezekiah. He stood by the aqueduct of the upper pool on the highway of the fuller's field.%%
  \verse{36:3} He then came unto Eliakim, son of Hilkijah (who's over the household affairs), Shebna the scribe, and Joah the son of Asaph (the clerk).%%
  \verse{36:4} Rabshakeh said to them, ``Please say to Hezekiah, `Thus says the great king, the king of Assyria, ``What is this confidence in which you trust?%%
  \verse{36:5} You say, `It's just lip service,'\lit{a word for the lips} but there is strength for war. Now, who do you rely on that you have revolted against me?\ie{who do you trust that you feel comfortable in revolting against me?}%%
  \verse{36:6} You trust in the staff of a broken reed~--- on Egypt~--- whom if a man leans it goes into his hand and pierces it: such is Pharaoh, king of Egypt, to his allies.%%
  \verse{36:7} If you say to me, `We rely upon the \textsc{Lord} our God'~--- is it not he whose high places and altars Hezekiah has removed, saying to Judah and Jerusalem, `You shall worship before this altar'?''%%
  \verse{36:8} And now, I pray thee, make a bargain that my master, the king of Assyria, and I will give you up to two thousand horses if you're able to put riders on them.\ed{He knows they don't have that many riders and is taunting them with the Assyrian's superiority.}%%
  \verse{36:9} How then will you turn away the face of the least of my master's servants when you trust in Egypt for chariots and horsemen?%%
  \verse{36:10} And now, have I come up against this land without the \textsc{Lord} to destroy it? The \textsc{Lord} said to me, ``Go up against this land and destroy it.''\thinspace'\thinspace''%%
  \verse{36:11} Eliakim, Shebna, and Joah said to Rabshakeh, ``Please speak to your servants in Aramaic\ed{Aramaean, Syriac}\lit{Syrian language} because we can understand it. But don't speak to us in Hebrew\lit{the Jewish language} while in earshot of the people on the wall.''%%
  \verse{36:12} And Rabshakeh said, ``Did my master send me to speak these words to just you and your master and not to the men who sit on the wall so that they can eat their own feces and drink their own piss with you?''\ed{Remember that they are besieged.}%%
  \verse{36:13} Rabshakeh stood and cried in Hebrew with a loud voice\ed{So that all the people can hear and understand him, hence why they wanted him to speak Aramaic. This is a pretty diplomatically dick move.} and said, ``Hear the words of the great king, the king of Assyria!%%
  \verse{36:14} Thus says the king, `Don't let Hezekiah deceive you; he is not able to deliver you.%%
  \verse{36:15} And don't let Hezekiah make you trust in the \textsc{Lord}, saying, ``The \textsc{Lord} shall certainly deliver us. He will not let this city be given into the hand of the king of Assyria.''%%
  \verse{36:16} Don't listen to Hezekiah.' Thus says the king of Assyria, `Make a deal with me: come out to me and everyone shall eat of his own vine and his own fig tree and drink from their own cisterns%%
  \verse{36:17} until I come and take you away to a land like your own land~--- a land of corn and wine, a land of bread and vineyards.%%
  \verse{36:18} Don't let Hezekiah persuade you, saying, ``The \textsc{Lord} will deliver us.'' Have any of the Gentile gods delivered their land from the hand of the king of Assyria?%%
  \verse{36:19} Where are the gods of Hamath and Arpad? Where are the gods of Sepharvaim? Have they delivered Samaria from out of my hand?%%
  \verse{36:20} Who are they among the gods of the land that have delivered their own land from out of my hand that the \textsc{Lord} should deliver Jerusalem from out of my hand?''%%
  \verse{36:21} They kept silent and did not answer him a word for the king's commandment was, saying, ``Don't answer him.''%%
  \verse{36:22} Eliakim (the son of Hilkijah) who is in charge of the house, Shebna the scribe, and Joah the son of Asaph the clerk came in to Hezekiah with rent clothes and reported to him the words of Rabshakeh.%%
\end{inparaenum}
