\heading{52}{Isaiah calls Israel to repentance~--- they shall be made great by the Lord}

\begin{inparaenum}
  \pa \verse{52:1} Awake, awake!%
  \halot{עור}{\textbf{stir onesf.\ up}, \textbf{be awake}, \textbf{astir}, \textbf{lively}}
  Put on\pa your strength, O Zion!
  
  \pb Put on your beautiful garments,\pa O Jerusalem, the holy city.%%
  
  \pb For from here on out, no more shall enter into you\pa the uncircumcised and unclean.%%
  
  \pa \verse{52:2} Shake the dust off yourself. Arise,\pa sit down, O Jerusalem.%%
  
  \pb Loose yourselves from the chains on your neck,\pa O captive daughter of Zion.\bigskip%%
  
  \pc \verse{52:3} For thus saith the \lord, ``You've been sold for nothing,\ie{at no cost, for nothing} but you shall be redeemed without money.''%%
  
  \pc \verse{52:4} For thus saith the \lord\ God, ``My people went down to Egypt at first to sojourn%%
  
  \pc there, but lately Assyria has oppressed them without reason. \pa \verse{52:5} So now, what do I have here?'' declares the \lord,%%
  
  \pc ``For My people have been taken away for nothing! Their rulers cause a howling,'' declares the \lord, ``and will continually%%
  
  \pc scorn\alt{reject} My name.\pa \verse{52:6} Therefore shall My people know My name. Therefore, in that day,%%
  
  \pc \ed{Understood: they shall know} that I am the one who says, ``Here I am.''\lit{Behold! Me.}%%
  
  \pa \verse{52:7} How beautiful upon the mountains\pa are the feet of the one announcing\alt{bringing} tidings,\ed{Generally good news, but sometimes bad news; however, the context here is clear that it's good news.}%%
  
  \pb that declares peace;\alt{makes peace known;} who announces good tidings,\pa who makes salvation known.%%
  
  \pb Who says to Zion:\pa ``Your God reigns!''%%
  
  \pa \verse{52:8} Your watchmen lift up their voice,\pa together they sing aloud.%%
  
  \pb They shall see eye to eye\pa when the \lord\ returns to Zion.%%
  
  \pa \verse{52:9} Be cheerful, rejoice together,\pa O ruins\alt{wasted places} of Jerusalem,%%
  
  \pb for the \lord\ will comfort His people.\pa He has redeemed Jerusalem.%%
  
  \pa \verse{52:10} The \lord\ shall make bare His holy arm\pa in the eyes of all the nations.%%
  
  \pb All the ends of the earth shall see\pa our God's salvation.\bigskip%%
  
  \pb \verse{52:11} Depart! Depart! Get our from there!\pa Don't touch the unclean!%%
  
  \pc Go out from the midst of her. Be clean,\pa you who carry the vessels of the \lord.%
  \ed{This is referring to the temple vessels carried back from exile, so this is directed at the priests and Levites.}%%
  
  \pb \verse{52:12} For you shall not go out hastily,\pa neither shall you go out by flight,%%
  
  \pc because the \lord\ shall go before you~---\pa the God of Israel shall be your rearguard.\bigskip%%
  
  \pa \verse{52:13} My servant shall deal wisely.\alt{act prudently.}\pa He shall be exalted and lifted up~--- he shall be made very high.%%
  
  \pa \verse{52:14} As many as were astonished at You%
  \ca{1 c 2 Mss \peshitta\targum\ \Hebrew{עלָיו}}{1 with 2 Hebrew manuscripts (of the?) Peshitta and Targum have ``at Him''}%
  ~---%%
  
  \pb His appearance was marred by man,\pa and His form by the children of men\alt{children of Adam}~---%%
  \ca{frt tr ad fin 53,2}{[this line is] perhaps transposed to the end of \vref{Is}{53}{2}}%
  
  \pa \verse{52:15} so shall He prosper\alt{astonish}\lit{sprinkle} many nations:\pa kings shall shut their mouths concerning Him:%%
  
  \pb for that which was not told unto them, they've seen;\pa and that which they've not heard, they've understood.\alt{considered.}%%
\end{inparaenum}
