\heading{53}{Isaiah proclaims Christ and His atonement~--- the Lord shall be despised of men~--- He shall bear our griefs, afflictions, sins, and pain~--- He shall at last be exalted}

\begin{inparaenum}
  \pa \verse{53:1} Who has believed what we've heard?\pa Unto whom has the arm of the \lord\ been revealed?%
  \ed{The second hemistich is also a question: the \Hebrew{וְ} at the beginning of it is the sign that it's parallel with the first hemistich.}%
  
  \pa \verse{53:2} He shall grow up before Him like a tender plant~---\pa as a root from dry ground.%
  
  \pb He has no form, no majesty;\alt{honor}\pa when we see\alt{observe} Him, there is no beauty that we should desire Him.%
  \ie{see Him with any appreciation}%
  \ed{The critical apparatus from \vref{Is}{52}{14} says that the line ``His appearance was marred by man,\pa and His form by the children of men'' is transposed to here in the Peshitta and Targum.}%
  
  \pa \verse{53:3} He is despised and ostracized\alt{left} of men:\pa a man of sorrow,\alt{pain} who knows\alt{is acquainted with}\ed{Saying He \textit{knows} grief really personifies His pain and suffering.} grief.\alt{sorrow, sickness.}%
  
  \pb As one from whom we hid our%
  \ed{It's understood that it's either us or other people hiding their respective faces.}
  faces.\pa He is despised. We did not esteem Him.%
  
  \pb \verse{53:4} Surely\alt{Assuredly} He has borne our griefs\alt{infirmities, sicknesses}\pa and carried our sorrows.\alt{pains.}%
  
  \pb Yet we have regarded\alt{viewed, esteemed} Him stricken\alt{plagued}~---\pa smitten of God~--- and afflicted.%
  
  \pa \verse{53:5} He was pierced\alt{wounded} for our transgressions;\pa He was bruised for our iniquities.%
  
  \pb Upon Him was the punishment that brought us peace:\alt{The chastisement of our peace was upon Him:}\pa by His stripes\alt{bruises} we are healed.%
  
  \pa \verse{53:6} Like sheep we've all gone astray,%
  \alt{wandered off}
  \alt{All of us~--- like sheep~--- have gone astray,}%
  \pa everyone has turned their own way.\alt{each man has taken his own path.}%
  
  \pb The \lord\ has placed\alt{laid} upon Him\pa all of our iniquity.%
  
  \pb \verse{53:7} He was oppressed and he was afflicted,\pa yet He never opened His mouth.%
  
  \pb He was led\alt{brought} like a lamb to the slaughter.\pa Like a sheep is dumb before its shearers,%
  
  \pb so he didn't open His mouth.%
  
  \pa \verse{53:8} He was taken for\alt{from, by} oppression and judgment.\pa Who shall speak of\alt{talk about, declare, ponder} His descendants?%
  
  \pb For He has been cut off from the land of the living.\pa He is stricken\alt{plagued} by\alt{for} my people's transgression.%
  
  \pa \verse{53:9} He\ed{This is referring to some other actor, possibly the Father} appoints His\ie{the suffering servant} has given His grave with the wicked.\pa He is with the rich in death.%
  
  \pb However, He has done no violence;\pa neither is there deception\alt{deceit} in His mouth.%
  
  \pb \verse{53:10} The \lord\ delights to bruise Him and make Him suffer.\alt{sick}\pa Though you make His soul\alt{life} an offering for guilt,\alt{sin,}%
  
  \pb He shall see His offspring.%
  \cf{Abinadi's explanation in \vref{Mos}{15}{10--18}}
  His days shall be prolonged.\pa The pleasure of the \lord\alt{The \lord's good favor} shall prosper in His hand.%
  
  \pa \verse{53:11} Through hard work He shall be satisfied.%
  \alt{He shall see the labor of His soul.}
  \pa By\alt{Through} His knowledge%
  
  \pb shall my Righteous Servant instruct many in righteousness.\pa He shall bear their iniquities.%
  
  \pa \verse{53:12} Therefore, I shall give Him room\alt{a portion} among many.\pa He apportions spoil unto the mighty.%%
  
  \pb Because that He exposed\alt{poured out} His soul unto death\pa and He was numbered among the transgressors.%%
  
  \pb He hath borne the sins of many.\pa He has made intercession for transgressors.%
  \alt{He has interceded for those who transgress.}%
\end{inparaenum}
