\heading{7}{Syria, Ephraim, and the son of Remaliah conspire against the house of David~--- xxxx}

\begin{inparaenum}
  \verse{7:1} In the days of Ahaz, son of Jotham, son of Uzziah king of Judah, that Rezin the king of Syria and Pekah, the son of Remaliah, king of Israel, went up to Jerusalem to wage war against it, but he was not able to\ca{\qumran\super{a}\septuagint\peshitta\vulgate\ et \caref{2~R}{16}{5} \Hebrew{יָכְלוּ}}{Qumran, the Septuagint, the Peshitta, the Vulgate, and \vref{2~Ki}{16}{5} have ``they were not able to''} endure against it.\ca{\missing\ \caref{2~R}{16}{5}}{[``against it'' is] missing in \vref{2~Ki}{16}{5}}%%
  \verse{7:2} Is it told to the house of David, saying, ``Syria is in cahoots with Ephraim.'' xxxx%%
  \verse{7:3} The \textsc{Lord} said to Isaiah, ``Please go and meet Ahaz~--- you and your son Shear-jashub~--- to the end of the aqueduct\alt{conduit} of the upper pool, to the highway of the fuller's field.''%%
  \verse{7:4} Say to him, ``Take heed and be quiet! Don't let your heart be timid before these two, smoking firebrands\ed{used to stir a fire} because of the fierce anger of Rezin, Syria, and the son of Remaliah.%%
  \verse{7:5} Because that Syria counseled evil upon you, Ephraim and the son of Remaliah,\ie{all three counseling against the house of David} saying,%%
  \verse{7:6} `We're going up to Judah and shall vex\alt{harass, rouse} it: we shall rend it unto ourselves and cause a king to reign in its midst~--- the son of Tabeal.'\smallskip%%
  
  \pvab{\verse{7:7} Thus said the Lord \textsc{God},}{`It shall neither stand nor come to pass!'}%%
  
  \pvab{\verse{7:8} For Syria's head is Damascus;}{Damascus' head is Rezin.}%%
  
  \pvbb{Ephraim will be a broken people}{within sixty-five years.}%%
  
  \pvab{\verse{7:9} Ephraim's head is Samaria;}{Samaria's head is the son of Remaliah.}%%
  
  \pvbb{If you don't believe}{you shall surely not be established.''}%%
  
  \verse{7:10} The \textsc{Lord} again spoke to Ahaz, saying,%%
  \verse{7:11} ``Ask the \textsc{Lord} your God for a sign: ask to the depths of Sheol, raise it on high.''\ie{you can ask for whatever you want: full range}%%
  \verse{7:12} And Ahaz said, ``I will neither ask nor try\alt{test, task} the \textsc{Lord}.''%%
  \verse{7:13} So He said, ``Please hear,\alt{Hear now} O house of David: is wearying man such a small thing\alt{so trivial} that you will also weary my God?%%
  \verse{7:14} Therefore, the Lord Himself shall give you a sign: a young woman\ed{at least a woman of sexual maturity; it's from the Septuagint that we get the rendering ``virgin'' although it has that meaning in Hebrew. The Greek \Greek{παρθένος} and Latin \textit{virgo}, a woman who has not known man. In Hebrew, a young woman or maiden, whether or not she's known man} shall conceive and bring forth a son and name him Immanuel.\ed{this might not be Messianic}%%
  \verse{7:15} He shall eat butter and honey and thereby know to refuse evil and fixate on\alt{choose} good.%%
  \verse{7:16} For before the child knows to refuse the evil and choose the good, the land you're vexed with\alt{which you fear} shall be forsaken because of its two kings.%%
  \verse{7:17} The \textsc{Lord} will bring upon you, your people, and your father's house days which have not come since the day when Ephraim turned from Judah~--- even the king of Assyria.\smallskip%%
  
  \pvaa{\verse{7:18} In that day}%%
  
  \pvbb{the \textsc{Lord} will hiss for a fly}{that is in the extremity of the brooks of Egypt}%%
  
  \pvba{and for a bee that is in the land of Assyria.}%%
  
  \pvab{\verse{7:19} They've come. They're all rested}{in the desolate valleys and in the holes of the rocks,}%%
  
  \pvba{in the thorn bushes and in all the pastures.}%%
  
  \pvab{\verse{7:20} In that day the Lord will shave}{with a razor~--- which is hired from beyond the river\footnotemark\ by the king of Assyria~---}%%
  \ie{the Euphrates}%%
  
  \pvbb{the head, the hair of the legs\footnotemark}{and also the beard will he take away in that day.}%%
  \alt{feet}%%
  
  \pvab{\verse{7:21} In that day}{a man will nourish a calf and two sheep;}%%
  
  \pvab{\verse{7:22} from the abundance of milk yielded}{he shall eat butter.}%%
  
  \pvbb{For butter and honey shall be eaten}{by everyone who's left in the midst of the land.}%%
  
  \pvaa{\verse{7:23} In that day,}%%
  
  \pvbb{every place in which there were}{a thousand vines at a thousand silver pieces}%%
  
  \pvba{shall become briers and thorns.}%% What?
  
  \pvab{\verse{7:24} They shall come thither with bows and arrows}{because all the land has become briers and thorns.}%%
  
  \pvbb{\verse{7:25} All the mountains which have been cultivated with a hoe,}{the fear of brier and thorn shall not come there.}%%
  
  \pvba{It shall be for the sending forth of oxen and the treading of sheep.}%%
\end{inparaenum}
