\heading{1}{Naomi's husband and sons die~--- one of her daughters-in-law, Orpah, leaves~--- the other, Ruth, promises to stay}

\begin{inparaenum}
    \verse{1:1} In the days when the judges judged there was a famine in the land. A man of Bethlehem-Judah went to sojourn in the field of Moab\footnote{Moabite territory or country}~--- he, and his wife, and his two sons.%%
    \verse{1:2} And the name of the man~--- Elimelech; and the name of his wife~--- Naomi; and the names of his two sons~--- Mahlon and Chilion: Ephrathites from Bethlehem-Judah. And they entered the field of Moab and they stayed\footnote{continued, sojourned} there.%%
    \verse{1:3} And Naomi's husband, Elimelech, died, and she was left with\footnote{and} her two sons.%%
    \verse{1:4} And they took to\footnote{raised up unto} themselves Moabite wives, the name of the one, Orpah; and the name of the other, Ruth. And they stayed there about ten years.%%
    \verse{1:5} Both Mahlon and Chilion died. And the woman was bereft of her two boys and her husband.%%
    \verse{1:6} And\footnote{Then} she rose up with her daughters-in-law to return to the land of Moab, for she had heard that in the land of Moab the \textsc{Lord} had visited His people by giving them bread.%%
    \verse{1:7} She went out from the place where she was with her two daughters-in-law and they went on the road to return to the land of Judah.%%
    \verse{1:8} And Naomi said to her two daughters-in-law to go return to their mothers' houses. ``The \textsc{Lord} will do kindly with you according as ye have done with the dead and myself.%%
    \verse{1:9} The \textsc{Lord} will give you rest in your husbands' houses.'' Then she kissed them, and they lifted up their voices and wept.%%
    \verse{1:10} But they said to her, ``Surely we will return to thy people.''%%
    \verse{1:11} But Naomi said, ``Return, my daughters. Why would you go with me? Do I yet have sons in my womb that they may be husbands to you?%%
    \verse{1:12} Return, my daughters, for I have grown too old to have a husband \footnote{I have grown old from having a man.}. If I were to say I have hope, and if I were to have a man tonight, and have sons,%%
    \verse{1:13} would you wait for them until they were grown and keep yourselves from a husband? Not so, my daughters, for it grieveth me greatly for your sakes that the hand of the \textsc{Lord} is against me.''%%
    \verse{1:14} They lifted up their voices and wept again. And Orpah kissed her mother-in-law, but Ruth clung to her.%%
    \verse{1:15} And she\footnote{Naomi} said, ``Your sister-in-law\footnote{The Hebrew, \Hebrew{בּמת}, denotes a relationship by marriage, not necessarily that they are sisters.} is returning to her people and to her gods. Return after your sister-in-law.''%%
    \verse{1:16} And Ruth said, ``Do not ask me to abandon you nor to turn from following thee: for wherever you go, will I go; and where you lodge, I will lodge; your people shall be my people; and your God shall be my God;%%
    \verse{1:17} and where thou diest, will I die and be buried there: the \textsc{Lord} do to me, and more also, if anything but death parts me and thee.''%%
    \verse{1:18} When she saw how bold\footnote{steadfastly minded} she\footnote{Ruth} was to go with her, she ceased speaking with her.%%
    \verse{1:19} And the two of them went to Bethlehem. And when they got to Bethlehem the city was abuzz about them, saying, ``Is this Naomi?''%%
    \verse{1:20} And she said unto them, ``Don't call me Naomi. Call me Mara, for the Almighty hath treated me harshly.%%
    \verse{1:21} I left full, and empty the \textsc{Lord} returned me. Why would you call me Naomi since the \textsc{Lord} hath afflicted me and the Almighty hath broken me in pieces?''%%
    \verse{1:22} And Naomi returned with Ruth the Moabitess, her daughter-in-law, with her to the land of Moab and they entered into Bethlehem in the beginning of barley harvest.%%
\end{inparaenum}
