\heading{3}{Naomi has Ruth entreat herself to Boaz, their levir~--- Boaz tells Ruth that there is another levir who has more immediate responsibility over her than he~--- he comes up with a plan to redeem her himself~--- she spends the night and he sends her off}

\begin{inparaenum}
    \verse{3:1} So Naomi, her mother-in-law, said to her, ``My daughter, rest that it may be well for you.%%
    \verse{3:2} Is not Boaz our kinsman among whose maidservants you were with? He is threshing barley at the threshing floor tonight.%%
    \verse{3:3} Wash and anoint thyself, put your clothes on, and get down to the threshing floor, but don't make thyself known to the man until he's done eating and drinking.%%
    \verse{3:4} And when he lies down you shall know the place where he lieth. And you shall enter\footnote{go in} and uncover his feet and lie down. And he shall tell you that which you shall do.''%%
    \verse{3:5} And she\footnote{Ruth} said unto her, ``All that thou hast said unto me I will do.''%%
    \verse{3:6} And she went down to the threshing floor and she did just about everything that her mother-in-law told her.%%
    \verse{3:7} And Boaz ate and drank and was happy. So he entered to lie down at the edge of the heap of grain. And she entered in quietly\footnote{secretly} and uncovered his feet and lied down.%%
    \verse{3:8} And night came, and the man was shocked and turned and there was a woman lying at his feet.%%
    \verse{3:9} And he said, ``Who is there?'' And she said, ``I am Ruth, thy handmaiden. Spread the hem of thy garment over thy maidservant for you have the right of a levirite.''%%
    \verse{3:10} And he said, ``Blessed be thou of the \textsc{Lord}, my daughter. You've done a finer act of kindness at the end than in the beginning in not going after the young men,\footnote{KB: fully grown, vigorous, still unmarried.} be they rich or poor.%%
    \verse{3:11} Now, my daughter, fear not\footnote{do not be afraid} for all I have said unto you, that will I do. All of my people think that thou art a virtuous woman.%%
    \verse{3:12} And now, truly I am a levir; however, there's a levir nearer than I.%%
    \verse{3:13} Stay the night. If he redeems you in the morning, great. If he's not inclined to redeem you, then as the \textsc{Lord} lives, I will redeem you. Lie down till the morning.''%%
    \verse{3:14} She laid at his feet till morning and got up before you can tell who's who.\footnote{lit., before you can tell a man from another.} He said, ``Don't let it be known that a woman came to the threshing floor.''%%
    \verse{3:15} He said, ``Give me the scarf from off your head.'' And he measured out for her six measures of barley. Then he\footnote{The Septuagint version can be translated as either ``he'' or ``she.'' The BHS leaves no such ambiguity and can only be rendered as ``he.''} went into the city.%%
    \verse{3:16} And she came to her mother-in-law and she\footnote{Naomi} said, ``How did it go, my daughter?'' And then she told her everything that the man had done for her,%%
    \verse{3:17} saying\footnote{Literally, ``and she said''}, ``He gave me six measures because he said to me, `Don't go to your mother-in-law empty-handed.'\thinspace''%%
    \verse{3:18} Then she said, ``Sit, my daughter, until you know how the thing will play out, for the man will not be quiet until he has finished the matter today.''%%
\end{inparaenum}
